### 2016-09-12 paper.tex
Initial commit

% Template for ICASSP-2017 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{algorithm2e}
\usepackage{pbox}
\usepackage{multirow}

% Example definitions.
% --------------------
\def\x{{\mathbf x}}
\def\L{{\cal L}}

% Title.
% ------
\title{INVARIANT REPRESENTATIONS FOR SPEECH RECOGNITION}
%
% Single address.
% ---------------
%\name{Dmitriy Serdyuk $^{\dagger}$ \qquad Kartik Audhkhasi$^{\star}$ \and Philemon Brakel$^{\dagger}$ \and Bhuvana Ramabhadran$^{\star}$  \and Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
%\address{$^{\dagger}$ Universit\'e de Montr\'eal, $^{\star}$ IBM}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
\twoauthors
  {Dmitriy Serdyuk $^{\dagger}$\thanks{Dmitriy Serdyuk performed the work
  while at IBM}, Kartik Audhkhasi$^{\star}$, Philemon Brakel$^{\dagger}$, Bhuvana Ramabhadran$^{\star}$, Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
  {$^{\dagger}$ Universit\'e de Montr\'eal \\
  $^{\star}$ IBM}
  {}
  {}

\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
    Robustness of a speech recognition system is a challenging task and modern
    automatic speech recognition (ASR) systems need to be robust to different recording conditions,
    microphone type, accents, and noisy recording environment. We mitigate
    this problem in a fashion similar to recent research on image generation using
    Generative Adversarial Networks and domain adaptation ideas extending
    adversarial gradient-based training.
    This work focuses on investigation of neural architectures which produce
    representations invariant to certain conditions as gender and noise type. We
    experimentally evaluate proposed architectures on the acoustic modelling task
    for speech recognition on the Aurora4 dataset.
\end{abstract}
%
\begin{keywords}
speech recognition, deep neural networks, invariant representations, adversarial training
\end{keywords}
%
\section{Introduction}
\label{sec:intro}

    The main idea of this work is that instead of training a robust system,
    it is possible to train a system which will be invariant to these 
    conditions. This method of training requires the labels in the dataset
    which indicate which condition a particular recording was taken in, the 
    test data doesn't need to contain these labels. The method was proposed in
    the context of image classification in~\cite{ganin2014unsupervised} and 
    will be explained in Section~\ref{sec:gradient-reverse} and the application 
    to speech recognition is discussed
    in Section~\ref{sec:invariant-speech}. It is tightly connected to the idea
    of generative adversarial networks (GAN) proposed 
    in~\cite{goodfellow2014generative} thus we start with the discussion
    on this type of generative networks in Section~\ref{sec:gans}

    %TODO

\section{Generative Adversarial Networks}
\label{sec:gans}
    One of the most insightful recent advantages in the area of generative models
    is the models of the family of generative adversarial networks (GAN).
    A GAN consists of two networks: generator and discriminator. The generator 
    network $G$ has an
    input of randomly generated vector and is asked to produce an image 
    similar to the images in the training set. The discriminator network $D$
    can either receive a generated image from the generator $G$ or an image
    from the training set, the task of the discriminator is to distinguish
    between a generated image and a ``real'' image taken from the dataset. Thus
    the discriminator is just a classifier network with a sigmoid output
    and can be trained with the gradient methods (see 
    Section~\ref{sec:optimization}). The main idea of the generative adversarial
    training is that it is possible to propagate the gradient of the discriminator
    network further to the the generator network staking them on top of each other.
    This leads to an efficient training procedure and GANs achieve the state of the
    art that time on handwritten digit generation MNIST dataset.

    In other words, these two networks are competing with each other the 
    generator is trying to deceive a the discriminator network while it tries
    to do its best to recognize if there was a deception. This game theoretical
    similarity gave a name of adversarial networks. The objective of adversarial
    training is following
    \begin{align*}
        \min_G \max_D V(D, G) = \mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
          \mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
    \end{align*}

    One of the properties of GAN networks that there is no straightforward way 
    to evaluate the quality of image generation process since the GAN doesn't 
    output the loglikelihood of its samples. The paper estimated the likelihood
    using the Parzen windows estimator.

    In practice, GANs might be hard to train due to the fact that in a case when 
    the discriminator is more powerful that the generator, the later doesn't 
    receive any gradient thus can not be improved. Usually this problem is solved
    using a smart schedule which stops learning the discriminator as soon as it
    is too far ahead of the generator. Other possible solutions include ``soft''
    targets bounded at some constant rather than between 0 and 1 and using a 
    different proxy for the cost.

\section{Gradient Reverse Method}
\label{sec:gradient-reverse}

    A work~\cite{ganin2014unsupervised} proposes a method of training a network 
    which can be adapted to new domains. The training set consists of the images
    labeled with classes and the separate domain labels. The test set doesn't need
    to have the domain labels. For example, it may be first domain
    consisting of MNIST digits and second domain of MNIST digits on a different
    background. The network has a fork-like structure: the image is fed to the
    first network which produces a hidden representation $h$, then there are two
    networks branching of the representation $h$: a domain classifier network and 
    a target classifier. The goal of the training is to produce the hidden 
    representation invariant to the domain labels so that information doesn't 
    interfere the target classifier which is used in the test time.

    The network is trained with three goals: the hidden representation $h$ should
    be helpful for the target classifier, but harmful for the domain classifier,
    the domain classifier should have a good classification accuracy. More 
    formally, we define the objective as
    \begin{equation}
        L = L(\hat{y}, y; \theta_R, \theta_C) + 
        \alpha L(\hat{d}, d; \theta_D) -
        \beta L(\hat{d}, d; \theta_R),
    \end{equation}
    where $y$ is the ground truth class, $d$ is the domain label, corresponding
    hatted variables are the network predictions, $\theta$ are the network 
    parameters, and $R, C, D$ are the subsets of parameters in the representation,
    classifier and domain adaptation networks respectively. The hyper-parameters
    $\alpha$ and $\beta$ denote the influence of every part of the cost.

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\bibliographystyle{IEEE}
\bibliography{strings,refs}

\end{document}

### 2016-09-13 paper.tex
Fix title
Do not use second bibtex file
Update abstract

@@
 % ------
-\title{INVARIANT REPRESENTATIONS FOR SPEECH RECOGNITION}
+\title{INVARIANT REPRESENTATIONS FOR NOISY SPEECH RECOGNITION}
 %
@@
     Robustness of a speech recognition system is a challenging task and modern
-    automatic speech recognition (ASR) systems need to be robust to different recording conditions,
+    automatic speech recognition (ASR) systems need to be resilient to different recording conditions,
     microphone type, accents, and noisy recording environment. We mitigate
@@
     Generative Adversarial Networks and domain adaptation ideas extending
-    adversarial gradient-based training.
+    adversarial gradient-based training. A recent work of Ganin et al. proposes to
+    use adversarial training procedure for image domain adaptation making an intermediate
+    representation to improve the target performance and at the same time
+    to deteriorate the domain classifier performance, which is a separate neural
+    network taking the intermediate representation as the input.
+
     This work focuses on investigation of neural architectures which produce
-    representations invariant to certain conditions as gender and noise type. We
+    representations invariant to certain conditions as the noise type for speech
+    recognition. Similar
+    to the Ganin et al. work we train the intermediate representation to be invariant
+    to noise conditions training three neural networks: the encoder, the recognizer, and the
+    noise classifier to produce the representation produced by the encoder to decrease
+    the performance of the classifier making the representation invariant to noise conditions. We
     experimentally evaluate proposed architectures on the acoustic modelling task
-    for speech recognition on the Aurora4 dataset.
+    for speech recognition on a popular benchmark Aurora4.
 \end{abstract}
@@
 \bibliographystyle{IEEE}
-\bibliography{strings,refs}
+\bibliography{refs}
 

### 2016-09-13 paper.tex
bit of extra intro text

--- 2016-09-12 paper.tex
@@
 \label{sec:intro}
+    One of the most challenging aspects of automatic speech recognition (ASR),
+    is the mismatch between the training and testing conditions. During
+    testing, a system may encounter new recording conditions, microphone types,
+    accents and types of background noises. Even if all testing conditions also appeared
+    in the train data, their relative distribution is likely to be
+    different. For this reason, it's important to develop systems that are
+    robust towards changes in recording conditions in general.
+
+    While many types of model and feature based adaptation methods like MLLR and fMLLR
+    \cite{leggetter1995maximum} have been proposed, most of these methods have
+    been designed with GMM-HMM systems in mind. Since the enormous success of
+    Deep Neural Network (DNN) systems for ASR \cite{hinton2012deep}, GMM-based
+    model adaptation is not possible anymore, while feature-based methods may
+    not be optimal anymore.
+    Moreover, the discriminative gradient-based optimization used to train
+    DNNs, in combination with the ability of DNNs to learn highly non-linear
+    feature transformations, allows for more flexibility when it comes to the
+    construction of training objectives that promote domain invariant
+    representations.
 
@@
     training is that it is possible to propagate the gradient of the discriminator
-    network further to the the generator network staking them on top of each other.
+    network further to the generator network staking them on top of each other.
     This leads to an efficient training procedure and GANs achieve the state of the
@@
     In other words, these two networks are competing with each other the 
-    generator is trying to deceive a the discriminator network while it tries
+    generator is trying to deceive the discriminator network while it tries
     to do its best to recognize if there was a deception. This game theoretical
@@
     to evaluate the quality of image generation process since the GAN doesn't 
-    output the loglikelihood of its samples. The paper estimated the likelihood
+    output the log-likelihood of its samples. The paper estimated the likelihood
     using the Parzen windows estimator.

### 2016-09-13 paper.tex
Add aknowledgements and fix makefile

--- 2016-09-13 paper.tex
@@
   {Dmitriy Serdyuk $^{\dagger}$\thanks{Dmitriy Serdyuk performed the work
-  while at IBM}, Kartik Audhkhasi$^{\star}$, Philemon Brakel$^{\dagger}$, Bhuvana Ramabhadran$^{\star}$, Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
+  while at IBM}, Kartik Audhkhasi$^{\star}$, Phil\'emon Brakel$^{\dagger}$, Bhuvana Ramabhadran$^{\star}$, Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
   {$^{\dagger}$ Universit\'e de Montr\'eal \\
@@
 
+\section{Related Work}
+
+\section{Experiments}
+
+\section{Conclusion}
+
+\section*{Acknowledgments}
+
+We would like to thank Yaroslav Ganin, David Warde-Farley for insightful discussions,
+developers of Theano~\cite{2016arXiv160502688short}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 
+for great instruments.
+
 % References should be produced using the bibtex program from suitable

### 2016-09-13 paper.tex

@@
 \section{Experiments}
+We experimentally evaluate the approach discussed in Section~\ref{sec:gradient-reverse} 
+on a noisy speech recognition dataset Aurora-4~\cite{parihar2002aurora}. This dataset
+is based on the Wall Street Journal (WSJ0 is available at LDC). Additive noise of 
+six categories (airport, babble, car, restaurant, street, train) was introduced to the 
+clean data. We used mel-filterbank features with their deltas and delta-deltas spliced
+over 5 frames on each side making 11 consecutive frames and 1320 input features. 
+We did not use any speaker normalization of input features. We pretrained 5 layer
+speech recognition network with 2048 rectified linear units at every layer using Adam 
+optimization algorithm for 5 epochs. After pretraining, we turned on the invariance term.
+
+First expriment we performed is the gender invariance. We added a 2 layer network 
+branching off the last layer of the pretrained speech recognition network and 
+trained it until convergence with SGD. Although, the performance remained unchanged,
+the t-SNE visualization on Figure~\ref{fig:tsne} demonstrates that the hidden units
+are less separated in the projected space.
+
+\begin{figure}
+    \centering
+    %TODO: add image
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{tsne-original.png}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{tsne-invariant.png}
+    \end{subfigure}%
+    \caption{t-SNE visualization of the last hidden layer. On the left, the baseline
+    model. On the right, the model trained with the adversarial objective. Colors 
+    represent genders.}
+    \label{fig:tsne}
+\end{figure}
 

### 2016-09-13 paper.tex
Add experiment description and table

@@
 % ---------------
-%\name{Dmitriy Serdyuk $^{\dagger}$ \qquad Kartik Audhkhasi$^{\star}$ \and Philemon Brakel$^{\dagger}$ \and Bhuvana Ramabhadran$^{\star}$  \and Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
-%\address{$^{\dagger}$ Universit\'e de Montr\'eal, $^{\star}$ IBM}
-%
-% For example:
-% ------------
-%\address{School\\
-%	Department\\
-%	Address}
-%
-% Two addresses (uncomment and modify for two-address case).
-% ----------------------------------------------------------
-\twoauthors
-  {Dmitriy Serdyuk $^{\dagger}$\thanks{Dmitriy Serdyuk performed the work
+\name{Dmitriy Serdyuk $^{\dagger}$\thanks{Dmitriy Serdyuk performed the work
   while at IBM}, Kartik Audhkhasi$^{\star}$, Phil\'emon Brakel$^{\dagger}$, Bhuvana Ramabhadran$^{\star}$, Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
-  {$^{\dagger}$ Universit\'e de Montr\'eal \\
+\address{$^{\dagger}$ Universit\'e de Montr\'eal\\
          $^{\star}$ IBM}
-  {}
-  {}
 
@@
     experimentally evaluate proposed architectures on the acoustic modelling task
-    for speech recognition on a popular benchmark Aurora4.
+    for speech recognition on a popular benchmark Aurora-4.
 \end{abstract}
@@
 on a noisy speech recognition dataset Aurora-4~\cite{parihar2002aurora}. This dataset
-is based on the Wall Street Journal (WSJ0 is available at LDC). Additive noise of 
+is based on the Wall Street Journal corpus (WSJ0 is available at LDC). Additive noise of 
 six categories (airport, babble, car, restaurant, street, train) was introduced to the 
 clean data. We used mel-filterbank features with their deltas and delta-deltas spliced
-over 5 frames on each side making 11 consecutive frames and 1320 input features. 
+over 5 frames on each side making 11 consecutive frames and 1320 input features in total. 
 We did not use any speaker normalization of input features. We pretrained 5 layer
@@
 
-First expriment we performed is the gender invariance. We added a 2 layer network 
+The first expriment we performed is the gender invariance. We added a 2 layer network 
 branching off the last layer of the pretrained speech recognition network and 
 trained it until convergence with SGD. Although, the performance remained unchanged,
-the t-SNE visualization on Figure~\ref{fig:tsne} demonstrates that the hidden units
+the t-SNE visualization in Figure~\ref{fig:tsne} demonstrates that the hidden units
 are less separated in the projected space.
+
+Next, we trained a model for the noise condition invariance. We used an architecture
+similar to one described above with a binary invariance classifier target denoting 
+whether the utterance is clean or noisy. We oversampled the noisy utterances to 
+address the fact that the labels are not ballanced, the clean subset has approximately 4400
+when every noisy condition subset contains only 443 utterances. The results are
+summarized in Table~\ref{tab:results}.
 
@@
     \centering
-    %TODO: add image
     \begin{subfigure}[b]{0.5\linewidth}
@@
     \end{subfigure}%
-    \caption{t-SNE visualization of the last hidden layer. On the left, the baseline
+    \caption{t-SNE visualization of the last hidden layer using the clean development
+    subset. On the left, the baseline
     model. On the right, the model trained with the adversarial objective. Colors 
@@
 \end{figure}
+
+\begin{table}
+    \centering
+    \caption{Results for word error rate (WER\%) on Aurora-4 dataset.}
+    \label{tab:results}
+    \begin{tabular}{r|cc|cc}
+    \multirow{2}{*}{Subset} & \multicolumn{2}{c}{Orig}  & \multicolumn{2}{|c}{Inv}\\ 
+               & wv1  & wv2  & wv1  & wv2 \\
+    \hline 
+    clean      & 4.04 & 6.76 & & \\
+    airport    & 7.44 & 15.36& & \\
+    babble     & 7.86 & 15.88& & \\
+    car        & 4.17 & 8.99 & & \\
+    restaurant & 9.25 & 17.32& & \\
+    street     & 7.79 & 16.14& & \\
+    train      & 8.18 & 17.26& & \\
+    \hline
+    Average    & \multicolumn{2}{c}{10.46}& & 
+    \end{tabular}
+    %TODO: add numbers!
+\end{table}
 

### 2016-09-13 paper.tex
Add model illustration

@@
 
+\subsection{Adversarial Cost Types}
+
+%TODO: explain -log(\hat{y}) and log(1 - \hat{y})
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.7\linewidth]{model.pdf}
+    \caption{The model consists of three neural networks. The encoder $E$ produces
+    the intermediate representation $h$ which used in the recognizer $R$ and 
+    in the discriminator $D$. The hidden representation $h$ is trained to improve
+    the recognition and minimize the discrimination accuracy. The discriminator
+    is a classifier trained to maximize its accuracy on the gender or the noise type
+    recognition task.}
+    \label{fig:model}
+\end{figure}
+
 \section{Related Work}

### 2016-10-21 paper.tex
Use nips style

@@
 \documentclass{article}
-\usepackage{spconf,amsmath,graphicx}
+
+\usepackage[final]{nips_2016}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage[draft]{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+
+\usepackage{amsmath,graphicx}
 \usepackage{bm}
-\usepackage{hyperref}
 \usepackage{url}
@@
 \usepackage{amsthm}
-\usepackage{amsfonts}
 \usepackage{graphicx}
 \usepackage{subcaption}
-\usepackage{hyperref}
 \usepackage{algorithm2e}
@@
 % ---------------
-\name{Dmitriy Serdyuk $^{\dagger}$\thanks{Dmitriy Serdyuk performed the work
-  while at IBM}, Kartik Audhkhasi$^{\star}$, Phil\'emon Brakel$^{\dagger}$, Bhuvana Ramabhadran$^{\star}$, Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
-\address{$^{\dagger}$ Universit\'e de Montr\'eal\\
-         $^{\star}$ IBM}
+%\name{Dmitriy Serdyuk $^{\dagger}$\thanks{Dmitriy Serdyuk performed the work
+%  while at IBM}, Kartik Audhkhasi$^{\star}$, Phil\'emon Brakel$^{\dagger}$, Bhuvana Ramabhadran$^{\star}$, Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
+%\address{$^{\dagger}$ Universit\'e de Montr\'eal\\
+%         $^{\star}$ IBM}
+\author{
+  Dmitriy Serdyuk\thanks{Dmitriy Serdyuk performed the work 
+    during an internship at IBM} \\
+  MILA\\
+  Universit\'e de Montr\'eal\\
+  Montr\'eal, QC H3T 1J4 \\
+  \texttt{serdyuk@iro.umontreal.ca} \\
+  \And
+  Kartik Audhkhasi \\
+  IBM \\
+  Yorktown Heights, NY 10598\\
+  \texttt{kaudhkha@us.ibm.com} \\
+  \And
+  Phil\'emon Brakel \\
+  MILA\\
+  Universit\'e de Montr\'eal\\
+  Montr\'eal, QC H3T 1J4 \\
+  \texttt{pbpop3@gmail.com} \\
+  \And
+  Bhuvana Ramabhadran\\
+  IBM \\
+  Yorktown Heights, NY 10598\\
+  \texttt{bhuvana@us.ibm.com} \\
+  \And
+  Yoshua Bengio\\
+  MILA, CIFAR Fellow\\
+  Universit\'e de Montr\'eal\\
+  Montr\'eal, QC H3T 1J4 \\
+  \texttt{findme@inter.net} \\
+}
 
@@
     network taking the intermediate representation as the input.
-
     This work focuses on investigation of neural architectures which produce
@@
 %
-\begin{keywords}
-speech recognition, deep neural networks, invariant representations, adversarial training
-\end{keywords}
+%\begin{keywords}
+%speech recognition, deep neural networks, invariant representations, adversarial training
+%\end{keywords}
 %
@@
     the discriminator is just a classifier network with a sigmoid output
-    and can be trained with the gradient methods (see 
-    Section~\ref{sec:optimization}). The main idea of the generative adversarial
+    and can be trained with the gradient methods. The main idea of the generative adversarial
     training is that it is possible to propagate the gradient of the discriminator
-    network further to the generator network staking them on top of each other.
-    This leads to an efficient training procedure and GANs achieve the state of the
+    network further to the generator network stacking them on top of each other.
+    This leads to an efficient training procedure and GANs achieved the state of the
     art that time on handwritten digit generation MNIST dataset.
@@
     to do its best to recognize if there was a deception. This game theoretical
-    similarity gave a name of adversarial networks. The objective of adversarial
-    training is following
+    similarity gave a name of adversarial networks. If we put it formally, the 
+    objective of adversarial training is
     \begin{align*}
-        \min_G \max_D V(D, G) = \mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
+        \min_G \max_D V(D, G) = \mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + 
             \mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
@@
 
-    One of the properties of GAN networks that there is no straightforward way 
-    to evaluate the quality of image generation process since the GAN doesn't 
-    output the log-likelihood of its samples. The paper estimated the likelihood
-    using the Parzen windows estimator.
+    %One of the properties of GAN networks that there is no straightforward way 
+    %to evaluate the quality of image generation process since the GAN doesn't 
+    %output the log-likelihood of its samples. The paper estimated the likelihood
+    %using the Parzen windows estimator.
 
@@
 
+\subsection{Invariant represenations for speech recognition}
+\label{sec:invariant-speech}
+
 \subsection{Adversarial Cost Types}
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{model.pdf}
+    \includegraphics[width=0.3\linewidth]{model.pdf}
     \caption{The model consists of three neural networks. The encoder $E$ produces
@@
 \section{Related Work}
+    Earlier approaches include various works on denoising including
+
+    In a work~\cite{yusuke2016adversarial} an acoustic model trained in an adversarial fashion is
+    applied to an in-house dataset based on Wall Street Journal 1 corrupted with additive noise.
 
@@
 on a noisy speech recognition dataset Aurora-4~\cite{parihar2002aurora}. This dataset
-is based on the Wall Street Journal corpus (WSJ0 is available at LDC). Additive noise of 
+is based on the Wall Street Journal corpus (WSJ0 is available at LDC). Noise of 
 six categories (airport, babble, car, restaurant, street, train) was introduced to the 
-clean data. We used mel-filterbank features with their deltas and delta-deltas spliced
+clean data. This noise was added and filtering simulating sampling frequency was applyed 
+both to clean and noisy data. 
+We used mel-filterbank features with their deltas and delta-deltas spliced
 over 5 frames on each side making 11 consecutive frames and 1320 input features in total. 
+Global mean and variance normalization was applied to these features.
 We did not use any speaker normalization of input features. We pretrained 5 layer
@@
     \hline 
-    clean      & 4.04 & 6.76 & & \\
-    airport    & 7.44 & 15.36& & \\
-    babble     & 7.86 & 15.88& & \\
-    car        & 4.17 & 8.99 & & \\
-    restaurant & 9.25 & 17.32& & \\
-    street     & 7.79 & 16.14& & \\
-    train      & 8.18 & 17.26& & \\
+    clean      & 4.04 & 6.76 & 5.03& 7.53 \\
+    airport    & 7.44 & 15.36& 8.22& 16.03\\
+    babble     & 7.86 & 15.88& 9.06& 17.82\\
+    car        & 4.17 & 8.99 & 5.21& 9.81 \\
+    restaurant & 9.25 & 17.32& 12.2& 19.60\\
+    street     & 7.79 & 16.14& 9.34& 17.56\\
+    train      & 8.18 & 17.26& 9.60& 18.66\\
     \hline
@@
 
-We would like to thank Yaroslav Ganin, David Warde-Farley for insightful discussions,
+We would like to thank Yaroslav Ganin, David Warde-Farley, Samuel Thomas for insightful discussions,
 developers of Theano~\cite{2016arXiv160502688short}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 
@@
 % -------------------------------------------------------------------------
-\bibliographystyle{IEEE}
+%\bibliographystyle{IEEE}
+\bibliographystyle{authordate1}
 \bibliography{refs}

### 2016-10-22 paper.tex
Changes in model description

@@
 % ------
-\title{INVARIANT REPRESENTATIONS FOR NOISY SPEECH RECOGNITION}
+\title{Invariant Representations for Noisy Speech Recognition}
 %
@@
   Dmitriy Serdyuk\thanks{Dmitriy Serdyuk performed the work 
-    during an internship at IBM} \\
+    during an internship at IBM Watson} \\
   MILA\\
@@
   Kartik Audhkhasi \\
-  IBM \\
+  IBM Watson Research Center \\
   Yorktown Heights, NY 10598\\
@@
   Bhuvana Ramabhadran\\
-  IBM \\
+  IBM  Watson Research Center\\
+  Yorktown Heights, NY 10598\\
+  \texttt{bhuvana@us.ibm.com} \\
+  \And
+  Samuel Thomas\\
+  IBM  Watson Research Center\\
   Yorktown Heights, NY 10598\\
@@
 
-    %One of the properties of GAN networks that there is no straightforward way 
-    %to evaluate the quality of image generation process since the GAN doesn't 
-    %output the log-likelihood of its samples. The paper estimated the likelihood
-    %using the Parzen windows estimator.
-
     In practice, GANs might be hard to train due to the fact that in a case when 
@@
     \begin{equation}
-        L = L(\hat{y}, y; \theta_R, \theta_C) + 
+        L = L(\hat{y}, y; \theta_R, \theta_E) + 
         \alpha L(\hat{d}, d; \theta_D) -
-        \beta L(\hat{d}, d; \theta_R),
+        \beta L(\hat{d}, d; \theta_E),
     \end{equation}
@@
     hatted variables are the network predictions, $\theta$ are the network 
-    parameters, and $R, C, D$ are the subsets of parameters in the representation,
-    classifier and domain adaptation networks respectively. The hyper-parameters
+    parameters, and $E, R, D$ are the subsets of parameters in the encoder,
+    recogniser and the domain adaptation networks respectively. The hyper-parameters
     $\alpha$ and $\beta$ denote the influence of every part of the cost.
 
-\subsection{Invariant represenations for speech recognition}
+\subsection{Invariant Represenations for Speech Recognition}
 \label{sec:invariant-speech}
 
-\subsection{Adversarial Cost Types}
+In this work we focus on speech recognition using DNN-HMM hybrid systems. The
+recording conditions, speaker identity, or gener might be considred to be domains
+and adptation task corresponds to making the hidden layer representations be 
+invariant with respect to those conditions. We hypothesize that this method of
+training helps to generalize on unseen conditiions and requires only very small
+amount of labeling the conditions, see Section~\ref{sec:experiments} for experimental
+evaluation.
 
-%TODO: explain -log(\hat{y}) and log(1 - \hat{y})
+The model we use is depicted on Figure~\ref{fig:model}. It is a feedforward neural network
+which is trained to predict the HMM state tweaked with a branch for the invariance
+training. This branch is discarded in the testing phase. In our experiments we
+used the noise condition as the domain label merging all noise type to one label
+and clean type to another.
+
+The particular cost we used for our training is
+    \begin{equation}
+        L = L_1(\hat{y}, y; \theta_R, \theta_E) + 
+        L_2(\hat{d}, d; \theta_D) +
+        \beta L_3(\hat{d}, d; \theta_E),
+        \label{eq:cost}
+    \end{equation}
+where $L_3 = - d\log(1 - \hat{d}) - (1-d)\log(\hat{d})$ maximizes the probablity
+of an incorrect domain classification. The terms for $L_1$ and $L_2$ are 
+regular cross-entropy terms which are minimized with respect to the parameters
+after semicolon.
+We don't use any coefficient for first two terms because it is equivalent to 
+changing the learning rate.
 
@@
 \section{Experiments}
+\label{sec:experiments}
 We experimentally evaluate the approach discussed in Section~\ref{sec:gradient-reverse} 
 on a noisy speech recognition dataset Aurora-4~\cite{parihar2002aurora}. This dataset
-is based on the Wall Street Journal corpus (WSJ0 is available at LDC). Noise of 
-six categories (airport, babble, car, restaurant, street, train) was introduced to the 
-clean data. This noise was added and filtering simulating sampling frequency was applyed 
-both to clean and noisy data. 
-We used mel-filterbank features with their deltas and delta-deltas spliced
-over 5 frames on each side making 11 consecutive frames and 1320 input features in total. 
-Global mean and variance normalization was applied to these features.
-We did not use any speaker normalization of input features. We pretrained 5 layer
-speech recognition network with 2048 rectified linear units at every layer using Adam 
-optimization algorithm for 5 epochs. After pretraining, we turned on the invariance term.
+is based on the Wall Street Journal corpus (WSJ0 is available at LDC). It contains noise of 
+six categories which was added to clean
+data. Every clean and noisy utterance filtered to simulate sampling frequency. The train
+set contains approximately 4400 clean utterances and 446 utterances of each noise condition,
+in total 2676 noisy utterances.
+The test
+set contains clean data, the same 6 noises, and data recorded with a different microphone
+both clean and noise corrupted.
 
-The first expriment we performed is the gender invariance. We added a 2 layer network 
-branching off the last layer of the pretrained speech recognition network and 
-trained it until convergence with SGD. Although, the performance remained unchanged,
-the t-SNE visualization in Figure~\ref{fig:tsne} demonstrates that the hidden units
-are less separated in the projected space.
+We used 16kHz training audio. We extracted mel-filterbank features with their deltas and 
+delta-deltas spliced over 5 frames on each side making 11 consecutive frames and 1320 input 
+features in total. Global mean and variance normalization was applied to these features.
+We did not use any speaker normalization of input features. We trained 6 layer
+speech recognition network with 2048 rectified linear units at every layer using momentum
+accelerated stochastic gradient descent for 15 epochs. We annealed the learning rate accroding 
+to the development set dividing it by 2 when the cross entropy did not improve.
 
-Next, we trained a model for the noise condition invariance. We used an architecture
-similar to one described above with a binary invariance classifier target denoting 
-whether the utterance is clean or noisy. We oversampled the noisy utterances to 
-address the fact that the labels are not ballanced, the clean subset has approximately 4400
-when every noisy condition subset contains only 443 utterances. The results are
-summarized in Table~\ref{tab:results}.
+In order to evaluate the impact of our method on generalization to new, unseen noises 
+we performs 6 groups of experiments with different set of seen noises. We always train
+on clean data and add noise conditions one-by-one in order: airport, babble, car, 
+restaurant, street, train. The last group includes all noises therefore matches the
+standard multi condition training setup. For every group of experiments we trained the
+baseline and the invariance model where we branch out at 4th layer to a logistic 
+binary classifier clean versus noisy data. Due to the imballance between clean and
+noisy parts of the dataset, we had to oversample noisy frames, every minibatch contained
+half of clean and half of noisy data points.
 
-\begin{figure}
-    \centering
-    \begin{subfigure}[b]{0.5\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{tsne-original.png}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.5\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{tsne-invariant.png}
-    \end{subfigure}%
-    \caption{t-SNE visualization of the last hidden layer using the clean development
-    subset. On the left, the baseline
-    model. On the right, the model trained with the adversarial objective. Colors 
-    represent genders.}
-    \label{fig:tsne}
-\end{figure}
+The results are summarized in Table~\ref{tab:results}. 
+In our experiments we did not use layerwise pretraining not uncommonly used for small
+datasets due to the difficulties of combining it with the invariance training. Our preliminary
+experiments show that the network is not able to leave the parameter region it reaches after 
+pretraining.
+
+%\begin{figure}
+%    \centering
+%    \begin{subfigure}[b]{0.5\linewidth}
+%        \centering
+%        \includegraphics[width=\linewidth]{tsne-original.png}
+%    \end{subfigure}%
+%    \begin{subfigure}[b]{0.5\linewidth}
+%        \centering
+%        \includegraphics[width=\linewidth]{tsne-invariant.png}
+%    \end{subfigure}%
+%    \caption{t-SNE visualization of the last hidden layer using the clean development
+%    subset. On the left, the baseline
+%    model. On the right, the model trained with the adversarial objective. Colors 
+%    represent genders.}
+%    \label{fig:tsne}
+%\end{figure}
 
@@
     \centering
-    \caption{Results for word error rate (WER\%) on Aurora-4 dataset.}
+    \caption{Average word error rate (WER\%) on Aurora-4 dataset on all test conditions,
+        including seen and unseen noise and unseen microphone. First column
+        is the number of noise conditions used for the training set. The last row is a 
+        preliminary experiment with layer-wise pretraining close to state-of-the-art
+        model and a corresponding invariance training starting with a pretrained model.}
     \label{tab:results}
-    \begin{tabular}{r|cc|cc}
-    \multirow{2}{*}{Subset} & \multicolumn{2}{c}{Orig}  & \multicolumn{2}{|c}{Inv}\\ 
-               & wv1  & wv2  & wv1  & wv2 \\
+    \begin{tabular}{r|cc}
+    %\multirow{2}{*}{Subset} & \multicolumn{2}{c}{Orig}  & \multicolumn{2}{|c}{Inv}\\ 
+    %           & wv1  & wv2  & wv1  & wv2 \\
+    Noise       &Invariant    &Baseline\\
     \hline
-    clean      & 4.04 & 6.76 & 5.03& 7.53 \\
-    airport    & 7.44 & 15.36& 8.22& 16.03\\
-    babble     & 7.86 & 15.88& 9.06& 17.82\\
-    car        & 4.17 & 8.99 & 5.21& 9.81 \\
-    restaurant & 9.25 & 17.32& 12.2& 19.60\\
-    street     & 7.79 & 16.14& 9.34& 17.56\\
-    train      & 8.18 & 17.26& 9.60& 18.66\\
-    \hline
-    Average    & \multicolumn{2}{c}{10.46}& & 
+    1           &16.36*       &18.14 \\
+    2           &15.56*       &17.39 \\
+    3           &14.29        &14.67 \\
+    4           &13.61        &13.84 \\         
+    5           &13.41        &13.02 \\         
+    6           &12.62        &12.60 \\
+    \hline\hline
+    6, pre-train&11.88        &11.99
     \end{tabular}
@@
 
-\section{Conclusion}
+\section{Discussion}
+    We demonstrate that the invariance training helps to generalize to unseen noise
+    conditions and improves in
+
+    The task of speech recognition differs from the image recognition
+    adaptation in an aspect of the domain adaptation network training dynamics. 
+    While the most common problem for the GAN type of networks trained on images 
+    is overfitting, the noise type classification networks suffers from 
+    underfitting. This makes the gradient of the $L_3$ term in Eq.~\ref{eq:cost}
+    unrealiable and noisy.
 
@@
 
-We would like to thank Yaroslav Ganin, David Warde-Farley, Samuel Thomas for insightful discussions,
+We would like to thank Yaroslav Ganin, David Warde-Farley for insightful discussions,
 developers of Theano~\cite{2016arXiv160502688short}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 

### 2016-10-23 paper.tex
Updated numbers and added plots

@@
     \centering
-    \includegraphics[width=0.3\linewidth]{model.pdf}
+    \captionsetup[subfigure]{oneside,margin={0.3cm,0cm}}
+    \begin{subfigure}[b]{0.4\linewidth}
+        \centering
+        \includegraphics[width=0.9\linewidth]{model.pdf}
         \caption{The model consists of three neural networks. The encoder $E$ produces
@@
         \label{fig:model}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.6\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{wer_avg.pdf}
+        \includegraphics[width=\linewidth]{wer_seen_unseen.pdf}
+        \caption{Up: Average performance of the model depending on the number of noise
+            conditions used for training. Bottom: Average performance on seen condition versus unseen.
+            Testing was performed on all wv1 conditions.
+            }
+        \label{fig:model}
+    \end{subfigure}%
 \end{figure}
@@
         including seen and unseen noise and unseen microphone. First column
-        is the number of noise conditions used for the training set. The last row is a 
+        is the number of noise conditions used for the training. The last row is a 
         preliminary experiment with layer-wise pretraining close to state-of-the-art
@@
     \label{tab:results}
-    \begin{tabular}{r|cc}
+    \begin{tabular}{r|cc||cc|cc|cc|cc}
     %\multirow{2}{*}{Subset} & \multicolumn{2}{c}{Orig}  & \multicolumn{2}{|c}{Inv}\\ 
     %           & wv1  & wv2  & wv1  & wv2 \\
-    Noise       &Invariant    &Baseline\\
+        Noise       &Inv&BL&  \multicolumn{2}{c|}{A} & \multicolumn{2}{c|}{B} & \multicolumn{2}{c|}{C} & \multicolumn{2}{c}{D}\\
+               & & &  Inv & BL & Inv & BL & Inv & BL & Inv & BL\\
     \hline
-    1           &16.36*       &18.14 \\
-    2           &15.56*       &17.39 \\
-    3           &14.29        &14.67 \\
-    4           &13.61        &13.84 \\         
-    5           &13.41        &13.02 \\         
-    6           &12.62        &12.60 \\
+    1           &16.36        &18.14 &6.54&7.57    &12.71& 14.09   & 11.45&   13.10    & 22.47 &   24.80    \\
+    2           &15.56        &17.39 &5.90&  6.58 &   11.69   &13.28   &11.12   &13.51   &21.79   &23.96 \\
+    3           &14.24        &14.67 &5.45 & 5.08&    10.76&   12.44&   9.75&    9.84 &   19.93&   19.30\\
+    4           &13.61        &13.84 & 5.08 &5.29    &9.73    &9.97    &9.49    &9.56    &19.49   &19.90\\         
+    5           &13.41        &13.02 & 5.12 &5.34    &9.52    &9.42    &9.55    &8.67    &19.33   &18.65\\         
+    6           &12.62        &12.60 & 4.80 &4.61    &9.04    &8.86    &8.76    &8.59    &18.16   &18.21\\
     \hline\hline
-    6, pre-train&11.88        &11.99
+    6* &11.88        &11.99
     \end{tabular}

### 2016-10-23 paper.tex
Changed image sizes and edited abstract

@@
     representations invariant to certain conditions as the noise type for speech
-    recognition. Similar
-    to the Ganin et al. work we train the intermediate representation to be invariant
-    to noise conditions training three neural networks: the encoder, the recognizer, and the
-    noise classifier to produce the representation produced by the encoder to decrease
-    the performance of the classifier making the representation invariant to noise conditions. We
+    recognition.  We
     experimentally evaluate proposed architectures on the acoustic modelling task
@@
     \captionsetup[subfigure]{oneside,margin={0.3cm,0cm}}
-    \begin{subfigure}[b]{0.4\linewidth}
+    \begin{subfigure}[b]{0.3\linewidth}
         \centering
@@
     \end{subfigure}%
-    \begin{subfigure}[b]{0.6\linewidth}
+    \begin{subfigure}[b]{0.7\linewidth}
         \centering

### 2016-10-23 paper.tex
Extend introduction, fix figures

@@
     in~\cite{goodfellow2014generative} thus we start with the discussion
-    on this type of generative networks in Section~\ref{sec:gans}
-
-    %TODO
+    on this type of generative networks in Section~\ref{sec:gans}, continue
+    with the description of the gradient reverse method in 
+    Section~\ref{sec:gradient-reverse}. Main contributions of this work
+    include the development of an acoustic model benefiting from invariant
+    representations constructed similarly to the gradient reverse method
+    (see Section~\ref{sec:invariant-speech}); and thorough experimentation
+    on a publicly available Aurora-4 dataset for noisy speech recognition 
+    (see Sectibn~\ref{sec:experiments}).
 
@@
     and can be trained with the gradient methods. The main idea of the generative adversarial
-    training is that it is possible to propagate the gradient of the discriminator
+    training is the possibility to propagate the gradient of the discriminator
     network further to the generator network stacking them on top of each other.
-    This leads to an efficient training procedure and GANs achieved the state of the
-    art that time on handwritten digit generation MNIST dataset.
 
@@
 training helps to generalize on unseen conditiions and requires only very small
-amount of labeling the conditions, see Section~\ref{sec:experiments} for experimental
-evaluation.
+amount of supervision -- the condition labels.
 
@@
             }
-        \label{fig:model}
-    \end{subfigure}%
+        \label{fig:results}
+    \end{subfigure}
+    \caption{Model and results.}
 \end{figure}
@@
 experiments show that the network is not able to leave the parameter region it reaches after 
-pretraining.
-
-%\begin{figure}
-%    \centering
-%    \begin{subfigure}[b]{0.5\linewidth}
-%        \centering
-%        \includegraphics[width=\linewidth]{tsne-original.png}
-%    \end{subfigure}%
-%    \begin{subfigure}[b]{0.5\linewidth}
-%        \centering
-%        \includegraphics[width=\linewidth]{tsne-invariant.png}
-%    \end{subfigure}%
-%    \caption{t-SNE visualization of the last hidden layer using the clean development
-%    subset. On the left, the baseline
-%    model. On the right, the model trained with the adversarial objective. Colors 
-%    represent genders.}
-%    \label{fig:tsne}
-%\end{figure}
+pretraining. The Figure~\ref{fig:results} visualizes the average performance and the
+performance on seen conditions versus unseen. We conclude that the best performance
+gain can be acheived in the case of small amount of noisy samples in the training set.
 

### 2016-10-23 paper.tex
Discuss future work

@@
     \end{subfigure}
-    \caption{Model and results.}
+    \caption{Model structure and results.}
 \end{figure}
@@
 standard multi condition training setup. For every group of experiments we trained the
-baseline and the invariance model where we branch out at 4th layer to a logistic 
+baseline and the invariance model where we branch out at 4th layer to an  
 binary classifier clean versus noisy data. Due to the imballance between clean and
@@
     We demonstrate that the invariance training helps to generalize to unseen noise
-    conditions and improves in
+    conditions and improves in the case of a small amount of training noise. Future
+    research on this topic includes experimentation with layerwise pretraining,
+    more architecture exploration, and the form of the invariance promotion cost term.
 

### 2016-10-23 paper.tex
Edited paper.tex

--- 2016-10-23 paper.tex
@@
   Kartik Audhkhasi \\
-  IBM Watson Research Center \\
+  IBM Watson \\
   Yorktown Heights, NY 10598\\
@@
   Bhuvana Ramabhadran\\
-  IBM  Watson Research Center\\
+  IBM Watson\\
   Yorktown Heights, NY 10598\\
@@
   Samuel Thomas\\
-  IBM  Watson Research Center\\
+  IBM Watson\\
   Yorktown Heights, NY 10598\\
-  \texttt{bhuvana@us.ibm.com} \\
+  \texttt{samthom@us.ibm.com} \\
   \And
@@
 \begin{abstract}
-    Robustness of a speech recognition system is a challenging task and modern
-    automatic speech recognition (ASR) systems need to be resilient to different recording conditions,
-    microphone type, accents, and noisy recording environment. We mitigate
-    this problem in a fashion similar to recent research on image generation using
+    Modern automatic speech recognition (ASR) systems need to be robust under acoustic variability arising from environmental, speaker, channel and recording conditions.
+Ensuring such robustness to variability is a challenge in modern day neural network-based ASR systems, especially when all types of variability are not seen during training.
+We attempt to address this problem by encouraging the neural network acoustic model to learn invariant feature representations.
+    We use ideas from recent research on image generation using
     Generative Adversarial Networks and domain adaptation ideas extending
-    adversarial gradient-based training. A recent work of Ganin et al. proposes to
-    use adversarial training procedure for image domain adaptation making an intermediate
-    representation to improve the target performance and at the same time
-    to deteriorate the domain classifier performance, which is a separate neural
-    network taking the intermediate representation as the input.
-    This work focuses on investigation of neural architectures which produce
-    representations invariant to certain conditions as the noise type for speech
-    recognition.  We
-    experimentally evaluate proposed architectures on the acoustic modelling task
-    for speech recognition on a popular benchmark Aurora-4.
+    adversarial gradient-based training. A recent work from Ganin et al. proposes to
+    use adversarial training for image domain adaptation by using an intermediate
+    representation from the main target classification network to deteriorate the domain 
+    classifier performance through a separate neural network.
+    Our work focuses on investigating neural architectures which produce
+    representations invariant to noise conditions for ASR.  We
+    experimentally evaluate the proposed architecture on a Aurora-4, a popular benchmark for
+    noise robust ASR. We show that our method generalizes better than the standard multi-condition training especially when only a few noise categories are seen during training.
 \end{abstract}
@@
 \label{sec:intro}
-    One of the most challenging aspects of automatic speech recognition (ASR),
-    is the mismatch between the training and testing conditions. During
-    testing, a system may encounter new recording conditions, microphone types,
-    accents and types of background noises. Even if all testing conditions also appeared
-    in the train data, their relative distribution is likely to be
-    different. For this reason, it's important to develop systems that are
-    robust towards changes in recording conditions in general.
+    One of the most challenging aspects of automatic speech recognition (ASR)
+    is the mismatch between the training and testing acoustic conditions. During
+    testing, a system may encounter new recording conditions, microphone types, speakers,
+    accents and types of background noises. Even in case all testing conditions have also appeared
+    in the training data, their acoustic statistics likely to be
+    different. For this reason, it's important to develop ASR systems that are
+    robust towards such undesirable acoustic variabilities.
 
-    While many types of model and feature based adaptation methods like MLLR and fMLLR
-    \cite{leggetter1995maximum} have been proposed, most of these methods have
-    been designed with GMM-HMM systems in mind. Since the enormous success of
-    Deep Neural Network (DNN) systems for ASR \cite{hinton2012deep}, GMM-based
-    model adaptation is not possible anymore, while feature-based methods may
-    not be optimal anymore.
+    While many types of model and feature based adaptation methods like maximum likelihood linear regression (MLLR) and feature-space MLLR (fMLLR)
+    \cite{leggetter1995maximum} have been proposed, most of these methods were
+    designed with Gaussian mixture model-hidden Markov model (GMM-HMM) systems in mind. Since the enormous success of
+    Deep Neural Network (DNN) acoustic models for ASR \cite{hinton2012deep}, GMM-based
+    model adaptation and feature-space adaptation are not optimal for the overall hybrid DNN-HMM system.
     Moreover, the discriminative gradient-based optimization used to train
     DNNs, in combination with the ability of DNNs to learn highly non-linear
-    feature transformations, allows for more flexibility when it comes to the
-    construction of training objectives that promote domain invariant
+    feature transformations, allows for greater flexibility in
+    constructing training objective functions that promote learning of noise invariant
     representations.
 
-    The main idea of this work is that instead of training a robust system,
-    it is possible to train a system which will be invariant to these 
-    conditions. This method of training requires the labels in the dataset
-    which indicate which condition a particular recording was taken in, the 
-    test data doesn't need to contain these labels. The method was proposed in
-    the context of image classification in~\cite{ganin2014unsupervised} and 
-    will be explained in Section~\ref{sec:gradient-reverse} and the application 
-    to speech recognition is discussed
-    in Section~\ref{sec:invariant-speech}. It is tightly connected to the idea
+    The main idea of this work is that instead of explicitly using noise robust acoustic features,
+    we force the acoustic model to learn a representation invariant to noise 
+    conditions. This training requires the noise condition labels labels in the training dataset
+    , though, the 
+    test data does not need to contain these labels. Since this work is tightly connected to the idea
     of generative adversarial networks (GAN) proposed 
-    in~\cite{goodfellow2014generative} thus we start with the discussion
-    on this type of generative networks in Section~\ref{sec:gans}
-
-    %TODO
+    in~\cite{goodfellow2014generative}, Section~\ref{sec:gans} presents an overview of GANs. Section~\ref{sec:gradient-reverse} then presents the gradient reverse method proposed in
+    the context of image classification in~\cite{ganin2014unsupervised}. We then present the application 
+    of this method to speech recognition
+    in Section~\ref{sec:invariant-speech}. Section~\ref{sec:experiments} presents the development of an acoustic model learning invariant
+    representations constructed similarly to the gradient reverse method
+    from Section~\ref{sec:invariant-speech}. This section presents thorough experimentation
+    on a publicly available Aurora-4 dataset for noisy speech recognition.
 
@@
 \label{sec:gans}
-    One of the most insightful recent advantages in the area of generative models
-    is the models of the family of generative adversarial networks (GAN).
+    One of the most insightful recent developments in the area of generative models
+    is the family of generative adversarial networks (GANs).
     A GAN consists of two networks: generator and discriminator. The generator 
     network $G$ has an
-    input of randomly generated vector and is asked to produce an image 
+    input of randomly-generated feature vectors and is asked to produce data, e.g. an image, 
     similar to the images in the training set. The discriminator network $D$
     can either receive a generated image from the generator $G$ or an image
-    from the training set, the task of the discriminator is to distinguish
-    between a generated image and a ``real'' image taken from the dataset. Thus
-    the discriminator is just a classifier network with a sigmoid output
-    and can be trained with the gradient methods. The main idea of the generative adversarial
-    training is that it is possible to propagate the gradient of the discriminator
-    network further to the generator network stacking them on top of each other.
-    This leads to an efficient training procedure and GANs achieved the state of the
-    art that time on handwritten digit generation MNIST dataset.
+    from the training set. Its task is to distinguish
+    between the ``fake'' generated image and the ``real'' image taken from the dataset. Thus,
+    the discriminator is just a classifier network with a sigmoid output neuron
+    and can be trained with gradient backpropagation.
 
-    In other words, these two networks are competing with each other the 
-    generator is trying to deceive the discriminator network while it tries
-    to do its best to recognize if there was a deception. This game theoretical
-    similarity gave a name of adversarial networks. If we put it formally, the 
-    objective of adversarial training is
+    The main idea of generative adversarial
+    training is to stack the discriminator and generator network on top of each other and propagate the gradient of the discriminator
+    network further to the generator network. In other words, these two networks are competing with each other. The 
+    generator is trying to deceive the discriminator network, while the discriminator tries
+    to do its best to recognize if there was a deception, similar to adversarial game-theoretic settings.    
+    Formally, the objective function of GAN training is
     \begin{align*}
@@
     \end{align*}
+	{\bf Please explain how the min-max applies to the two terms.
+I assume that we are maximizing first term with respect to D and doing min over D/max over G for the second term.
+Hence maximizing the first term increases the likelihood of the discriminator output (single sigmoid neuron) for true training data.
+Maximizing the second term decreases the likelihood of the discriminator output for fake training data.
+But then, what does the min over G mean?
+That you are maximizing the generator likelihood of the fake data?}
 
-    In practice, GANs might be hard to train due to the fact that in a case when 
-    the discriminator is more powerful that the generator, the later doesn't 
-    receive any gradient thus can not be improved. Usually this problem is solved
-    using a smart schedule which stops learning the discriminator as soon as it
-    is too far ahead of the generator. Other possible solutions include ``soft''
-    targets bounded at some constant rather than between 0 and 1 and using a 
-    different proxy for the cost.
+    In practice, GANs might be hard to train due to the fact that when 
+    the discriminator is more powerful that the generator, the latter doesn't 
+    receive any gradient and thus can not be improved. One way to solve this problem is to
+    use a smart schedule which stops trainign the discriminator as soon as it
+    is much better trained than the generator. Other possible solutions include ``soft''
+    targets for the discriminator and using a 
+    different proxy for the cost {\bf I don't understand different proxy for the cost}.
 
@@
 
-    A work~\cite{ganin2014unsupervised} proposes a method of training a network 
+    {\bf We need to write a couple of sentences on how this is motivated by GAN.} Prior work by~\cite{ganin2014unsupervised} proposed a method of training a network 
     which can be adapted to new domains. The training set consists of the images
-    labeled with classes and the separate domain labels. The test set doesn't need
-    to have the domain labels. For example, it may be first domain
-    consisting of MNIST digits and second domain of MNIST digits on a different
-    background. The network has a fork-like structure: the image is fed to the
-    first network which produces a hidden representation $h$, then there are two
-    networks branching of the representation $h$: a domain classifier network and 
-    a target classifier. The goal of the training is to produce the hidden 
-    representation invariant to the domain labels so that information doesn't 
-    interfere the target classifier which is used in the test time.
+    labeled with classes of interest and separate domain labels. For example, the domain may refer to image background. The testing set does not need
+    to have the domain labels. The network has a fork-like structure: the image is fed to the
+    first network which produces a hidden representation $h$. Then this representation $h$ is input to two separate networks: a domain classifier network (D) and 
+    a target classifier network (R). The goal of training is to learn the hidden 
+    representation that is invariant to the domain labels and performs well on the target classification task, so that the domain information doesn't 
+    interfere with the target classifier at test time.
 
     The network is trained with three goals: the hidden representation $h$ should
-    be helpful for the target classifier, but harmful for the domain classifier,
-    the domain classifier should have a good classification accuracy. More 
-    formally, we define the objective as
+    be helpful for the target classifier, harmful for the domain classifier,
+    and the domain classifier should have a good classification accuracy. More 
+    formally, the authors define the loss function as
     \begin{equation}
@@
     where $y$ is the ground truth class, $d$ is the domain label, corresponding
-    hatted variables are the network predictions, $\theta$ are the network 
-    parameters, and $E, R, D$ are the subsets of parameters in the encoder,
-    recogniser and the domain adaptation networks respectively. The hyper-parameters
-    $\alpha$ and $\beta$ denote the influence of every part of the cost.
+    hat variables are the network predictions, and $\theta_E, \theta_R$ and $\theta_D$ are the subsets of  parameters for the encoder,
+    recogniser and the domain classifier networks respectively. The hyper-parameters
+    $\alpha$ and $\beta$ denote the relative influence of the loss functions terms.
 
-\subsection{Invariant Represenations for Speech Recognition}
+\subsection{Invariant Representions for Speech Recognition}
 \label{sec:invariant-speech}
 
-In this work we focus on speech recognition using DNN-HMM hybrid systems. The
-recording conditions, speaker identity, or gener might be considred to be domains
-and adptation task corresponds to making the hidden layer representations be 
-invariant with respect to those conditions. We hypothesize that this method of
-training helps to generalize on unseen conditiions and requires only very small
-amount of labeling the conditions, see Section~\ref{sec:experiments} for experimental
-evaluation.
+This work focuses on speech recognition using DNN-HMM hybrid systems. The context dependent (CD) HMM states are the class labels of interest. The
+recording conditions, speaker identity, or gender may be considered to be domains. Then the task is to make the hidden layer representations of the CD state classifier network 
+invariant with respect to these domains. We hypothesize that this adversarial method of
+training helps the CD state classifier to generalize better to unseen domain conditions and requires only a  
+small additional amount of supervision, i.e. the domain labels.
 
-The model we use is depicted on Figure~\ref{fig:model}. It is a feedforward neural network
-which is trained to predict the HMM state tweaked with a branch for the invariance
-training. This branch is discarded in the testing phase. In our experiments we
-used the noise condition as the domain label merging all noise type to one label
-and clean type to another.
-
-The particular cost we used for our training is
+Figure~\ref{fig:model} depicts the model, which is same as the model for the gradient reverse method presented in the previous section.
+It is a feedforward neural network trained to predict the CD HMM state, with a branch that predicts the domain (noise condition).
+This branch is discarded in the testing phase.
+In our experiments we
+used the noise condition as the domain label merging all noise types into one label
+and clean as the other label. Our training loss function is
     \begin{equation}
@@
     \end{equation}
-where $L_3 = - d\log(1 - \hat{d}) - (1-d)\log(\hat{d})$ maximizes the probablity
-of an incorrect domain classification. The terms for $L_1$ and $L_2$ are 
-regular cross-entropy terms which are minimized with respect to the parameters
-after semicolon.
-We don't use any coefficient for first two terms because it is equivalent to 
-changing the learning rate.
+where $L_3 = - d\log(1 - \hat{d}) - (1-d)\log(\hat{d})$ maximizes the probability
+of an incorrect domain classification {\bf Contrast this with term 3 of the original gradient reverse equation.}. The terms $L_1$ and $L_2$ are 
+regular cross-entropies which are minimized with corresponding parameters $\theta_E$ and $\theta_D$.
+For simplicity, we use only a single hyper-parameter - the weight of the third term.
 
@@
         the intermediate representation $h$ which used in the recognizer $R$ and 
-        in the discriminator $D$. The hidden representation $h$ is trained to improve
-        the recognition and minimize the discrimination accuracy. The discriminator
-        is a classifier trained to maximize its accuracy on the gender or the noise type
-        recognition task.}
+        in the domain discriminator $D$. The hidden representation $h$ is trained to improve
+        the recognition and minimize the domain discriminator accuracy. The domain discriminator
+        is a classifier trained to maximize its accuracy on the noise type
+        classification task.}
         \label{fig:model}
@@
         \includegraphics[width=\linewidth]{wer_seen_unseen.pdf}
-        \caption{Up: Average performance of the model depending on the number of noise
-            conditions used for training. Bottom: Average performance on seen condition versus unseen.
+        \caption{Up: Average performance of the baseline multi-condition and invariance model varying with  the number of noise
+            conditions used for training. Bottom: Average performance on seen versus unseen noise conditions.
             Testing was performed on all wv1 conditions.
             }
-        \label{fig:model}
-    \end{subfigure}%
+        \label{fig:results}
+    \end{subfigure}
+    \caption{Model structure for invariant training and ASR results.}
 \end{figure}
@@
 
+    {\bf Need to call out differences from this work: (1) We use the standard Aurora-4 training/testing conditions.
+(2) We also show the impact of seeing different number of noise types during training.
+(3) ...}
+
 \section{Experiments}
@@
 We experimentally evaluate the approach discussed in Section~\ref{sec:gradient-reverse} 
-on a noisy speech recognition dataset Aurora-4~\cite{parihar2002aurora}. This dataset
-is based on the Wall Street Journal corpus (WSJ0 is available at LDC). It contains noise of 
+on the Aurora-4~\cite{parihar2002aurora} noisy speech recognition dataset. Aurora-4
+is based on the Wall Street Journal corpus (WSJ0). It contains noises of 
 six categories which was added to clean
-data. Every clean and noisy utterance filtered to simulate sampling frequency. The train
+data. Every clean and noisy utterance is filtered to simulate sampling frequency {\bf simulate sampling frequency?}. The training
 set contains approximately 4400 clean utterances and 446 utterances of each noise condition,
-in total 2676 noisy utterances.
-The test
-set contains clean data, the same 6 noises, and data recorded with a different microphone
-both clean and noise corrupted.
+i.e. a total of 2676 noisy utterances.
+The testing
+set contains clean data, data corrupted by the 6 noises, and data recorded with a different microphone
+for both clean and noisy cases.
 
-We used 16kHz training audio. We extracted mel-filterbank features with their deltas and 
-delta-deltas spliced over 5 frames on each side making 11 consecutive frames and 1320 input 
-features in total. Global mean and variance normalization was applied to these features.
-We did not use any speaker normalization of input features. We trained 6 layer
-speech recognition network with 2048 rectified linear units at every layer using momentum
-accelerated stochastic gradient descent for 15 epochs. We annealed the learning rate accroding 
-to the development set dividing it by 2 when the cross entropy did not improve.
+We extracted 40-dimensional Mel-filterbank features with their deltas and 
+delta-deltas spliced over $\pm$5 frames, resulting in 1320 input 
+features. We performed speaker-independent global mean and variance normalization on these features. We trained 6 layer
+DNN with 2048 rectified linear units at every layer using momentum-accelerated stochastic gradient descent for 15 epochs. We annealed the learning rate according 
+to the development set cross-entropy, dividing it by 2 when the cross entropy did not improve.
 
-In order to evaluate the impact of our method on generalization to new, unseen noises 
-we performs 6 groups of experiments with different set of seen noises. We always train
-on clean data and add noise conditions one-by-one in order: airport, babble, car, 
-restaurant, street, train. The last group includes all noises therefore matches the
-standard multi condition training setup. For every group of experiments we trained the
-baseline and the invariance model where we branch out at 4th layer to a logistic 
-binary classifier clean versus noisy data. Due to the imballance between clean and
-noisy parts of the dataset, we had to oversample noisy frames, every minibatch contained
-half of clean and half of noisy data points.
+In order to evaluate the impact of our method on generalization to unseen noises,
+we performed 6 experiments with different set of seen noises. We always train
+on clean data and add noise conditions one-by-one in the following order: airport, babble, car, 
+restaurant, street, train. The last training group includes all noises therefore matches the
+standard multi-condition training setup. For every group of experiments we trained the
+baseline and the invariance model where we branch out at the $4^{th}$ layer to an  
+binary classifier predicting clean versus noisy data. Due to the imbalance between amounts of clean and
+noisy utterances, we had to oversample noisy frames. Hence every mini-batch contained
+equal number of clean and noisy speech frames.
 
-The results are summarized in Table~\ref{tab:results}. 
-In our experiments we did not use layerwise pretraining not uncommonly used for small
+Table~\ref{tab:results} summarizes the results.
+Figure~\ref{fig:results} visualizes the word error rate for the baseline multi-condition training and invariance training as the number of seen noise types varies.
+We conclude that the best performance
+gain is achieved when a small number of noise types are available during training.
+We conclude that invariance training is able to generalize better to unseen noise types compared with multi-condition training.
+
+We note that our experiments did not use layer-wise pre-training not uncommonly used for small
 datasets due to the difficulties of combining it with the invariance training. Our preliminary
-experiments show that the network is not able to leave the parameter region it reaches after 
-pretraining.
-
-%\begin{figure}
-%    \centering
-%    \begin{subfigure}[b]{0.5\linewidth}
-%        \centering
-%        \includegraphics[width=\linewidth]{tsne-original.png}
-%    \end{subfigure}%
-%    \begin{subfigure}[b]{0.5\linewidth}
-%        \centering
-%        \includegraphics[width=\linewidth]{tsne-invariant.png}
-%    \end{subfigure}%
-%    \caption{t-SNE visualization of the last hidden layer using the clean development
-%    subset. On the left, the baseline
-%    model. On the right, the model trained with the adversarial objective. Colors 
-%    represent genders.}
-%    \label{fig:tsne}
-%\end{figure}
+experiments on a pre-trained network when using all noise types (last row of Table~\ref{tab:results}) for training show modest gains in WER over the multi-condition baseline.
 
@@
 \section{Discussion}
-    We demonstrate that the invariance training helps to generalize to unseen noise
-    conditions and improves in
+    This paper presents the application of generative adversarial networks and invariance training for noise robust speech recognition.
+We show that invariance training helps the ASR system to generalize to better to unseen noise
+    conditions and improves word error rate in when a small number of noise types are seen during trainings. Future
+    research on this topic includes experimentation with layer-wise pre-training,
+    exploring alternative network architectures and invariance-promoting loss functions
 
-    The task of speech recognition differs from the image recognition
+    {\bf We need to make the message of this paragraph more clear.}The task of speech recognition differs from the image recognition
     adaptation in an aspect of the domain adaptation network training dynamics. 
@@
 developers of Theano~\cite{2016arXiv160502688short}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 
-for great instruments.
+for great toolkits.
 

### 2016-10-23 paper.tex
Address K's comments

@@
     \end{align*}
-	{\bf Please explain how the min-max applies to the two terms.
+    The maximization over the discriminator $D$ forms a usuall cross-entropy objective, the gradients are
+    computed with respect to the parameters of $D$. The parameters of $G$ are minimized using the gradients
+    propagated through the second term. The minimization over $G$ makes it to produce examples which $D$
+    classifiers as the training ones.
+	%{\bf Please explain how the min-max applies to the two terms.
 I assume that we are maximizing first term with respect to D and doing min over D/max over G for the second term.
@@
 That you are maximizing the generator likelihood of the fake data?}
-
-    In practice, GANs might be hard to train due to the fact that when 
-    the discriminator is more powerful that the generator, the latter doesn't 
-    receive any gradient and thus can not be improved. One way to solve this problem is to
-    use a smart schedule which stops trainign the discriminator as soon as it
-    is much better trained than the generator. Other possible solutions include ``soft''
-    targets for the discriminator and using a 
-    different proxy for the cost {\bf I don't understand different proxy for the cost}.
+    % Dima: deleted paragraph and explained min-max
 
@@
 
-    {\bf We need to write a couple of sentences on how this is motivated by GAN.} Prior work by~\cite{ganin2014unsupervised} proposed a method of training a network 
+    %{\bf We need to write a couple of sentences on how this is motivated by GAN.} 
+    %Dima: added a sentence at the end of this paragraph
+    Prior work by~\cite{ganin2014unsupervised} proposed a method of training a network 
     which can be adapted to new domains. The training set consists of the images
@@
     representation that is invariant to the domain labels and performs well on the target classification task, so that the domain information doesn't 
-    interfere with the target classifier at test time.
+    interfere with the target classifier at test time. Like the GAN objective forces the generation distribution be close to the data distribution,
+    the gradient reverse method makes domain distributions similar to each other.
 
@@
 \section{Related Work}
-    Earlier approaches include various works on denoising including
+    Generative adversarial networks were proposed and experimentally evaluated 
+    in~\cite{goodfellow2014generative}. This work was continued 
+    in work of~\cite{radford2015unsupervised}, which proposes several practical guidelines
+    for training GANs. This direction is further explored in~\cite{salimans2016improved}.
+
+    The conditioned GAN for generation task were utilized in~\cite{denton2015deep}. This
+    work proposes to improve images increasing the resolution step-by-step. 
+    The domain adaptation with the gradient reverse method was proposed in~\cite{ganin2014unsupervised}.
+
+    %TODO: cite denoising papers? some work on aurora 4?
 
@@
 six categories which was added to clean
-data. Every clean and noisy utterance is filtered to simulate sampling frequency {\bf simulate sampling frequency?}. The training
+data. Every clean and noisy utterance is filtered to simulate the frequency characteristics of the sampling frequency. 
+% Dima: is it correct now?
+%
+The training
 set contains approximately 4400 clean utterances and 446 utterances of each noise condition,
@@
 
-    {\bf We need to make the message of this paragraph more clear.}The task of speech recognition differs from the image recognition
-    adaptation in an aspect of the domain adaptation network training dynamics. 
-    While the most common problem for the GAN type of networks trained on images 
-    is overfitting, the noise type classification networks suffers from 
-    underfitting. This makes the gradient of the $L_3$ term in Eq.~\ref{eq:cost}
-    unrealiable and noisy.
+    %{\bf We need to make the message of this paragraph more clear.}
+    % Dima: tried to fix -- it is differnt => needs more investigation
+    Our experiments show that in contrast to the image recognition task,
+    the task of speech recognition,  the domain adaptation network suffers from
+    the underfitting. Therefore, the gradient of the $L_3$ term in Eq.~\ref{eq:cost}
+    unrealiable and noisy. This aspect needs further development.
 

### 2016-10-23 paper.tex
One more comment

@@
         \beta L(\hat{d}, d; \theta_E),
+        \label{eq:grm}
     \end{equation}
@@
 where $L_3 = - d\log(1 - \hat{d}) - (1-d)\log(\hat{d})$ maximizes the probability
-of an incorrect domain classification {\bf Contrast this with term 3 of the original gradient reverse equation.}. The terms $L_1$ and $L_2$ are 
+of an incorrect domain classification in contrast to the Eq.~\ref{eq:grm} where the 
+correct classification is minimized.
+%{\bf Contrast this with term 3 of the original gradient reverse equation.}. 
+The terms $L_1$ and $L_2$ are 
 regular cross-entropies which are minimized with corresponding parameters $\theta_E$ and $\theta_D$.

### 2016-10-23 paper.tex
Add citation of aurora4 paper
Fix typos

@@
     \end{align*}
-    The maximization over the discriminator $D$ forms a usuall cross-entropy objective, the gradients are
+    The maximization over the discriminator $D$ forms a usual cross-entropy objective, the gradients are
     computed with respect to the parameters of $D$. The parameters of $G$ are minimized using the gradients
@@
     hat variables are the network predictions, and $\theta_E, \theta_R$ and $\theta_D$ are the subsets of  parameters for the encoder,
-    recogniser and the domain classifier networks respectively. The hyper-parameters
+    recognizer and the domain classifier networks respectively. The hyper-parameters
     $\alpha$ and $\beta$ denote the relative influence of the loss functions terms.
 
-\subsection{Invariant Representions for Speech Recognition}
+\subsection{Invariant Representations for Speech Recognition}
 \label{sec:invariant-speech}
@@
 Figure~\ref{fig:model} depicts the model, which is same as the model for the gradient reverse method presented in the previous section.
-It is a feedforward neural network trained to predict the CD HMM state, with a branch that predicts the domain (noise condition).
+It is a feed-forward neural network trained to predict the CD HMM state, with a branch that predicts the domain (noise condition).
 This branch is discarded in the testing phase.
@@
     %TODO: cite denoising papers? some work on aurora 4?
+
+    The influence of representations produced by a neural network to internal noise reduction 
+    is discussed in~\cite{yu2013feature} and this work sets a baseline for experiments on
+    Aurora-4 dataset.
 
@@
 We show that invariance training helps the ASR system to generalize to better to unseen noise
-    conditions and improves word error rate in when a small number of noise types are seen during trainings. Future
+    conditions and improves word error rate in when a small number of noise types are seen during training. Future
     research on this topic includes experimentation with layer-wise pre-training,
@@
     the underfitting. Therefore, the gradient of the $L_3$ term in Eq.~\ref{eq:cost}
-    unrealiable and noisy. This aspect needs further development.
+    unreliable and noisy. This aspect needs further development.
 

### 2016-10-23 paper.tex
Fix one more typo
Rephrase related work

@@
   Dmitriy Serdyuk\thanks{Dmitriy Serdyuk performed the work 
-    during an internship at IBM Watson} \\
+    during an internship at IBM Watson.} \\
   MILA\\
@@
 regular cross-entropies which are minimized with corresponding parameters $\theta_E$ and $\theta_D$.
-For simplicity, we use only a single hyper-parameter - the weight of the third term.
+For simplicity, we use only a single hyper-parameter -- the weight of the third term.
 
@@
 
-    In a work~\cite{yusuke2016adversarial} an acoustic model trained in an adversarial fashion is
-    applied to an in-house dataset based on Wall Street Journal 1 corrupted with additive noise.
+    In a work~\cite{yusuke2016adversarial} an acoustic model: multilayer sigmoidal network is 
+    trained in an adversarial fashion for a task of dataset corrupted with additive noise.
 
-    {\bf Need to call out differences from this work: (1) We use the standard Aurora-4 training/testing conditions.
+    %{\bf Need to call out differences from this work: (1) We use the standard Aurora-4 training/testing conditions.
 (2) We also show the impact of seeing different number of noise types during training.

### 2016-10-23 paper.tex
Add files via upload
Rearranged sections and smoothed some wording; 4 pages.

@@
     during an internship at IBM Watson.} \\
-  MILA\\
+  MILA,
   Universit\'e de Montr\'eal\\
@@
   Kartik Audhkhasi \\
-  IBM Watson \\
-  Yorktown Heights, NY 10598\\
+  IBM Watson, \\
+  Yorktown Heights, NY \\
   \texttt{kaudhkha@us.ibm.com} \\
@@
   Universit\'e de Montr\'eal\\
-  Montr\'eal, QC H3T 1J4 \\
+  %Montr\'eal, QC H3T 1J4 \\
   \texttt{pbpop3@gmail.com} \\
@@
   Bhuvana Ramabhadran\\
-  IBM Watson\\
-  Yorktown Heights, NY 10598\\
+  IBM Watson,
+  Yorktown Heights, NY \\
   \texttt{bhuvana@us.ibm.com} \\
@@
   Samuel Thomas\\
-  IBM Watson\\
-  Yorktown Heights, NY 10598\\
+  IBM Watson,
+  Yorktown Heights, NY \\
   \texttt{samthom@us.ibm.com} \\
@@
   Yoshua Bengio\\
-  MILA, CIFAR Fellow\\
-  Universit\'e de Montr\'eal\\
-  Montr\'eal, QC H3T 1J4 \\
+  MILA, CIFAR Fellow,
+  Universit\'e de Montr\'eal,
+  %Montr\'eal, QC H3T 1J4 \\
   \texttt{findme@inter.net} \\
@@
     Our work focuses on investigating neural architectures which produce
-    representations invariant to noise conditions for ASR.  We
-    experimentally evaluate the proposed architecture on a Aurora-4, a popular benchmark for
+    representations invariant to noise conditions for ASR.  We evaluate the proposed architecture on the Aurora-4 task, a popular benchmark for
     noise robust ASR. We show that our method generalizes better than the standard multi-condition training especially when only a few noise categories are seen during training.
@@
     testing, a system may encounter new recording conditions, microphone types, speakers,
-    accents and types of background noises. Even in case all testing conditions have also appeared
-    in the training data, their acoustic statistics likely to be
-    different. For this reason, it's important to develop ASR systems that are
-    robust towards such undesirable acoustic variabilities.
+    accents and types of background noises.
+Furthermore, even if the test scenarios are seen during training, there can be significant variability in their statistics.
+Thus, its important to develop ASR systems that are invariant to unseen acoustic conditions.
+    %bhuv
+    %Even in case all testing conditions have also appeared
+    i%n the training data, their acoustic statistics likely to be
+    %different. For this reason, it's important to develop ASR systems that are
+    %robust towards such undesirable acoustic variabilities.
 
-    While many types of model and feature based adaptation methods like maximum likelihood linear regression (MLLR) and feature-space MLLR (fMLLR)
-    \cite{leggetter1995maximum} have been proposed, most of these methods were
-    designed with Gaussian mixture model-hidden Markov model (GMM-HMM) systems in mind. Since the enormous success of
-    Deep Neural Network (DNN) acoustic models for ASR \cite{hinton2012deep}, GMM-based
-    model adaptation and feature-space adaptation are not optimal for the overall hybrid DNN-HMM system.
-    Moreover, the discriminative gradient-based optimization used to train
-    DNNs, in combination with the ability of DNNs to learn highly non-linear
-    feature transformations, allows for greater flexibility in
-    constructing training objective functions that promote learning of noise invariant
-    representations.
+    %While many types of model and feature based adaptation methods such as, maximum likelihood linear regression (MLLR) and feature-space MLLR (fMLLR)
+    %\cite{leggetter1995maximum} have been proposed, most of these methods were
+    %designed with Gaussian mixture model-hidden Markov model (GMM-HMM) systems in mind. Since the enormous success of
+        %model adaptation and feature-space adaptation are not optimal for the overall hybrid DNN-HMM system.
+    %methods Moreover, the discriminative gradient-based optimization used to train
+    %DNNs, in combination with the ability of DNNs to learn highly non-linear
+%feature transformations, allows for greater flexibility in
+    %constructing training objective functions that promote learning of noise invariant
+    %representations.
+    Several model and feature based adaptation methods such as Maximum Likelihood Linear Regression (MLLR),  feature-based MLLR and iVectors~\cite{ivec} have been proposed to handle speaker variability; and Noise Adaptive Training (NAT)~\cite{msftnoise}  and Vector Taylor Series (VTS) to handle environment variability.
+With the increasing success of Deep Neural Network (DNN) acoustic models for ASR \cite{hinton2012deep, msft, ibm}, end-to-end systems are being proposed~\cite{essen, scatterTransforms, fromaudio} for modeling the acoustic conditions within a single network.
+This allows us to take advantage of the network's ability to learn highly non-linear feature transformations, with greater flexibility in constructing training objective functions that promote learning of noise invariant representations.
+The main idea of this work is to force the acoustic model to learn a representation invariant to noise conditions, instead of explicitly using noise robust acoustic features (Section~\ref{sec:invariant-speech}).
+This type of noise-invariant training requires noise-condition labels during training only.
+It is related to the idea of generative adversarial networks (GAN) and the gradient reverse method proposed in proposed  in~\cite{goodfellow2014generative}  and ~\cite{ganin2014unsupervised} respectively (Section~\ref{sec:relatedwork}).
+We present results on the Aurora-4 speech recognition task ( Section~\ref{sec:experiments} and summarize our findings in Section~\ref{sec:discussion}.
 
-    The main idea of this work is that instead of explicitly using noise robust acoustic features,
-    we force the acoustic model to learn a representation invariant to noise 
-    conditions. This training requires the noise condition labels labels in the training dataset
-    , though, the 
-    test data does not need to contain these labels. Since this work is tightly connected to the idea
-    of generative adversarial networks (GAN) proposed 
-    in~\cite{goodfellow2014generative}, Section~\ref{sec:gans} presents an overview of GANs. Section~\ref{sec:gradient-reverse} then presents the gradient reverse method proposed in
-    the context of image classification in~\cite{ganin2014unsupervised}. We then present the application 
-    of this method to speech recognition
-    in Section~\ref{sec:invariant-speech}. Section~\ref{sec:experiments} presents the development of an acoustic model learning invariant
-    representations constructed similarly to the gradient reverse method
-    from Section~\ref{sec:invariant-speech}. This section presents thorough experimentation
-    on a publicly available Aurora-4 dataset for noisy speech recognition.
+    %Since this work is tightly connected to the idea
+    %f generative adversarial networks (GAN) proposed 
+    %in~\cite{goodfellow2014generative}, Section~\ref{sec:gans} presents an overview of GANs. Section~\ref{sec:gradient-reverse} then presents the gradient reverse method proposed in
+    %the context of image classification in~\cite{ganin2014unsupervised}. We then present the application 
+    %of this method to speech recognition
+    %in Section~\ref{sec:invariant-speech}. Section~\ref{sec:experiments} presents the development of an acoustic model learning invariant
+    %representations constructed similarly to the gradient reverse method
+    %from Section~\ref{sec:invariant-speech}. This section presents thorough experimentation
+    %on a publicly available Aurora-4 dataset for noisy speech recognition.
 
-\section{Generative Adversarial Networks}
-\label{sec:gans}
-    One of the most insightful recent developments in the area of generative models
-    is the family of generative adversarial networks (GANs).
-    A GAN consists of two networks: generator and discriminator. The generator 
-    network $G$ has an
+\section{Related Work}
+\label{sec:relatedwork}
+{\bf Generative Adversarial Networks} consist of two networks: generator and discriminator. 
+    %One of the most insightful recent developments in the area of generative models
+    %is the family of generative adversarial networks (GANs).
+    %A GAN consists of two networks: generator and discriminator. 
+    The generator network $G$ has an
     input of randomly-generated feature vectors and is asked to produce data, e.g. an image, 
@@
     between the ``fake'' generated image and the ``real'' image taken from the dataset. Thus,
-    the discriminator is just a classifier network with a sigmoid output neuron
+    the discriminator is just a classifier network with a sigmoid output layer
     and can be trained with gradient backpropagation.
@@
 
-\section{Gradient Reverse Method}
-\label{sec:gradient-reverse}
+%\section{Gradient Reverse Method}
+%\label{sec:gradient-reverse}
 
@@
     %Dima: added a sentence at the end of this paragraph
+
+    % bhuv :We should say how this is different form multi-task learning
     Prior work by~\cite{ganin2014unsupervised} proposed a method of training a network 
-    which can be adapted to new domains. The training set consists of the images
-    labeled with classes of interest and separate domain labels. For example, the domain may refer to image background. The testing set does not need
-    to have the domain labels. The network has a fork-like structure: the image is fed to the
+    which can be adapted to new domains. The training data consists of the images
+    labeled with classes of interest and separate domain (image background) labels. 
+    %For example, the domain may refer to image background. The testing set does not need
+   % to have the domain labels. 
+   The network has a fork-like structure: the image is fed to the
     first network which produces a hidden representation $h$. Then this representation $h$ is input to two separate networks: a domain classifier network (D) and 
@@
     representation that is invariant to the domain labels and performs well on the target classification task, so that the domain information doesn't 
-    interfere with the target classifier at test time. Like the GAN objective forces the generation distribution be close to the data distribution,
-    the gradient reverse method makes domain distributions similar to each other.
+    interfere with the target classifier at test time. Similar to the GAN objective, which forces the generation distribution be close to the data distribution,
+    the {\bf gadient reverse method} makes domain distributions similar to each other.
 
@@
 
-\subsection{Invariant Representations for Speech Recognition}
+%% Denoising Auto encoder and
+
+The influence of representations produced by a neural network to internal noise reduction is discussed in~\cite{yu2013feature} and this work sets a baseline for experiments on  Aurora-4 dataset.
+Recently, in~\cite{yusuke2016adversarial} a multilayer sigmoidal network is trained in an adversarial fashion on a in-house transcription task corrupted by noise.
+    %% STILL TODO
+    %{\bf Need to call out differences from this work: (1) We use the standard Aurora-4 training/testing conditions.
+(2) We also show the impact of seeing different number of noise types during training.
+(3) ...}
+\section{Invariant Representations for Speech Recognition}
 \label{sec:invariant-speech}
 
-This work focuses on speech recognition using DNN-HMM hybrid systems. The context dependent (CD) HMM states are the class labels of interest. The
-recording conditions, speaker identity, or gender may be considered to be domains. Then the task is to make the hidden layer representations of the CD state classifier network 
+Most ASR systems are DNN-HMM hybrid systems. The context dependent (CD) HMM states (acoustic model) are the class labels of interest. The
+recording conditions, speaker identity, or gender represent the domains in GANs. The task is to make the hidden layer representations of the HMM state classifier network 
 invariant with respect to these domains. We hypothesize that this adversarial method of
-training helps the CD state classifier to generalize better to unseen domain conditions and requires only a  
+training helps the HMM state classifier to generalize better to unseen domain conditions and requires only a  
 small additional amount of supervision, i.e. the domain labels.
 
-Figure~\ref{fig:model} depicts the model, which is same as the model for the gradient reverse method presented in the previous section.
+Figure~\ref{fig:model} depicts the model, which is same as the model for the gradient reverse method.
 It is a feed-forward neural network trained to predict the CD HMM state, with a branch that predicts the domain (noise condition).
@@
 
-\section{Related Work}
-    Generative adversarial networks were proposed and experimentally evaluated 
-    in~\cite{goodfellow2014generative}. This work was continued 
-    in work of~\cite{radford2015unsupervised}, which proposes several practical guidelines
-    for training GANs. This direction is further explored in~\cite{salimans2016improved}.
+% Merged into earlier section
+%\section{Related Work}
+ %   Generative adversarial networks were proposed and experimentally evaluated 
+  %  in~\cite{goodfellow2014generative}. This work was continued 
+   % in work of~\cite{radford2015unsupervised}, which proposes several practical guidelines
+    %for training GANs. This direction is further explored in~\cite{salimans2016improved}.
 
-    The conditioned GAN for generation task were utilized in~\cite{denton2015deep}. This
-    work proposes to improve images increasing the resolution step-by-step. 
-    The domain adaptation with the gradient reverse method was proposed in~\cite{ganin2014unsupervised}.
+ %   The conditioned GAN for generation task were utilized in~\cite{denton2015deep}. This
+  %  work proposes to improve images increasing the resolution step-by-step. 
+   % The domain adaptation with the gradient reverse method was proposed in~\cite{ganin2014unsupervised}.
 
@@
 
-    The influence of representations produced by a neural network to internal noise reduction 
-    is discussed in~\cite{yu2013feature} and this work sets a baseline for experiments on
-    Aurora-4 dataset.
+  %  The influence of representations produced by a neural network to internal noise reduction 
+  % is discussed in~\cite{yu2013feature} and this work sets a baseline for experiments on
+   % Aurora-4 dataset.
 
-    In a work~\cite{yusuke2016adversarial} an acoustic model: multilayer sigmoidal network is 
-    trained in an adversarial fashion for a task of dataset corrupted with additive noise.
+   % In a work~\cite{yusuke2016adversarial} an acoustic model: multilayer sigmoidal network is 
+   % trained in an adversarial fashion for a task of dataset corrupted with additive noise.
 
@@
 We experimentally evaluate the approach discussed in Section~\ref{sec:gradient-reverse} 
-on the Aurora-4~\cite{parihar2002aurora} noisy speech recognition dataset. Aurora-4
+on the well-benchmarked Aurora-4~\cite{parihar2002aurora} noisy speech recognition task. Aurora-4
 is based on the Wall Street Journal corpus (WSJ0). It contains noises of 
-six categories which was added to clean
-data. Every clean and noisy utterance is filtered to simulate the frequency characteristics of the sampling frequency. 
+six categories which was added to clean data. Every clean and noisy utterance is filtered to simulate the frequency characteristics %%% of the sampling frequency. ?????  don;t follow % bhuv 
 % Dima: is it correct now?
@@
 The training
-set contains approximately 4400 clean utterances and 446 utterances of each noise condition,
+data  contains 4400 clean utterances and 446 utterances for each noise condition,
 i.e. a total of 2676 noisy utterances.
-The testing
-set contains clean data, data corrupted by the 6 noises, and data recorded with a different microphone
-for both clean and noisy cases.
+The test set consists of clean data, data corrupted by 6 noise types, and data recorded with a different microphone for both clean and noisy cases.
 
-We extracted 40-dimensional Mel-filterbank features with their deltas and 
+For both clean and noisy data, we extract 40-dimensional Mel-filterbank features with their deltas and 
 delta-deltas spliced over $\pm$5 frames, resulting in 1320 input 
-features. We performed speaker-independent global mean and variance normalization on these features. We trained 6 layer
-DNN with 2048 rectified linear units at every layer using momentum-accelerated stochastic gradient descent for 15 epochs. We annealed the learning rate according 
-to the development set cross-entropy, dividing it by 2 when the cross entropy did not improve.
+features that are subsequently mean and variance normalized.  The baseline acoustic model is a  6-layer 
+DNN with 2048 rectified linear units at every layer. It is trained using momentum-accelerated stochastic gradient descent for 15 epochs witj new-bob annealing.
+
+%We annealed the learning rate according 
+%to the development set cross-entropy, dividing it by 2 when the cross entropy did not improve.
 
 In order to evaluate the impact of our method on generalization to unseen noises,
-we performed 6 experiments with different set of seen noises. We always train
-on clean data and add noise conditions one-by-one in the following order: airport, babble, car, 
-restaurant, street, train. The last training group includes all noises therefore matches the
-standard multi-condition training setup. For every group of experiments we trained the
+we performed 6 experiments with different set of seen noises. The networks are trained
+on clean data, with each noise condition added one-by-one in the following order: airport, babble, car, 
+restaurant, street, and train. The last training group includes all noises therefore matches the
+standard multi-condition training setup. For every training group, we trained the
 baseline and the invariance model where we branch out at the $4^{th}$ layer to an  
 binary classifier predicting clean versus noisy data. Due to the imbalance between amounts of clean and
-noisy utterances, we had to oversample noisy frames. Hence every mini-batch contained
+noisy utterances, we had to oversample noisy frames to ensure that every mini-batch contained
 equal number of clean and noisy speech frames.
@@
 gain is achieved when a small number of noise types are available during training.
-We conclude that invariance training is able to generalize better to unseen noise types compared with multi-condition training.
+It can be seen that invariance training is able to generalize better to unseen noise types compared with multi-condition training.
 
-We note that our experiments did not use layer-wise pre-training not uncommonly used for small
-datasets due to the difficulties of combining it with the invariance training. Our preliminary
-experiments on a pre-trained network when using all noise types (last row of Table~\ref{tab:results}) for training show modest gains in WER over the multi-condition baseline.
+We note that our experiments did not use layer-wise pre-training, commonly used for small
+datasets. The baseline WERs reported are very close to the state-of-the-art. %due to the difficulties of combining it with the invariance training.
+Our preliminary experiments on a pre-trained network (better overall WER) when using all noise types (last row of Table~\ref{tab:results}) for training show the same trend as the non-pretrained networks.
 
@@
 \section{Discussion}
+\label{sec:discussion}
     This paper presents the application of generative adversarial networks and invariance training for noise robust speech recognition.
-We show that invariance training helps the ASR system to generalize to better to unseen noise
-    conditions and improves word error rate in when a small number of noise types are seen during training. Future
-    research on this topic includes experimentation with layer-wise pre-training,
-    exploring alternative network architectures and invariance-promoting loss functions
+We show that invariance training helps the ASR system to generalize better to unseen noise conditions and improves word error rate when a small number of noise types are seen during training.
+Our experiments show that in contrast to the image recognition task, in speech recognition,  the domain adaptation network suffers from underfitting.
+Therefore, the gradient of the $L_3$ term in Eq.~\ref{eq:cost} is unreliable and noisy.
+Future research includes enhancements to the domain adaptation network while exploring alternative network architectures and invariance-promoting loss functions.
 
@@
     % Dima: tried to fix -- it is differnt => needs more investigation
-    Our experiments show that in contrast to the image recognition task,
-    the task of speech recognition,  the domain adaptation network suffers from
-    the underfitting. Therefore, the gradient of the $L_3$ term in Eq.~\ref{eq:cost}
-    unreliable and noisy. This aspect needs further development.
 

### 2016-10-24 paper.tex
Fix typos, remove eq 2

@@
     %Even in case all testing conditions have also appeared
-    i%n the training data, their acoustic statistics likely to be
+    %in the training data, their acoustic statistics likely to be
     %different. For this reason, it's important to develop ASR systems that are
@@
     %representations.
-    Several model and feature based adaptation methods such as Maximum Likelihood Linear Regression (MLLR),  feature-based MLLR and iVectors~\cite{ivec} have been proposed to handle speaker variability; and Noise Adaptive Training (NAT)~\cite{msftnoise}  and Vector Taylor Series (VTS) to handle environment variability.
-With the increasing success of Deep Neural Network (DNN) acoustic models for ASR \cite{hinton2012deep, msft, ibm}, end-to-end systems are being proposed~\cite{essen, scatterTransforms, fromaudio} for modeling the acoustic conditions within a single network.
-This allows us to take advantage of the network's ability to learn highly non-linear feature transformations, with greater flexibility in constructing training objective functions that promote learning of noise invariant representations.
-The main idea of this work is to force the acoustic model to learn a representation invariant to noise conditions, instead of explicitly using noise robust acoustic features (Section~\ref{sec:invariant-speech}).
-This type of noise-invariant training requires noise-condition labels during training only.
-It is related to the idea of generative adversarial networks (GAN) and the gradient reverse method proposed in proposed  in~\cite{goodfellow2014generative}  and ~\cite{ganin2014unsupervised} respectively (Section~\ref{sec:relatedwork}).
-We present results on the Aurora-4 speech recognition task ( Section~\ref{sec:experiments} and summarize our findings in Section~\ref{sec:discussion}.
+    Several model and feature based adaptation methods such as Maximum Likelihood Linear Regression (MLLR),  feature-based MLLR 
+    and iVectors~\cite{ivec} have been proposed to handle speaker variability; and Noise Adaptive Training (NAT)~\cite{msftnoise}  
+    and Vector Taylor Series (VTS) to handle environment variability. With the increasing success of Deep Neural Network (DNN) 
+    acoustic models for ASR \cite{hinton2012deep, msft, ibm}, end-to-end systems are being proposed~\cite{essen, scatterTransforms, fromaudio} 
+    for modeling the acoustic conditions within a single network. This allows us to take advantage of the network's ability to learn 
+    highly non-linear feature transformations, with greater flexibility in constructing training objective functions that promote learning 
+    of noise invariant representations. The main idea of this work is to force the acoustic model to learn a representation invariant to 
+    noise conditions, instead of explicitly using noise robust acoustic features (Section~\ref{sec:invariant-speech}). This type of 
+    noise-invariant training requires noise-condition labels during training only. It is related to the idea of generative adversarial 
+    networks (GAN) and the gradient reverse method proposed in~\cite{goodfellow2014generative} and~\cite{ganin2014unsupervised} respectively 
+    (Section~\ref{sec:relatedwork}). We present results on the Aurora-4 speech recognition task in Section~\ref{sec:experiments} and summarize 
+    our findings in Section~\ref{sec:discussion}.
 
@@
 \label{sec:relatedwork}
-{\bf Generative Adversarial Networks} consist of two networks: generator and discriminator. 
+\emph{Generative Adversarial Networks} consist of two networks: generator and discriminator. 
     %One of the most insightful recent developments in the area of generative models
@@
     The generator network $G$ has an
-    input of randomly-generated feature vectors and is asked to produce data, e.g. an image, 
+    input of randomly-generated feature vectors and is asked to produce a sample, e.g. an image, 
     similar to the images in the training set. The discriminator network $D$
@@
     the discriminator is just a classifier network with a sigmoid output layer
-    and can be trained with gradient backpropagation.
+    and can be trained with gradient backpropagation. This gradient can be propagated further
+    to the generator network.
 
-    The main idea of generative adversarial
-    training is to stack the discriminator and generator network on top of each other and propagate the gradient of the discriminator
-    network further to the generator network. In other words, these two networks are competing with each other. The 
+    Two networks in the GAN setup are competing with each other: the 
     generator is trying to deceive the discriminator network, while the discriminator tries
@@
     classifiers as the training ones.
+
+    Several practical guidelines were proposed for optimizing GANs in~\cite{radford2015unsupervised} and 
+    further explored in~\cite{salimans2016improved}.
 	%{\bf Please explain how the min-max applies to the two terms.
@@
    % to have the domain labels. 
-   The network has a fork-like structure: the image is fed to the
+    The network has a $Y$-like structure: the image is fed to the
     first network which produces a hidden representation $h$. Then this representation $h$ is input to two separate networks: a domain classifier network (D) and 
@@
     interfere with the target classifier at test time. Similar to the GAN objective, which forces the generation distribution be close to the data distribution,
-    the {\bf gadient reverse method} makes domain distributions similar to each other.
+    the \emph{gadient reverse method} makes domain distributions similar to each other.
 
@@
     \begin{equation}
-        L = L(\hat{y}, y; \theta_R, \theta_E) + 
-        \alpha L(\hat{d}, d; \theta_D) -
-        \beta L(\hat{d}, d; \theta_E),
+        L = L_1(\hat{y}, y; \theta_R, \theta_E) + 
+        \alpha L_2(\hat{d}, d; \theta_D) -
+        \beta L_3(\hat{d}, d; \theta_E),
         \label{eq:grm}
@@
 The influence of representations produced by a neural network to internal noise reduction is discussed in~\cite{yu2013feature} and this work sets a baseline for experiments on  Aurora-4 dataset.
-Recently, in~\cite{yusuke2016adversarial} a multilayer sigmoidal network is trained in an adversarial fashion on a in-house transcription task corrupted by noise.
+Recently, in~\cite{yusuke2016adversarial} a multilayer sigmoidal network is trained in an adversarial fashion on an in-house transcription task corrupted by noise.
     %% STILL TODO
@@
 used the noise condition as the domain label merging all noise types into one label
-and clean as the other label. Our training loss function is
-    \begin{equation}
-        L = L_1(\hat{y}, y; \theta_R, \theta_E) + 
-        L_2(\hat{d}, d; \theta_D) +
-        \beta L_3(\hat{d}, d; \theta_E),
-        \label{eq:cost}
-    \end{equation}
-where $L_3 = - d\log(1 - \hat{d}) - (1-d)\log(\hat{d})$ maximizes the probability
+and clean as the other label. Our training loss function is  Eq.~\ref{eq:grm} with $\alpha$ set to 1 and
+$L_3$ to $-d\log(1 - \hat{d}) - (1-d)\log(\hat{d})$. $L_3$ term maximizes the probability
 of an incorrect domain classification in contrast to the Eq.~\ref{eq:grm} where the 
@@
 \label{sec:experiments}
-We experimentally evaluate the approach discussed in Section~\ref{sec:gradient-reverse} 
-on the well-benchmarked Aurora-4~\cite{parihar2002aurora} noisy speech recognition task. Aurora-4
+We experimentally evaluate our approach   
+on the well-benchmarked Aurora-4 \citep{parihar2002aurora} noisy speech recognition task. Aurora-4
 is based on the Wall Street Journal corpus (WSJ0). It contains noises of 
@@
 features that are subsequently mean and variance normalized.  The baseline acoustic model is a  6-layer 
-DNN with 2048 rectified linear units at every layer. It is trained using momentum-accelerated stochastic gradient descent for 15 epochs witj new-bob annealing.
+DNN with 2048 rectified linear units at every layer. It is trained using momentum-accelerated stochastic gradient descent for 15 epochs with new-bob annealing.
 
@@
 \label{sec:discussion}
-    This paper presents the application of generative adversarial networks and invariance training for noise robust speech recognition.
-We show that invariance training helps the ASR system to generalize better to unseen noise conditions and improves word error rate when a small number of noise types are seen during training.
-Our experiments show that in contrast to the image recognition task, in speech recognition,  the domain adaptation network suffers from underfitting.
-Therefore, the gradient of the $L_3$ term in Eq.~\ref{eq:cost} is unreliable and noisy.
+    This paper presents the application of generative adversarial networks and invariance training for noise robust speech recognition. We show that invariance training 
+    helps the ASR system to generalize better to unseen noise conditions and improves word error rate when a small number of noise types are seen during training. Our 
+    experiments show that in contrast to the image recognition task, in speech recognition,  the domain adaptation network suffers from underfitting. Therefore, the 
+    gradient of the $L_3$ term in Eq.~\ref{eq:grm} is unreliable and noisy.
 Future research includes enhancements to the domain adaptation network while exploring alternative network architectures and invariance-promoting loss functions.

### 2016-10-24 paper.tex
Added all references

@@
 \begin{abstract}
-    Modern automatic speech recognition (ASR) systems need to be robust under acoustic variability arising from environmental, speaker, channel and recording conditions.
+    Modern automatic speech recognition (ASR) systems need to be robust under acoustic variability arising from environmental, speaker, channel, and recording conditions.
 Ensuring such robustness to variability is a challenge in modern day neural network-based ASR systems, especially when all types of variability are not seen during training.
@@
     Several model and feature based adaptation methods such as Maximum Likelihood Linear Regression (MLLR),  feature-based MLLR 
-    and iVectors~\cite{ivec} have been proposed to handle speaker variability; and Noise Adaptive Training (NAT)~\cite{msftnoise}  
-    and Vector Taylor Series (VTS) to handle environment variability. With the increasing success of Deep Neural Network (DNN) 
-    acoustic models for ASR \cite{hinton2012deep, msft, ibm}, end-to-end systems are being proposed~\cite{essen, scatterTransforms, fromaudio} 
+    and iVectors~\citep{saon2013speaker} have been proposed to handle speaker variability; and Noise Adaptive Training \citep[NAT;][]{kalinli2010noise}
+    and Vector Taylor Series \citep[VTS;][]{un1998speech} to handle environment variability. With the increasing success of Deep Neural Network (DNN) 
+    acoustic models for ASR \citep{hinton2012deep,seide2011conversational,sainath2011making}, end-to-end systems are being proposed~\citep{miao2015eesen,sainath2015learning} 
     for modeling the acoustic conditions within a single network. This allows us to take advantage of the network's ability to learn 
@@
     interfere with the target classifier at test time. Similar to the GAN objective, which forces the generation distribution be close to the data distribution,
-    the \emph{gadient reverse method} makes domain distributions similar to each other.
+    the \emph{gradient reverse method} makes domain distributions similar to each other.
 
@@
 and clean as the other label. Our training loss function is  Eq.~\ref{eq:grm} with $\alpha$ set to 1 and
-$L_3$ to $-d\log(1 - \hat{d}) - (1-d)\log(\hat{d})$. $L_3$ term maximizes the probability
-of an incorrect domain classification in contrast to the Eq.~\ref{eq:grm} where the 
+$L_3$ to $d\log(1 - \hat{d}) + (1-d)\log(\hat{d})$. $L_3$ term maximizes the probability
+of an incorrect domain classification in contrast to the gradient reverse where the 
 correct classification is minimized.
@@
         is the number of noise conditions used for the training. The last row is a 
-        preliminary experiment with layer-wise pretraining close to state-of-the-art
+        preliminary experiment with layer-wise pre-training close to state-of-the-art
         model and a corresponding invariance training starting with a pretrained model.}

### 2016-11-27 paper.tex
Small fixes
Update numbers

@@
   Yorktown Heights, NY \\
-  \texttt{kaudhkha@us.ibm.com} \\
   \And
@@
   %Montr\'eal, QC H3T 1J4 \\
-  \texttt{pbpop3@gmail.com} \\
   \And
@@
   Yorktown Heights, NY \\
-  \texttt{bhuvana@us.ibm.com} \\
   \And
@@
   Yorktown Heights, NY \\
-  \texttt{samthom@us.ibm.com} \\
   \And
@@
   MILA, CIFAR Fellow,
-  Universit\'e de Montr\'eal,
+  Universit\'e de Montr\'eal
   %Montr\'eal, QC H3T 1J4 \\
-  \texttt{findme@inter.net} \\
 }
@@
 used the noise condition as the domain label merging all noise types into one label
-and clean as the other label. Our training loss function is  Eq.~\ref{eq:grm} with $\alpha$ set to 1 and
-$L_3$ to $d\log(1 - \hat{d}) + (1-d)\log(\hat{d})$. $L_3$ term maximizes the probability
+and clean as the other label. Our training loss function is Eq.~\ref{eq:grm} with 
+$L_3$ set to $d\log(1 - \hat{d}) + (1-d)\log(\hat{d})$ for stability during training. 
+$L_3$ term maximizes the probability
 of an incorrect domain classification in contrast to the gradient reverse where the 
 correct classification is minimized.
-%{\bf Contrast this with term 3 of the original gradient reverse equation.}. 
 The terms $L_1$ and $L_2$ are 
@@
             conditions used for training. Bottom: Average performance on seen versus unseen noise conditions.
-            Testing was performed on all wv1 conditions.
+            Testing was performed on all wv1 conditions (Sennheiser microphone).
             }
@@
 features that are subsequently mean and variance normalized.  The baseline acoustic model is a  6-layer 
-DNN with 2048 rectified linear units at every layer. It is trained using momentum-accelerated stochastic gradient descent for 15 epochs with new-bob annealing.
+DNN with 2048 rectified linear units at every layer. It is trained using momentum-accelerated stochastic gradient descent for 15 epochs with new-bob 
+annealing~\citep[as in][]{morgan1995continuous,sainath2011making}.
 
@@
     \hline\hline
-    6* &11.88        &11.99
+    6* &11.85        &11.99    &4.52    &4.76    &8.76    &8.76    &7.79    &8.57    &16.84&    16.99
     \end{tabular}

### 2016-11-30 poster.tex
Add poster

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Jacobs Landscape Poster
% LaTeX Template
% Version 1.0 (29/03/13)
%
% Created by:
% Computational Physics and Biophysics Group, Jacobs University
% https://teamwork.jacobs-university.de:8443/confluence/display/CoPandBiG/LaTeX+Poster
% 
% Further modified by:
% Nathaniel Johnston (nathaniel@njohnston.ca)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final]{beamer}

\usepackage[scale=1.24]{beamerposter} % Use the beamerposter package for laying out the poster

\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamercolor{block title}{fg=black,bg=white} % Colors of the block titles

\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks
% Many more colors are available for use in beamerthemeconfposter.sty

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{48in} % A0 width: 46.8in
\setlength{\paperheight}{36in} % A0 height: 33.1in
\setlength{\sepwid}{0.024\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.22\paperwidth} % Width of one column
\setlength{\twocolwid}{0.464\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

\usepackage{graphicx}  % Required for including images

\usepackage{booktabs} % Top and bottom rules for tables
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{amsmath,graphicx}
\usepackage{bm}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
%\usepackage{subcaption}
\usepackage{algorithm2e}
\usepackage{pbox}
\usepackage{multirow}

\usepackage{booktabs, multicol, multirow}
\usepackage{dirtree}
\usepackage{booktabs}
\usepackage{graphicx} 
\usepackage{amsmath}
\usepackage{booktabs, multicol, multirow}
\usepackage{pbox}
\usepackage{algorithm2e}
\usepackage[export]{adjustbox}
\renewcommand\DTstyle{}

\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\argmin}{argmin}

\newcounter{savedenum}
\newcommand*{\saveenum}{\setcounter{savedenum}{\theenumi}}
\newcommand*{\resume}{\setcounter{enumi}{\thesavedenum}}

%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------

\title{Invariant Representations for Noisy Speech Recognition} % Poster title

\author{Dmitriy Serdyuk${}^{12}$, Kartik Audhkhasi${}^2$, Phil\'emon Brakel${}^1$,
    Bhuvana Ramabhadran${}^2$, Samuel Thomas${}^2$, \\
    Yoshua Bengio${}^{1 \dagger}$}

    \institute{${}^1$ MILA, Universit\'{e} de Montr\'{e}al, ${}^2$ IBM Watson, ${}^\dagger$ CIFAR Fellow} % Institution(s)

%----------------------------------------------------------------------------------------

\begin{document}
\addtobeamertemplate{headline}{} 
{
\begin{tikzpicture}[remember picture,overlay] 
\node [shift={(-13 cm,-7cm)}] at (current page.north east)
% \node [shift={(-10 cm,-5cm)}] at (current page.north west)
{\includegraphics[height=7cm]{ibmpos_blue}}; 
%\node [shift={(-42 cm,-11cm)}] at (current page.north)
%{\includegraphics[height=7cm]{UdeM_CMJK.eps}}; 
\node [shift={(-49 cm,-6cm)}] at (current page.north)
% \node [shift={(-10 cm,-5cm)}] at (current page.north west)
{\includegraphics[height=9cm]{MILA_2C.eps}}; 
\end{tikzpicture} 
}

\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\begin{columns}[t] % The whole poster consists of three major columns, the second of which is split into two columns twice - the t option aligns each column's content to the top

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The first column

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

\begin{block}{Introduction}
    \begin{itemize}
        \item 
    Modern automatic speech recognition (ASR) systems need to be robust under 
    acoustic variability arising from environmental, speaker, channel, and recording 
    conditions. Ensuring such robustness to variability is a challenge in modern day 
    neural network-based ASR systems, especially when all types of variability are 
    not seen during training.

        \item We attempt to address this problem by encouraging the 
    neural network acoustic model to learn invariant feature representations.

        \item We use ideas from recent research on image generation using
    Generative Adversarial Networks and domain adaptation ideas extending
    adversarial gradient-based training. A recent work from Ganin et al. proposes to
    use adversarial training for image domain adaptation by using an intermediate
    representation from the main target classification network to deteriorate the 
    domain 
    classifier performance through a separate neural network.
    Our work focuses on investigating neural architectures which produce
    representations invariant to noise conditions for ASR.  
        \item We evaluate the proposed 
    architecture on the Aurora-4 task, a popular benchmark for
    noise robust ASR. We show that our method generalizes better than the standard 
    multi-condition training especially when only a few noise categories are seen 
    during training.
    \end{itemize}

\end{block}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\begin{block}{Generative Adversarial Networks}
    \begin{itemize}
        \item The generator network $G$ has an
    input of randomly-generated feature vectors and is asked to produce a sample, e.g. an image, 
    similar to the images in the training set. The discriminator network $D$
    can either receive a generated image from the generator $G$ or an image
    from the training set. Its task is to distinguish
    between the ``fake'' generated image and the ``real'' image taken from the dataset~\cite{goodfellow2014generative}. 
    \begin{align*}
        \min_G \max_D V(D, G) =& \mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
            &\mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
    \end{align*}

    \end{itemize}

    \begin{block}{Domain Adaptation with Reverse Gradient}
        \begin{itemize}
    \item 
    Prior work by~\cite{ganin2014unsupervised} proposed a method of training a network 
    which can be adapted to new domains. The training data consists of the images
    labeled with classes of interest and separate domain (image background) labels. 
    %For example, the domain may refer to image background. The testing set does not need
   % to have the domain labels. 
    The network has a $Y$-like structure: the image is fed to the
    first network which produces a hidden representation $h$. Then this representation $h$ is input to two separate networks: a domain classifier network (D) and 
    a target classifier network (R). The goal of training is to learn the hidden 
    representation that is invariant to the domain labels and performs well on the target classification task, so that the domain information doesn't 
    interfere with the target classifier at test time. Similar to the GAN objective, which forces the generation distribution be close to the data distribution,
    the \emph{gradient reverse method} makes domain distributions similar to each other.

        \end{itemize}

    \end{block}
\end{block}

\end{column} % End of the first column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\twocolwid}

    \begin{columns}[t]

        \begin{column}{\onecolwid} % The second column
            \begin{block}{{\Large Invariant Representation Network}}
                \begin{itemize}
                    \item The model consists of three neural networks. The encoder $E$ produces
                        the intermediate representation $h$ which used in the recognizer $R$ and 
                        in the domain discriminator $D$. The hidden representation $h$ is trained to improve
                        the recognition and minimize the domain discriminator accuracy. The domain discriminator
                        is a classifier trained to maximize its accuracy on the noise type
                        classification task.
                \end{itemize}
                \begin{equation*}
                    L = L_1(\hat{y}, y; \theta_R, \theta_E) + 
                    \alpha L_2(\hat{d}, d; \theta_D) -
                    \beta L_3(\hat{d}, d; \theta_E)
                \end{equation*}

                \begin{figure}
                    \centering
                    \includegraphics[width=0.9\linewidth]{model.pdf}
                \end{figure}

                \begin{itemize}
                    \item The model is a DNN-HMM system with a feed-forward neural network trained to predict the 
                        CD HMM state (2000 classes).
                    \item Input: 40-dimensional Mel-filterbank features with their deltas and 
                        delta-deltas spliced over $\pm$5 frames (total dimension is 1320).
                    \item Baseline: 6-layer DNN with 2048 rectified linear units at every layer.
                    \item The right branch that predicts the domain (noise condition). This branch is discarded 
                        in the testing phase. 
                    \item The noise condition is the domain label. We are merging all noise types into one label
                        and clean as the other label.
                    \item The invariance term is
                        \begin{equation*}
                            L_3 = d\log(1 - \hat{d}) + (1-d)\log(\hat{d}),
                        \end{equation*}
                        where $d$ is the domain label and $\hat{d}$ is the network prediction.
                \end{itemize}
            \end{block}

        \end{column} % end of second column

        \begin{column}{\sepwid}\end{column} % Empty spacer column

        \begin{column}{\onecolwid} % The third column
            \begin{block}{\Large Dataset}
                \begin{itemize}
                    \item Aurora-4 is based on the Wall Street Journal corpus (WSJ0). 
                    \item It contains noises of 
                        6 categories which was added to clean data. Every clean and noisy utterance is 
                        filtered to simulate the frequency characteristics. 
                    \item The training data: 
                        \begin{itemize}
                            \item One microphone: Sennheiser.
                            \item 4400 clean utterances.
                            \item 446 utterances for each of 6 noise condition.
                            \item Total of 2676 noisy utterances.
                        \end{itemize}
                    \item The testing set:
                        \begin{itemize}
                            \item Two microphones.
                            \item 330 utterances for each condition, each microphone.
                        \end{itemize}
                \end{itemize}
            \end{block}

            \begin{block}{{\Large Experiments on Aurora-4}}
                In order to evaluate the impact of our method on generalization to unseen noises,
                we performed 6 experiments with different set of seen noises. The networks are trained
                on clean data, with each noise condition added one-by-one in the following order: airport, babble, car, 
                restaurant, street, and train.

                Average performance of the baseline multi-condition and invariance model varying with the 
                number of noise conditions used for training. Bottom: Average performance on seen versus 
                unseen noise conditions.
                Testing was performed on all wv1 conditions (Sennheiser microphone).
                \begin{figure}
                    \centering
                    \includegraphics[width=\linewidth]{wer_avg.pdf}
                \end{figure}
                \begin{figure}
                    \centering
                    \includegraphics[width=\linewidth]{wer_seen_unseen.pdf}
                \end{figure}
            \end{block}
        \end{column} % end of third column
    \end{columns}
\end{column}

\begin{column}{\sepwid}\end{column} % Empty spacer column
\begin{column}{\onecolwid} % the fourth column
\begin{block}{Results}

    Average word error rate (WER\%) on Aurora-4 dataset on all test conditions,
        including seen and unseen noise and unseen microphone. First column
        is the number of noise conditions used for the training. The last row is a 
        preliminary experiment with layer-wise pre-training close to state-of-the-art
        model and a corresponding invariance training starting with a pretrained model.
\begin{table}
    \centering
    \label{tab:results}
    \begin{tabular}{r|cc||cc|cc|cc|cc}
        Noise       &Inv&BL&  \multicolumn{2}{c|}{A} & \multicolumn{2}{c|}{B} & \multicolumn{2}{c|}{C} & \multicolumn{2}{c}{D}\\
        cond  & & &  Inv & BL & Inv & BL & Inv & BL & Inv & BL\\
    \hline
    1           &16.36        &18.14 &6.54&7.57    &12.71& 14.09   & 11.45&   13.10    & 22.47 &   24.80    \\
    2           &15.56        &17.39 &5.90&  6.58 &   11.69   &13.28   &11.12   &13.51   &21.79   &23.96 \\
    3           &14.24        &14.67 &5.45 & 5.08&    10.76&   12.44&   9.75&    9.84 &   19.93&   19.30\\
    4           &13.61        &13.84 & 5.08 &5.29    &9.73    &9.97    &9.49    &9.56    &19.49   &19.90\\         
    5           &13.41        &13.02 & 5.12 &5.34    &9.52    &9.42    &9.55    &8.67    &19.33   &18.65\\         
    6           &12.62        &12.60 & 4.80 &4.61    &9.04    &8.86    &8.76    &8.59    &18.16   &18.21\\
    \hline\hline
    6*          &11.85        &11.99 &4.52    &4.76    &8.76    &8.76    &7.79    &8.57    &16.84&    16.99
    \end{tabular}
\end{table}
\end{block}

\begin{block}{Conclusions}
    \begin{itemize}
        \item We show that invariance training helps the ASR system to generalize better to unseen noise 
            conditions and improves word error rate when a small number of noise types are seen during 
            training.

        \item Related work~\cite{yusuke2016adversarial} investigates a similar approach for domain 
            adaptation to noise types and the SNR levels on an in-house dataset based on WSJ.

        \item Our experiments show that in contrast to the image recognition task, in speech recognition, the 
    gradient of the $L_3$ term is unreliable and noisy.

        \item Future research includes enhancements to the domain adaptation network while exploring 
            alternative network architectures and invariance-promoting loss functions.
    \end{itemize}
\end{block}
%----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------
\begin{block}{References}
    \bibliographystyle{amsalpha}
    \bibliography{refs.bib}
\end{block}
%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color

%----------------------------------------------------------------------------------------
%	CONTACT INFORMATION
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------

\end{column} % End of the third column

\end{columns} % End of all the columns in the poster

\end{frame} % End of the enclosing frame

\end{document}

### 2016-11-30 paper.tex
Itemize gradient reverse and network architecture sections

@@
 is based on the Wall Street Journal corpus (WSJ0). It contains noises of 
-six categories which was added to clean data. Every clean and noisy utterance is filtered to simulate the frequency characteristics %%% of the sampling frequency. ?????  don;t follow % bhuv 
-% Dima: is it correct now?
-%
-The training
+six categories which was added to clean data. Every clean and noisy utterance is filtered to simulate the frequency characteristics.The training
 data  contains 4400 clean utterances and 446 utterances for each noise condition,
@@
 
-    %{\bf We need to make the message of this paragraph more clear.}
-    % Dima: tried to fix -- it is differnt => needs more investigation
-
 \section*{Acknowledgments}

### 2016-11-30 poster.tex
Itemize gradient reverse and network architecture sections

@@
     \begin{itemize}
+        \item Two networks: the generator and the discriminator.
         \item The generator network $G$ has an
     input of randomly-generated feature vectors and is asked to produce a sample, e.g. an image, 
-    similar to the images in the training set. The discriminator network $D$
+    similar to the images in the training set.
+
+        \item The discriminator network $D$
     can either receive a generated image from the generator $G$ or an image
@@
     \begin{align*}
-        \min_G \max_D V(D, G) =& \mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
-            &\mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
+        \min_G \max_D V(D, G) = & \mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
+            & \mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
     \end{align*}
@@
             Prior work by~\cite{ganin2014unsupervised} proposed a method of training a network 
-    which can be adapted to new domains. The training data consists of the images
-    labeled with classes of interest and separate domain (image background) labels. 
-    %For example, the domain may refer to image background. The testing set does not need
-   % to have the domain labels. 
-    The network has a $Y$-like structure: the image is fed to the
-    first network which produces a hidden representation $h$. Then this representation $h$ is input to two separate networks: a domain classifier network (D) and 
-    a target classifier network (R). The goal of training is to learn the hidden 
-    representation that is invariant to the domain labels and performs well on the target classification task, so that the domain information doesn't 
-    interfere with the target classifier at test time. Similar to the GAN objective, which forces the generation distribution be close to the data distribution,
+            which can be adapted to new domains.
+
+            \item $Y$-like network: the image is fed to the
+            encoder network ($E$) which produces a hidden representation $h$.
+
+            \item This representation $h$ is input to two separate networks: a domain classifier network ($D$) and 
+            a target classifier network ($R$).
+
+            \item The goal of training is to learn the hidden 
+            representation that is invariant to the domain labels and performs well on the target classification 
+            task.
+
+            \item Similar to the GAN objective, which forces the generation distribution be close to the data distribution,
             the \emph{gradient reverse method} makes domain distributions similar to each other.
-
         \end{itemize}
@@
                 \begin{itemize}
-                    \item The model consists of three neural networks. The encoder $E$ produces
+                    \item The model consists of three neural networks. 
+                    \item The encoder $E$ produces
                         the intermediate representation $h$ which used in the recognizer $R$ and 
-                        in the domain discriminator $D$. The hidden representation $h$ is trained to improve
-                        the recognition and minimize the domain discriminator accuracy. The domain discriminator
+                        in the domain discriminator $D$. 
+                    \item The hidden representation $h$ is trained to improve
+                        the recognition and minimize the domain discriminator performance. 
+                    \item The domain discriminator
                         is a classifier trained to maximize its accuracy on the noise type
@@
                 Average performance of the baseline multi-condition and invariance model varying with the 
-                number of noise conditions used for training. Bottom: Average performance on seen versus 
+                number of noise conditions used for training. Average performance on seen versus 
                 unseen noise conditions.
@@
 %----------------------------------------------------------------------------------
-\begin{block}{References}
+\begin{block}{Important References}
     \bibliographystyle{amsalpha}

### 2016-11-30 poster.tex
Prettify experiments section

@@
             \begin{block}{{\Large Experiments on Aurora-4}}
-                In order to evaluate the impact of our method on generalization to unseen noises,
-                we performed 6 experiments with different set of seen noises. The networks are trained
+                \begin{itemize}
+                    \item We performed 6 experiments with different set of seen noises. 
+                    \item The networks are trained
                 on clean data, with each noise condition added one-by-one in the following order: airport, babble, car, 
-                restaurant, street, and train.
+                restaurant, street, and train:
+                    \begin{enumerate}
+                        \item clean + airport
+                        \item clean + airport + babble
+                        \item \ldots
+                    \end{enumerate}
+                \end{itemize}
 
-                Average performance of the baseline multi-condition and invariance model varying with the 
-                number of noise conditions used for training. Average performance on seen versus 
-                unseen noise conditions.
-                Testing was performed on all wv1 conditions (Sennheiser microphone).
                 \begin{figure}
@@
                 \end{figure}
+                Average performance of the baseline multi-condition and invariance model varying with the 
+                number of noise conditions used for training. 
                 \begin{figure}
@@
                 \end{figure}
+                Average performance on seen versus 
+                unseen noise conditions.
+                Testing was performed on all wv1 conditions (Sennheiser microphone).
             \end{block}

### 2016-11-30 poster.tex
Improve introduction

@@
         \item 
-    Modern automatic speech recognition (ASR) systems need to be robust under 
-    acoustic variability arising from environmental, speaker, channel, and recording 
-    conditions. Ensuring such robustness to variability is a challenge in modern day 
+            \textcolor{dblue}{Robustness} to the acoustic variability arising from environmental, speaker, channel, 
+    and recording 
+    conditions is a challenge in modern day 
     neural network-based ASR systems, especially when all types of variability are 
@@
         \item We attempt to address this problem by encouraging the 
-    neural network acoustic model to learn invariant feature representations.
+            neural network acoustic model to learn \textcolor{dblue}{invariant} feature representations.
 
         \item We use ideas from recent research on image generation using
-    Generative Adversarial Networks and domain adaptation ideas extending
-    adversarial gradient-based training. A recent work from Ganin et al. proposes to
-    use adversarial training for image domain adaptation by using an intermediate
-    representation from the main target classification network to deteriorate the 
-    domain 
-    classifier performance through a separate neural network.
-    Our work focuses on investigating neural architectures which produce
-    representations invariant to noise conditions for ASR.  
+            \textcolor{dblue}{generative adversarial networks} and domain adaptation ideas extending
+    adversarial gradient-based training.  
         \item We evaluate the proposed 
     architecture on the Aurora-4 task, a popular benchmark for
-    noise robust ASR. We show that our method generalizes better than the standard 
-    multi-condition training especially when only a few noise categories are seen 
-    during training.
+    noise robust ASR. 
     \end{itemize}
@@
     from the training set. Its task is to distinguish
-    between the ``fake'' generated image and the ``real'' image taken from the dataset~\cite{goodfellow2014generative}. 
+    between the ``fake'' generated image and the ``real'' image taken from the dataset~\cite{goodfellow2014generative}:
     \begin{align*}
@@
         \begin{itemize}
-            \item 
-            Prior work by~\cite{ganin2014unsupervised} proposed a method of training a network 
-            which can be adapted to new domains.
-
-            \item $Y$-like network: the image is fed to the
+            \item \cite{ganin2014unsupervised} proposes $Y$-shaped network: the image is fed to the
             encoder network ($E$) which produces a hidden representation $h$.

### 2016-11-30 poster.tex
Add gan picture

@@
     \end{itemize}
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.5\linewidth]{gan.pdf}
+    \end{figure}
+\end{block}
 
@@
         \item Similar to the GAN objective, which forces the generation distribution be close to the data distribution,
-            the \emph{gradient reverse method} makes domain distributions similar to each other.
+        the \textcolor{dblue}{gradient reverse method} makes domain distributions similar to each other.
     \end{itemize}
 
-    \end{block}
 \end{block}

### 2016-12-01 poster.tex
Fix articles, emphasize definitions and important results

@@
 \begin{tikzpicture}[remember picture,overlay] 
-\node [shift={(-13 cm,-7cm)}] at (current page.north east)
+\node [shift={(-13 cm,-6cm)}] at (current page.north east)
 % \node [shift={(-10 cm,-5cm)}] at (current page.north west)
@@
     \begin{itemize}
-        \item Two networks: the generator and the discriminator.
+        \item Two networks: the \textcolor{dblue}{generator} and the \textcolor{dblue}{discriminator}.
         \item The generator network $G$ has an
@@
     \begin{itemize}
-        \item \cite{ganin2014unsupervised} proposes $Y$-shaped network: the image is fed to the
+        \item \cite{ganin2014unsupervised} proposes a $Y$-shaped network: the image is fed to the
         encoder network ($E$) which produces a hidden representation $h$.
@@
 
-        \item The goal of training is to learn the hidden 
+        \item The goal of training is to learn a hidden 
         representation that is invariant to the domain labels and performs well on the target classification 
@@
                 \begin{itemize}
-                    \item The model consists of three neural networks. 
-                    \item The encoder $E$ produces
-                        the intermediate representation $h$ which used in the recognizer $R$ and 
-                        in the domain discriminator $D$. 
+                    \item Our model consists of three neural networks. 
+                    \item The \textcolor{dblue}{encoder} $E$ produces
+                        the intermediate representation $h$ which used in the \textcolor{dblue}{recognizer} $R$ and 
+                        in the \textcolor{dblue}{domain discriminator} $D$. 
                     \item The hidden representation $h$ is trained to improve
@@
 
+                \vspace{2cm}
                 \begin{figure}
                     \centering
-                    \includegraphics[width=0.9\linewidth]{model.pdf}
+                    \includegraphics[width=0.7\linewidth]{model.pdf}
                 \end{figure}
+                \vspace{2cm}
 
@@
                         CD HMM state (2000 classes).
-                    \item Input: 40-dimensional Mel-filterbank features with their deltas and 
+                    \item Input: 40-dimensional mel-filterbank features with their deltas and 
                         delta-deltas spliced over $\pm$5 frames (total dimension is 1320).
@@
 
+                \vspace{1cm}
                 \begin{figure}
@@
                 number of noise conditions used for training. 
+                \vspace{4cm}
                 \begin{figure}
@@
     \label{tab:results}
-    \begin{tabular}{r|cc||cc|cc|cc|cc}
-        Noise       &Inv&BL&  \multicolumn{2}{c|}{A} & \multicolumn{2}{c|}{B} & \multicolumn{2}{c|}{C} & \multicolumn{2}{c}{D}\\
-        cond  & & &  Inv & BL & Inv & BL & Inv & BL & Inv & BL\\
+    \begin{tabular}{r|rr||rr|rr|rr|rr}
+        Noise       &&&  \multicolumn{2}{c|}{A} & \multicolumn{2}{c|}{B} & \multicolumn{2}{c|}{C} & \multicolumn{2}{c}{D}\\
+        cond  & Inv & BL &  Inv & BL & Inv & BL & Inv & BL & Inv & BL\\
     \hline
-    1           &16.36        &18.14 &6.54&7.57    &12.71& 14.09   & 11.45&   13.10    & 22.47 &   24.80    \\
-    2           &15.56        &17.39 &5.90&  6.58 &   11.69   &13.28   &11.12   &13.51   &21.79   &23.96 \\
-    3           &14.24        &14.67 &5.45 & 5.08&    10.76&   12.44&   9.75&    9.84 &   19.93&   19.30\\
-    4           &13.61        &13.84 & 5.08 &5.29    &9.73    &9.97    &9.49    &9.56    &19.49   &19.90\\         
-    5           &13.41        &13.02 & 5.12 &5.34    &9.52    &9.42    &9.55    &8.67    &19.33   &18.65\\         
-    6           &12.62        &12.60 & 4.80 &4.61    &9.04    &8.86    &8.76    &8.59    &18.16   &18.21\\
+    1           &\textcolor{dblue}{16.36}        &18.14 & 6.54  & 7.57    &12.71& 14.09   & 11.45&   13.10    & 22.47 &   24.80    \\
+        2           &\textcolor{dblue}{15.56}        &17.39 & 5.90  & 6.58    &   11.69   &13.28   &11.12   &13.51   &21.79   &23.96 \\
+        3           &\textcolor{dblue}{14.24}        &14.67 & 5.45  & 5.08    &   10.76&   12.44&   9.75&    9.84 &   19.93&   19.30\\
+        4           &\textcolor{dblue}{13.61}        &13.84 & 5.08  & 5.29    &9.73    &9.97    &9.49    &9.56    &19.49   &19.90\\         
+        5           &13.41        &\textcolor{dblue}{13.02} & 5.12  & 5.34    &9.52    &9.42    &9.55    &8.67    &19.33   &18.65\\         
+        6           &12.62        &\textcolor{dblue}{12.60} & 4.80  & 4.61    &9.04    &8.86    &8.76    &8.59    &18.16   &18.21\\
     \hline\hline
-    6*          &11.85        &11.99 &4.52    &4.76    &8.76    &8.76    &7.79    &8.57    &16.84&    16.99
+    6*          &\textcolor{dblue}{11.85}        &11.99 &4.52    &4.76    &8.76    &8.76    &7.79    &8.57    &16.84&    16.99
     \end{tabular}

### 2016-12-01 poster.tex
Make table prettier

@@
     \begin{tabular}{r|rr||rr|rr|rr|rr}
-        Noise       &&&  \multicolumn{2}{c|}{A} & \multicolumn{2}{c|}{B} & \multicolumn{2}{c|}{C} & \multicolumn{2}{c}{D}\\
+        Noise       &\multicolumn{2}{c||}{AVG}&  \multicolumn{2}{c|}{A} & \multicolumn{2}{c|}{B} & \multicolumn{2}{c|}{C} & \multicolumn{2}{c}{D}\\
         cond  & Inv & BL &  Inv & BL & Inv & BL & Inv & BL & Inv & BL\\

### 2017-03-21 interspeech/paper.tex

--- 2016-11-30 paper.tex
@@
-% Template for ICASSP-2017 paper; to be used with:
-%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
-%          IEEEbib.bst - IEEE bibliography style file.
-% --------------------------------------------------------------------------
-\documentclass{article}
+\documentclass[a4paper]{article}
 
-\usepackage[final]{nips_2016}
-
+\usepackage{natbib}
+\usepackage{INTERSPEECH_v2}
 \usepackage[utf8]{inputenc} % allow utf-8 input
@@
 \usepackage{multirow}
+\errorcontextlines 10000
 
-% Example definitions.
-% --------------------
-\def\x{{\mathbf x}}
-\def\L{{\cal L}}
-
-% Title.
-% ------
-\title{Invariant Representations for Noisy Speech Recognition}
-%
-% Single address.
-% ---------------
-%\name{Dmitriy Serdyuk $^{\dagger}$\thanks{Dmitriy Serdyuk performed the work
-%  while at IBM}, Kartik Audhkhasi$^{\star}$, Phil\'emon Brakel$^{\dagger}$, Bhuvana Ramabhadran$^{\star}$, Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
-%\address{$^{\dagger}$ Universit\'e de Montr\'eal\\
-%         $^{\star}$ IBM}
-\author{
-  Dmitriy Serdyuk\thanks{Dmitriy Serdyuk performed the work 
-    during an internship at IBM Watson.} \\
-  MILA,
-  Universit\'e de Montr\'eal\\
-  Montr\'eal, QC H3T 1J4 \\
-  \texttt{serdyuk@iro.umontreal.ca} \\
-  \And
-  Kartik Audhkhasi \\
-  IBM Watson, \\
-  Yorktown Heights, NY \\
-  \And
-  Phil\'emon Brakel \\
-  MILA\\
-  Universit\'e de Montr\'eal\\
-  %Montr\'eal, QC H3T 1J4 \\
-  \And
-  Bhuvana Ramabhadran\\
-  IBM Watson,
-  Yorktown Heights, NY \\
-  \And
-  Samuel Thomas\\
-  IBM Watson,
-  Yorktown Heights, NY \\
-  \And
-  Yoshua Bengio\\
-  MILA, CIFAR Fellow,
-  Universit\'e de Montr\'eal
-  %Montr\'eal, QC H3T 1J4 \\
-}
+\title{Adversarial Training of Invariant Features for Speech Recognition}
+\name{Dmitriy Serdyuk$^1$, Kartik Audhkhasi$^2$, 
+      Bhuvana Ramabhadran$^2$, Phil\'emon Brakel$^1$, Samuel Thomas$^2$,
+      Yoshua Bengio$^{1\dagger}$}
+\address{
+  $^1$MILA, Universit\'e de Montr\'eal, Canada\\
+  $^2$IBM Watson, USA\\
+  $^\dagger$CIFAR Fellow}
+\email{author@university.edu, coauthor@company.com}
 
 \begin{document}
-%\ninept
-%
+
 \maketitle
@@
 \begin{abstract}
-    Modern automatic speech recognition (ASR) systems need to be robust under acoustic variability arising from environmental, speaker, channel, and recording conditions.
-Ensuring such robustness to variability is a challenge in modern day neural network-based ASR systems, especially when all types of variability are not seen during training.
-We attempt to address this problem by encouraging the neural network acoustic model to learn invariant feature representations.
-    We use ideas from recent research on image generation using
-    Generative Adversarial Networks and domain adaptation ideas extending
-    adversarial gradient-based training. A recent work from Ganin et al. proposes to
-    use adversarial training for image domain adaptation by using an intermediate
-    representation from the main target classification network to deteriorate the domain 
-    classifier performance through a separate neural network.
-    Our work focuses on investigating neural architectures which produce
-    representations invariant to noise conditions for ASR.  We evaluate the proposed architecture on the Aurora-4 task, a popular benchmark for
-    noise robust ASR. We show that our method generalizes better than the standard multi-condition training especially when only a few noise categories are seen during training.
+Adversarial training for invariant representations for speech recognition
+
+Recent advances in domain adaptation allow us to construct networks that are
+invariant to certain labeled factors. The reverse gradient algorithm uses an
+adversarial learning approach to train a network to produce the desired label
+as well as to deceive a second classifier that is trained on adaptation 
+labels. This is tightly connected to the ideas behind generative adversarial 
+networks (GANs). Therefore, some insights for training GANs are needed for the
+reverse gradient's successful application. We adapt this approach to train an 
+acoustic system that is invariant to undesirable factors such as the recording
+environment noise type. We also investigate speaker identity invariance and 
+conduct experiments on two popular datasets.
 \end{abstract}
-%
-%\begin{keywords}
-%speech recognition, deep neural networks, invariant representations, adversarial training
-%\end{keywords}
-%
+\noindent\textbf{Index Terms}: speech recognition, adversarial training, 
+reverse gradient
 \section{Introduction}
@@
     One of the most challenging aspects of automatic speech recognition (ASR)
-    is the mismatch between the training and testing acoustic conditions. During
-    testing, a system may encounter new recording conditions, microphone types, speakers,
-    accents and types of background noises.
-Furthermore, even if the test scenarios are seen during training, there can be significant variability in their statistics.
-Thus, its important to develop ASR systems that are invariant to unseen acoustic conditions.
-    %bhuv
-    %Even in case all testing conditions have also appeared
-    %in the training data, their acoustic statistics likely to be
-    %different. For this reason, it's important to develop ASR systems that are
-    %robust towards such undesirable acoustic variabilities.
+    is the mismatch between the training and testing acoustic conditions. 
+    During testing, a system may encounter new recording conditions, 
+    microphone types, speakers, accents and types of background noises. 
+    Furthermore, even if the test scenarios are seen during training, there 
+    can be significant variability in their statistics. Thus, its important to 
+    develop ASR systems that are invariant to unseen acoustic conditions.
 
-    %While many types of model and feature based adaptation methods such as, maximum likelihood linear regression (MLLR) and feature-space MLLR (fMLLR)
-    %\cite{leggetter1995maximum} have been proposed, most of these methods were
-    %designed with Gaussian mixture model-hidden Markov model (GMM-HMM) systems in mind. Since the enormous success of
-        %model adaptation and feature-space adaptation are not optimal for the overall hybrid DNN-HMM system.
-    %methods Moreover, the discriminative gradient-based optimization used to train
-    %DNNs, in combination with the ability of DNNs to learn highly non-linear
-%feature transformations, allows for greater flexibility in
-    %constructing training objective functions that promote learning of noise invariant
-    %representations.
-    Several model and feature based adaptation methods such as Maximum Likelihood Linear Regression (MLLR),  feature-based MLLR 
-    and iVectors~\citep{saon2013speaker} have been proposed to handle speaker variability; and Noise Adaptive Training \citep[NAT;][]{kalinli2010noise}
-    and Vector Taylor Series \citep[VTS;][]{un1998speech} to handle environment variability. With the increasing success of Deep Neural Network (DNN) 
-    acoustic models for ASR \citep{hinton2012deep,seide2011conversational,sainath2011making}, end-to-end systems are being proposed~\citep{miao2015eesen,sainath2015learning} 
-    for modeling the acoustic conditions within a single network. This allows us to take advantage of the network's ability to learn 
-    highly non-linear feature transformations, with greater flexibility in constructing training objective functions that promote learning 
-    of noise invariant representations. The main idea of this work is to force the acoustic model to learn a representation invariant to 
-    noise conditions, instead of explicitly using noise robust acoustic features (Section~\ref{sec:invariant-speech}). This type of 
-    noise-invariant training requires noise-condition labels during training only. It is related to the idea of generative adversarial 
-    networks (GAN) and the gradient reverse method proposed in~\cite{goodfellow2014generative} and~\cite{ganin2014unsupervised} respectively 
-    (Section~\ref{sec:relatedwork}). We present results on the Aurora-4 speech recognition task in Section~\ref{sec:experiments} and summarize 
+    Several model and feature based adaptation methods such as Maximum 
+    Likelihood Linear Regression (MLLR),  feature-based MLLR and 
+    iVectors~\citep{saon2013speaker} have been proposed to handle speaker 
+    variability; and Noise Adaptive Training \citep[NAT;][]{kalinli2010noise}
+    and Vector Taylor Series \citep[VTS;][]{un1998speech} to handle 
+    environment variability. With the increasing success of Deep Neural 
+    Network~(DNN) acoustic models for ASR 
+    \citep{hinton2012deep,seide2011conversational,sainath2011making}, 
+    end-to-end systems are being 
+    proposed~\citep{miao2015eesen,sainath2015learning} for modeling the 
+    acoustic conditions within a single network. This allows us to take 
+    advantage of the network's ability to learn highly non-linear feature 
+    transformations, with greater flexibility in constructing training 
+    objective functions that promote learning of noise invariant 
+    representations. The main idea of this work is to force the acoustic model 
+    to learn a representation invariant to noise conditions, instead of 
+    explicitly using noise robust acoustic features 
+    (Section~\ref{sec:invariant-speech}). This type of noise-invariant 
+    training requires noise-condition labels during training only. It is 
+    related to the idea of generative adversarial networks (GAN) and the 
+    gradient reverse method proposed in~\cite{goodfellow2014generative} 
+    and~\cite{ganin2014unsupervised} respectively 
+    (Section~\ref{sec:relatedwork}). We present results on the Aurora-4 speech 
+    recognition task in Section~\ref{sec:experiments} and summarize 
     our findings in Section~\ref{sec:discussion}.
-
-    %Since this work is tightly connected to the idea
-    %f generative adversarial networks (GAN) proposed 
-    %in~\cite{goodfellow2014generative}, Section~\ref{sec:gans} presents an overview of GANs. Section~\ref{sec:gradient-reverse} then presents the gradient reverse method proposed in
-    %the context of image classification in~\cite{ganin2014unsupervised}. We then present the application 
-    %of this method to speech recognition
-    %in Section~\ref{sec:invariant-speech}. Section~\ref{sec:experiments} presents the development of an acoustic model learning invariant
-    %representations constructed similarly to the gradient reverse method
-    %from Section~\ref{sec:invariant-speech}. This section presents thorough experimentation
-    %on a publicly available Aurora-4 dataset for noisy speech recognition.
 
@@
 \emph{Generative Adversarial Networks} consist of two networks: generator and discriminator. 
-    %One of the most insightful recent developments in the area of generative models
-    %is the family of generative adversarial networks (GANs).
-    %A GAN consists of two networks: generator and discriminator. 
     The generator network $G$ has an
@@
     \begin{align*}
-        \min_G \max_D V(D, G) = \mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + 
-            \mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
+        \min_G \max_D V(D, G) = &\mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
+            &\mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
     \end{align*}
@@
     further explored in~\cite{salimans2016improved}.
-	%{\bf Please explain how the min-max applies to the two terms.
-I assume that we are maximizing first term with respect to D and doing min over D/max over G for the second term.
-Hence maximizing the first term increases the likelihood of the discriminator output (single sigmoid neuron) for true training data.
-Maximizing the second term decreases the likelihood of the discriminator output for fake training data.
-But then, what does the min over G mean?
-That you are maximizing the generator likelihood of the fake data?}
-    % Dima: deleted paragraph and explained min-max
 
-%\section{Gradient Reverse Method}
-%\label{sec:gradient-reverse}
-
-    %{\bf We need to write a couple of sentences on how this is motivated by GAN.} 
-    %Dima: added a sentence at the end of this paragraph
-
-    % bhuv :We should say how this is different form multi-task learning
     Prior work by~\cite{ganin2014unsupervised} proposed a method of training a network 
@@
     labeled with classes of interest and separate domain (image background) labels. 
-    %For example, the domain may refer to image background. The testing set does not need
-   % to have the domain labels. 
     The network has a $Y$-like structure: the image is fed to the
-    first network which produces a hidden representation $h$. Then this representation $h$ is input to two separate networks: a domain classifier network (D) and 
+    first network which produces a hidden representation $h$. Then this 
+    representation $h$ is input to two separate networks: a domain classifier network (D) and 
     a target classifier network (R). The goal of training is to learn the hidden 
-    representation that is invariant to the domain labels and performs well on the target classification task, so that the domain information doesn't 
-    interfere with the target classifier at test time. Similar to the GAN objective, which forces the generation distribution be close to the data distribution,
+    representation that is invariant to the domain labels and performs well on 
+    the target classification task, so that the domain information doesn't 
+    interfere with the target classifier at test time. Similar to the GAN 
+    objective, which forces the generation distribution be close to the data distribution,
     the \emph{gradient reverse method} makes domain distributions similar to each other.
@@
     where $y$ is the ground truth class, $d$ is the domain label, corresponding
-    hat variables are the network predictions, and $\theta_E, \theta_R$ and $\theta_D$ are the subsets of  parameters for the encoder,
+    hat variables are the network predictions, and $\theta_E, \theta_R$ and 
+    $\theta_D$ are the subsets of  parameters for the encoder,
     recognizer and the domain classifier networks respectively. The hyper-parameters
@@
 
-%% Denoising Auto encoder and
+    The influence of representations produced by a neural network to internal 
+    noise reduction is discussed in~\cite{yu2013feature} and this work sets a 
+    baseline for experiments on  Aurora-4 dataset. Recently, 
+    in~\cite{yusuke2016adversarial} a multilayer sigmoidal network is trained 
+    in an adversarial fashion on an in-house transcription task corrupted by noise.
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.4\linewidth]{model.pdf}
+    \caption{The model consists of three neural networks. The encoder $E$ produces
+    the intermediate representation $h$ which used in the recognizer $R$ and 
+    in the domain discriminator $D$. The hidden representation $h$ is trained to improve
+    the recognition and minimize the domain discriminator accuracy. The domain discriminator
+    is a classifier trained to maximize its accuracy on the noise type
+    classification task.}
+    \label{fig:model}
+\end{figure}[t]
 
-The influence of representations produced by a neural network to internal noise reduction is discussed in~\cite{yu2013feature} and this work sets a baseline for experiments on  Aurora-4 dataset.
-Recently, in~\cite{yusuke2016adversarial} a multilayer sigmoidal network is trained in an adversarial fashion on an in-house transcription task corrupted by noise.
-    %% STILL TODO
-    %{\bf Need to call out differences from this work: (1) We use the standard Aurora-4 training/testing conditions.
-(2) We also show the impact of seeing different number of noise types during training.
-(3) ...}
 \section{Invariant Representations for Speech Recognition}
@@
 
-Most ASR systems are DNN-HMM hybrid systems. The context dependent (CD) HMM states (acoustic model) are the class labels of interest. The
-recording conditions, speaker identity, or gender represent the domains in GANs. The task is to make the hidden layer representations of the HMM state classifier network 
+Most ASR systems are DNN-HMM hybrid systems. The context dependent (CD) HMM 
+states (acoustic model) are the class labels of interest. The
+recording conditions, speaker identity, or gender represent the domains in 
+GANs. The task is to make the hidden layer representations of the HMM state classifier network 
 invariant with respect to these domains. We hypothesize that this adversarial method of
@@
 
-Figure~\ref{fig:model} depicts the model, which is same as the model for the gradient reverse method.
-It is a feed-forward neural network trained to predict the CD HMM state, with a branch that predicts the domain (noise condition).
-This branch is discarded in the testing phase.
-In our experiments we
+Figure~\ref{fig:model} depicts the model, which is same as the model for the 
+gradient reverse method. It is a feed-forward neural network trained to predict 
+the CD HMM state, with a branch that predicts the domain (noise condition). This 
+branch is discarded in the testing phase. In our experiments we
 used the noise condition as the domain label merging all noise types into one label
@@
 For simplicity, we use only a single hyper-parameter -- the weight of the third term.
-
 \begin{figure}
-    \centering
-    \captionsetup[subfigure]{oneside,margin={0.3cm,0cm}}
-    \begin{subfigure}[b]{0.3\linewidth}
-        \centering
-        \includegraphics[width=0.9\linewidth]{model.pdf}
-        \caption{The model consists of three neural networks. The encoder $E$ produces
-        the intermediate representation $h$ which used in the recognizer $R$ and 
-        in the domain discriminator $D$. The hidden representation $h$ is trained to improve
-        the recognition and minimize the domain discriminator accuracy. The domain discriminator
-        is a classifier trained to maximize its accuracy on the noise type
-        classification task.}
-        \label{fig:model}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.7\linewidth}
     \centering
@@
     \label{fig:results}
-    \end{subfigure}
-    \caption{Model structure for invariant training and ASR results.}
 \end{figure}
-
-% Merged into earlier section
-%\section{Related Work}
- %   Generative adversarial networks were proposed and experimentally evaluated 
-  %  in~\cite{goodfellow2014generative}. This work was continued 
-   % in work of~\cite{radford2015unsupervised}, which proposes several practical guidelines
-    %for training GANs. This direction is further explored in~\cite{salimans2016improved}.
-
- %   The conditioned GAN for generation task were utilized in~\cite{denton2015deep}. This
-  %  work proposes to improve images increasing the resolution step-by-step. 
-   % The domain adaptation with the gradient reverse method was proposed in~\cite{ganin2014unsupervised}.
-
-    %TODO: cite denoising papers? some work on aurora 4?
-
-  %  The influence of representations produced by a neural network to internal noise reduction 
-  % is discussed in~\cite{yu2013feature} and this work sets a baseline for experiments on
-   % Aurora-4 dataset.
-
-   % In a work~\cite{yusuke2016adversarial} an acoustic model: multilayer sigmoidal network is 
-   % trained in an adversarial fashion for a task of dataset corrupted with additive noise.
-
-    %{\bf Need to call out differences from this work: (1) We use the standard Aurora-4 training/testing conditions.
-(2) We also show the impact of seeing different number of noise types during training.
-(3) ...}
-
-\section{Experiments}
-\label{sec:experiments}
-We experimentally evaluate our approach   
-on the well-benchmarked Aurora-4 \citep{parihar2002aurora} noisy speech recognition task. Aurora-4
-is based on the Wall Street Journal corpus (WSJ0). It contains noises of 
-six categories which was added to clean data. Every clean and noisy utterance is filtered to simulate the frequency characteristics.The training
-data  contains 4400 clean utterances and 446 utterances for each noise condition,
-i.e. a total of 2676 noisy utterances.
-The test set consists of clean data, data corrupted by 6 noise types, and data recorded with a different microphone for both clean and noisy cases.
-
-For both clean and noisy data, we extract 40-dimensional Mel-filterbank features with their deltas and 
-delta-deltas spliced over $\pm$5 frames, resulting in 1320 input 
-features that are subsequently mean and variance normalized. The baseline acoustic model is a 6-layer 
-DNN with 2048 rectified linear units at every layer. It is trained using momentum-accelerated stochastic gradient descent for 15 epochs with new-bob 
-annealing~\citep[as in][]{morgan1995continuous,sainath2011making}.
-
-%We annealed the learning rate according 
-%to the development set cross-entropy, dividing it by 2 when the cross entropy did not improve.
-
-In order to evaluate the impact of our method on generalization to unseen noises,
-we performed 6 experiments with different set of seen noises. The networks are trained
-on clean data, with each noise condition added one-by-one in the following order: airport, babble, car, 
-restaurant, street, and train. The last training group includes all noises therefore matches the
-standard multi-condition training setup. For every training group, we trained the
-baseline and the invariance model where we branch out at the $4^{th}$ layer to an  
-binary classifier predicting clean versus noisy data. Due to the imbalance between amounts of clean and
-noisy utterances, we had to oversample noisy frames to ensure that every mini-batch contained
-equal number of clean and noisy speech frames.
-
-Table~\ref{tab:results} summarizes the results.
-Figure~\ref{fig:results} visualizes the word error rate for the baseline multi-condition training and invariance training as the number of seen noise types varies.
-We conclude that the best performance
-gain is achieved when a small number of noise types are available during training.
-It can be seen that invariance training is able to generalize better to unseen noise types compared with multi-condition training.
-
-We note that our experiments did not use layer-wise pre-training, commonly used for small
-datasets. The baseline WERs reported are very close to the state-of-the-art. %due to the difficulties of combining it with the invariance training.
-Our preliminary experiments on a pre-trained network (better overall WER) when using all noise types (last row of Table~\ref{tab:results}) for training show the same trend as the non-pretrained networks.
-
-\begin{table}
+\begin{table*}[t]
     \centering
@@
     \end{tabular}
-    %TODO: add numbers!
-\end{table}
+\end{table*}
+
+\section{Experiments}
+\label{sec:experiments}
+We experimentally evaluate our approach   
+on the well-benchmarked Aurora-4 \citep{parihar2002aurora} noisy speech recognition task. Aurora-4
+is based on the Wall Street Journal corpus (WSJ0). It contains noises of 
+six categories which was added to clean data. Every clean and noisy utterance is 
+filtered to simulate the frequency characteristics.The training
+data  contains 4400 clean utterances and 446 utterances for each noise condition,
+i.e. a total of 2676 noisy utterances.
+The test set consists of clean data, data corrupted by 6 noise types, and data 
+recorded with a different microphone for both clean and noisy cases.
+
+For both clean and noisy data, we extract 40-dimensional Mel-filterbank features with their deltas and 
+delta-deltas spliced over $\pm$5 frames, resulting in 1320 input 
+features that are subsequently mean and variance normalized. The baseline acoustic model is a 6-layer 
+DNN with 2048 rectified linear units at every layer. It is trained using 
+momentum-accelerated stochastic gradient descent for 15 epochs with new-bob 
+annealing~\citep[as in][]{morgan1995continuous,sainath2011making}.
+
+In order to evaluate the impact of our method on generalization to unseen noises,
+we performed 6 experiments with different set of seen noises. The networks are trained
+on clean data, with each noise condition added one-by-one in the following order: airport, babble, car, 
+restaurant, street, and train. The last training group includes all noises therefore matches the
+standard multi-condition training setup. For every training group, we trained the
+baseline and the invariance model where we branch out at the $4^{th}$ layer to an  
+binary classifier predicting clean versus noisy data. Due to the imbalance between amounts of clean and
+noisy utterances, we had to oversample noisy frames to ensure that every mini-batch contained
+equal number of clean and noisy speech frames.
+
+Table~\ref{tab:results} summarizes the results. Figure~\ref{fig:results} visualizes 
+the word error rate for the baseline multi-condition training and invariance training 
+as the number of seen noise types varies. We conclude that the best performance
+gain is achieved when a small number of noise types are available during training. 
+It can be seen that invariance training is able to generalize better to unseen 
+noise types compared with multi-condition training.
+
+We note that our experiments did not use layer-wise pre-training, commonly used for small
+datasets. The baseline WERs reported are very close to the state-of-the-art. 
+Our preliminary experiments on a pre-trained network (better overall WER) when 
+using all noise types (last row of Table~\ref{tab:results}) for training show 
+the same trend as the non-pretrained networks.
 
@@
 \label{sec:discussion}
-    This paper presents the application of generative adversarial networks and invariance training for noise robust speech recognition. We show that invariance training 
-    helps the ASR system to generalize better to unseen noise conditions and improves word error rate when a small number of noise types are seen during training. Our 
-    experiments show that in contrast to the image recognition task, in speech recognition,  the domain adaptation network suffers from underfitting. Therefore, the 
-    gradient of the $L_3$ term in Eq.~\ref{eq:grm} is unreliable and noisy.
-Future research includes enhancements to the domain adaptation network while exploring alternative network architectures and invariance-promoting loss functions.
+    This paper presents the application of generative adversarial networks and 
+    invariance training for noise robust speech recognition. We show that invariance training 
+    helps the ASR system to generalize better to unseen noise conditions and improves 
+    word error rate when a small number of noise types are seen during training. Our 
+    experiments show that in contrast to the image recognition task, in speech 
+    recognition,  the domain adaptation network suffers from underfitting. Therefore, the 
+    gradient of the $L_3$ term in Eq.~\ref{eq:grm} is unreliable and noisy. Future 
+    research includes enhancements to the domain adaptation network while exploring 
+    alternative network architectures and invariance-promoting loss functions.
 
-\section*{Acknowledgments}
+\section{Acknowledgments}
 
 We would like to thank Yaroslav Ganin, David Warde-Farley for insightful discussions,
-developers of Theano~\cite{2016arXiv160502688short}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 
+developers of Theano~\cite{theano2016theano}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 
 for great toolkits.
 
-% References should be produced using the bibtex program from suitable
-% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
-% style file from IEEE produces unsorted bibliography list.
-% -------------------------------------------------------------------------
-%\bibliographystyle{IEEE}
-\bibliographystyle{authordate1}
+\bibliographystyle{IEEEtranN}
+
 \bibliography{refs}
-
 \end{document}

### 2017-03-21 interspeech/paper.tex
small typos/fixes and some comments

@@
 \begin{abstract}
-Adversarial training for invariant representations for speech recognition
-
 Recent advances in domain adaptation allow us to construct networks that are
@@
     Furthermore, even if the test scenarios are seen during training, there 
-    can be significant variability in their statistics. Thus, its important to 
+    can be significant variability in their statistics. Thus, it's important to 
     develop ASR systems that are invariant to unseen acoustic conditions.
@@
     and Vector Taylor Series \citep[VTS;][]{un1998speech} to handle 
-    environment variability. With the increasing success of Deep Neural 
+    environment variability.
+    % The end-to-end part has become sort of irrelevant given that the
+    % experiments are actually with hybrid systems and this version of the paper will not
+    % have to be sold as being end-to-end anymore.
+    With the increasing success of Deep Neural 
     Network~(DNN) acoustic models for ASR 
@@
     objective functions that promote learning of noise invariant 
-    representations. The main idea of this work is to force the acoustic model 
-    to learn a representation invariant to noise conditions, instead of 
+    representations.
+    % Phil: The part below should include speaker invariance and mention the WSJ task
+    The main idea of this work is to force the acoustic model 
+    to learn a representation which is invariant to noise conditions, instead of 
     explicitly using noise robust acoustic features 
@@
 
+% This is more of an explanation of the method rather than a literature review.
+    % Perhaps we should change the section title.
 \section{Related Work}
 \label{sec:relatedwork}
-\emph{Generative Adversarial Networks} consist of two networks: generator and discriminator. 
+\emph{Generative Adversarial Networks} consist of two networks: the generator and the discriminator. 
     The generator network $G$ has an
-    input of randomly-generated feature vectors and is asked to produce a sample, e.g. an image, 
-    similar to the images in the training set. The discriminator network $D$
+    input of randomly-generated feature vectors and is asked to produce a
+    sample, e.g.\ an image, similar to the images in the training set. The discriminator network $D$
     can either receive a generated image from the generator $G$ or an image
@@
     and can be trained with gradient backpropagation. This gradient can be propagated further
-    to the generator network.
+    to the generator network (assuming that the output of the generator is
+    continuous).
 
-    Two networks in the GAN setup are competing with each other: the 
+    The two networks in the GAN setup are competing with each other: the 
     generator is trying to deceive the discriminator network, while the discriminator tries
@@
     The maximization over the discriminator $D$ forms a usual cross-entropy objective, the gradients are
-    computed with respect to the parameters of $D$. The parameters of $G$ are minimized using the gradients
-    propagated through the second term. The minimization over $G$ makes it to produce examples which $D$
-    classifiers as the training ones.
-
+    computed with respect to the parameters of $D$. The parameters of $G$ are
+    minimizing the classification objective using the gradients
+    propagated through the second term. The minimization over $G$ makes it
+    produce samples which $D$ classifies as originating from the train data.
     Several practical guidelines were proposed for optimizing GANs in~\cite{radford2015unsupervised} and 
@@
 
-    Prior work by~\cite{ganin2014unsupervised} proposed a method of training a network 
+    Prior work by~\cite{ganin2014unsupervised} proposed a method for training a network 
     which can be adapted to new domains. The training data consists of the images
@@
     representation $h$ is input to two separate networks: a domain classifier network (D) and 
-    a target classifier network (R). The goal of training is to learn the hidden 
-    representation that is invariant to the domain labels and performs well on 
+    a target classifier network (R). The goal of training is to learn a hidden 
+    representation that is invariant to the domain labels and still performs well on 
     the target classification task, so that the domain information doesn't 
@@
     objective, which forces the generation distribution be close to the data distribution,
-    the \emph{gradient reverse method} makes domain distributions similar to each other.
+    the \emph{gradient reverse method} makes domain specific distributions similar to each other.
 
@@
 
-    The influence of representations produced by a neural network to internal 
+    The effectiveness of representations produced by a neural network for internal 
     noise reduction is discussed in~\cite{yu2013feature} and this work sets a 
-    baseline for experiments on  Aurora-4 dataset. Recently, 
-    in~\cite{yusuke2016adversarial} a multilayer sigmoidal network is trained 
+    baseline for experiments on the Aurora-4 dataset. Recently, 
+    in~\cite{yusuke2016adversarial} a multilayer sigmoidal network was trained 
     in an adversarial fashion on an in-house transcription task corrupted by noise.
@@
 training helps the HMM state classifier to generalize better to unseen domain conditions and requires only a  
-small additional amount of supervision, i.e. the domain labels.
+small additional amount of supervision, i.e., the domain labels.
 
-Figure~\ref{fig:model} depicts the model, which is same as the model for the 
+Figure \ref{fig:model} depicts the model, which is same as the model for the 
 gradient reverse method. It is a feed-forward neural network trained to predict 
-the CD HMM state, with a branch that predicts the domain (noise condition). This 
+the CD-HMM state, with a branch that predicts the domain (noise condition). This 
 branch is discarded in the testing phase. In our experiments we
-used the noise condition as the domain label merging all noise types into one label
-and clean as the other label. Our training loss function is Eq.~\ref{eq:grm} with 
+used the noise condition as the domain label, merging all noise types into one label
+and defining `clean' as the other label. Our training loss function is Eq.~\ref{eq:grm} with 
 $L_3$ set to $d\log(1 - \hat{d}) + (1-d)\log(\hat{d})$ for stability during training. 
@@
 For simplicity, we use only a single hyper-parameter -- the weight of the third term.
+
+% This should be done with subfig instead of saying Top/Bottom in the caption
 \begin{figure}
@@
     \includegraphics[width=\linewidth]{wer_seen_unseen.pdf}
-    \caption{Up: Average performance of the baseline multi-condition and invariance model varying with  the number of noise
+    \caption{Top: Average performance of the baseline multi-condition and invariance model varying with  the number of noise
         conditions used for training. Bottom: Average performance on seen versus unseen noise conditions.
@@
 \end{figure}
+
 \begin{table*}[t]
     \centering
-    \caption{Average word error rate (WER\%) on Aurora-4 dataset on all test conditions,
-        including seen and unseen noise and unseen microphone. First column
-        is the number of noise conditions used for the training. The last row is a 
-        preliminary experiment with layer-wise pre-training close to state-of-the-art
+    \caption{Average word error rates (WER\%) on Aurora-4 dataset on all test conditions,
+        including seen and unseen noise and unseen microphone. The first column
+        specifies the number of noise conditions used for the training. The
+        results in the last row are from a preliminary experiment with
+        layer-wise pre-training, close to state-of-the-art
         model and a corresponding invariance training starting with a pretrained model.}
@@
 is based on the Wall Street Journal corpus (WSJ0). It contains noises of 
-six categories which was added to clean data. Every clean and noisy utterance is 
-filtered to simulate the frequency characteristics.The training
+six categories which were added to the clean data. Every clean and noisy utterance is 
+% what kind of frequency characteristics?
+filtered to simulate the frequency characteristics. The training
 data contains 4400 clean utterances and 446 utterances for each noise condition,
-i.e. a total of 2676 noisy utterances.
+i.e., a total of 2676 noisy utterances.
 The test set consists of clean data, data corrupted by 6 noise types, and data 
-recorded with a different microphone for both clean and noisy cases.
+recorded with a different microphone for both the clean and noisy conditions.
 
-For both clean and noisy data, we extract 40-dimensional Mel-filterbank features with their deltas and 
+For both the clean and noisy data, we extract 40-dimensional Mel-filterbank features with their deltas and 
 delta-deltas spliced over $\pm$5 frames, resulting in 1320 input 
@@
 
-In order to evaluate the impact of our method on generalization to unseen noises,
+To evaluate the impact of our method on generalization to unseen noises,
 we performed 6 experiments with different set of seen noises. The networks are trained
@@
 standard multi-condition training setup. For every training group, we trained the
-baseline and the invariance model where we branch out at the $4^{th}$ layer to an  
+baseline and the invariance model where we branch out at the $4^{th}$ layer to a
 binary classifier predicting clean versus noisy data. Due to the imbalance between amounts of clean and
@@
 
-Table~\ref{tab:results} summarizes the results. Figure~\ref{fig:results} visualizes 
-the word error rate for the baseline multi-condition training and invariance training 
+Table~\ref{tab:results} summarizes the results. Figure \ref{fig:results} visualizes 
+the word error rate (WER) for the baseline multi-condition training and invariance training 
 as the number of seen noise types varies. We conclude that the best performance
@@
     helps the ASR system to generalize better to unseen noise conditions and improves 
-    word error rate when a small number of noise types are seen during training. Our 
+    the word error rate when a small number of noise types are seen during training. Our 
     experiments show that in contrast to the image recognition task, in speech 

### 2017-03-21 interspeech/paper.tex
tense consistency in experiments

@@
 \label{sec:experiments}
-We experimentally evaluate our approach   
+We experimentally evaluated our approach   
 on the well-benchmarked Aurora-4 \citep{parihar2002aurora} noisy speech recognition task. Aurora-4
 is based on the Wall Street Journal corpus (WSJ0). It contains noises of 
-six categories which were added to the clean data. Every clean and noisy utterance is 
+six categories which were added to the clean data. Every clean and noisy
+utterance has been 
 % what kind of frequency characteristics?
@@
 
-For both the clean and noisy data, we extract 40-dimensional Mel-filterbank features with their deltas and 
+For both the clean and noisy data, we extracted 40-dimensional Mel-filterbank features with their deltas and 
 delta-deltas spliced over $\pm$5 frames, resulting in 1320 input 
-features that are subsequently mean and variance normalized. The baseline acoustic model is a 6-layer 
-DNN with 2048 rectified linear units at every layer. It is trained using 
+features that were subsequently mean and variance normalized. The baseline
+acoustic model was a 6-layer 
+DNN with 2048 rectified linear units at every layer. It was trained using 
 momentum-accelerated stochastic gradient descent for 15 epochs with new-bob 
@@
 To evaluate the impact of our method on generalization to unseen noises,
-we performed 6 experiments with different set of seen noises. The networks are trained
+we performed 6 experiments with different sets of noises seen during training.
+The networks were trained
 on clean data, with each noise condition added one-by-one in the following order: airport, babble, car, 
-restaurant, street, and train. The last training group includes all noises therefore matches the
+restaurant, street, and train. The last training group included all noises and therefore matched the
 standard multi-condition training setup. For every training group, we trained the
-baseline and the invariance model where we branch out at the $4^{th}$ layer to a
+baseline and the invariance model, where we branched out at the $4^{th}$ layer to a
 binary classifier predicting clean versus noisy data. Due to the imbalance between amounts of clean and
@@
 
-Table~\ref{tab:results} summarizes the results. Figure \ref{fig:results} visualizes 
+Table \ref{tab:results} summarizes the results. Figure \ref{fig:results} visualizes 
 the word error rate (WER) for the baseline multi-condition training and invariance training 

### 2017-03-21 interspeech/paper.tex
add back unbreakable spaces

@@
 
-Figure \ref{fig:model} depicts the model, which is same as the model for the 
+Figure~\ref{fig:model} depicts the model, which is same as the model for the 
 gradient reverse method. It is a feed-forward neural network trained to predict 
@@
 
-Table \ref{tab:results} summarizes the results. Figure \ref{fig:results} visualizes 
+Table~\ref{tab:results} summarizes the results. Figure~\ref{fig:results} visualizes 
 the word error rate (WER) for the baseline multi-condition training and invariance training 

### 2017-03-21 interspeech/paper.tex
Edit introduction and background sections

@@
   $^\dagger$CIFAR Fellow}
-\email{author@university.edu, coauthor@company.com}
+\email{serdyuk@iro.umontreal.ca}
 
@@
 acoustic system that is invariant to undesirable factors such as the recording
-environment noise type. We also investigate speaker identity invariance and 
-conduct experiments on two popular datasets.
+environment noise type. We conduct experiments on a popular dataset.
 \end{abstract}
@@
     With the increasing success of Deep Neural 
-    Network~(DNN) acoustic models for ASR 
-    \citep{hinton2012deep,seide2011conversational,sainath2011making}, 
-    end-to-end systems are being 
-    proposed~\citep{miao2015eesen,sainath2015learning} for modeling the 
-    acoustic conditions within a single network. This allows us to take 
+    Network~(DNN) acoustic models for ASR investigated 
+    in~\citep{hinton2012deep,seide2011conversational,sainath2011making}, 
+    and in more recent works~\citep{miao2015eesen,sainath2015learning} 
+    complicated structure of acoustic conditions is modeled within a single network.
+    This allows us to take 
     advantage of the network's ability to learn highly non-linear feature 
@@
     and~\cite{ganin2014unsupervised} respectively 
-    (Section~\ref{sec:relatedwork}). We present results on the Aurora-4 speech 
+    (Section~\ref{sec:background}). Then we discuss previous work in 
+    Section~\ref{sec:relatedwork}. 
+    We present results on the Aurora-4 speech 
     recognition task in Section~\ref{sec:experiments} and summarize 
@@
     % Perhaps we should change the section title.
-\section{Related Work}
-\label{sec:relatedwork}
+\section{Background}
+\label{sec:background}
 \emph{Generative Adversarial Networks} consist of two networks: the generator and the discriminator. 
@@
     from the training set. Its task is to distinguish
-    between the ``fake'' generated image and the ``real'' image taken from the dataset. Thus,
+    between the ``counterfeit'' generated image and the ``genuine'' image taken from the dataset. Thus,
     the discriminator is just a classifier network with a sigmoid output layer
@@
     Formally, the objective function of GAN training is
-    \begin{align*}
+    \begin{align}
         \min_G \max_D V(D, G) = &\mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
             &\mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
-    \end{align*}
+        \label{eq:gan}
+    \end{align}
     The maximization over the discriminator $D$ forms a usual cross-entropy objective, the gradients are
-    computed with respect to the parameters of $D$. The parameters of $G$ are
+    computed with respect to the parameters of $D$. An important property of
+    this objective is that the gradient of the composition~$D(G(\cdot))$ is well
+    defined. Therefore the generator~$G$ can be trained with the backpropagation
+    algorithm using the gradient signal from the discriminator~$D$.
+    The generator~$G$ is
     minimizing the classification objective using the gradients
@@
     produce samples which $D$ classifies as originating from the train data.
+
     Several practical guidelines were proposed for optimizing GANs in~\cite{radford2015unsupervised} and 
-    further explored in~\cite{salimans2016improved}.
+    further explored in~\cite{salimans2016improved}. The second term of Eq.~\ref{eq:gan}
+    is frequently exchanged with
+    \begin{equation}
+        - \mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (D(G(\bm{z})))].
+    \end{equation}
+    The gradient of this term has the same sign but it has better properties.
+    We experimented with severall losses for our discriminator and chose one
+    similar to Eq.~\ref{eq:gan} but we maintain the negative sign in our notation
+    to demonstrate the general direction of the gradient of the discriminator loss.
 
@@
     \label{fig:model}
-\end{figure}[t]
+\end{figure}
+
+\section{Related Work}
+\label{sec:relatedwork}
+
+%TODO
 

### 2017-03-21 interspeech/paper.tex
Nicer figure
Various fixes

@@
     related to the idea of generative adversarial networks (GAN) and the 
-    gradient reverse method proposed in~\cite{goodfellow2014generative} 
+    gradient reverse method proposed by~\cite{goodfellow2014generative} 
     and~\cite{ganin2014unsupervised} respectively 
@@
 \label{sec:background}
+\begin{figure*}
+    \centering
+
+    \captionsetup[subfigure]{oneside,margin={0.3cm,0cm}}
+    \begin{subfigure}[b]{0.49\linewidth}
+    \centering
+        \includegraphics[width=0.4\linewidth]{model.pdf}
+        \caption{The model consists of three neural networks. The encoder $E$ produces
+        the intermediate representation $h$ which used in the recognizer $R$ and 
+        in the domain discriminator $D$. The hidden representation $h$ is trained to improve
+        the recognition and minimize the domain discriminator accuracy. The domain discriminator
+        is a classifier trained to maximize its accuracy on the noise type
+        classification task.}
+        \label{fig:model}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.5\linewidth}
+        \centering
+        \includegraphics[width=0.7\linewidth]{wer_avg.pdf}
+        \includegraphics[width=0.7\linewidth]{wer_seen_unseen.pdf}
+        \caption{Top: Average performance of the baseline multi-condition and invariance model varying with  the number of noise
+            conditions used for training. Bottom: Average performance on seen versus unseen noise conditions.
+            Testing was performed on all wv1 conditions (Sennheiser microphone).
+            }
+        \label{fig:results}
+    \end{subfigure}
+    \caption{The model and summarized results.}
+\end{figure*}
 \emph{Generative Adversarial Networks} consist of two networks: the generator and the discriminator. 
@@
 
-    Several practical guidelines were proposed for optimizing GANs in~\cite{radford2015unsupervised} and 
-    further explored in~\cite{salimans2016improved}. The second term of Eq.~\ref{eq:gan}
+    Several practical guidelines were proposed for optimizing GANs by~\cite{radford2015unsupervised} and 
+    further explored by~\cite{salimans2016improved}. The second term of Eq.~\ref{eq:gan}
     is frequently exchanged with
@@
 
-    The effectiveness of representations produced by a neural network for internal 
-    noise reduction is discussed in~\cite{yu2013feature} and this work sets a 
-    baseline for experiments on the Aurora-4 dataset. Recently, 
-    in~\cite{yusuke2016adversarial} a multilayer sigmoidal network was trained 
-    in an adversarial fashion on an in-house transcription task corrupted by noise.
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=0.4\linewidth]{model.pdf}
-    \caption{The model consists of three neural networks. The encoder $E$ produces
-    the intermediate representation $h$ which used in the recognizer $R$ and 
-    in the domain discriminator $D$. The hidden representation $h$ is trained to improve
-    the recognition and minimize the domain discriminator accuracy. The domain discriminator
-    is a classifier trained to maximize its accuracy on the noise type
-    classification task.}
-    \label{fig:model}
-\end{figure}
-
 \section{Related Work}
 \label{sec:relatedwork}
+    %TODO
+
+    The effectiveness of representations produced by a neural network for internal 
+    noise reduction is discussed by~\cite{yu2013feature} and this work sets a 
+    baseline for experiments on the Aurora-4 dataset. Recently, 
+    in a work by~\cite{yusuke2016adversarial} a multilayer sigmoidal network was trained 
+    in an adversarial fashion on an in-house transcription task corrupted by noise.
 
     %TODO
+    \cite{saon2017english} use a form of adversarial training with a network
+    that predicts an i-vector using the mean square error loss.
 
@@
 For simplicity, we use only a single hyper-parameter -- the weight of the third term.
-
-% This should be done with subfig instead of saying Top/Bottom in the caption
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{wer_avg.pdf}
-    \includegraphics[width=\linewidth]{wer_seen_unseen.pdf}
-    \caption{Top: Average performance of the baseline multi-condition and invariance model varying with  the number of noise
-        conditions used for training. Bottom: Average performance on seen versus unseen noise conditions.
-        Testing was performed on all wv1 conditions (Sennheiser microphone).
-        }
-    \label{fig:results}
-\end{figure}
 
@@
 momentum-accelerated stochastic gradient descent for 15 epochs with new-bob 
-annealing~\citep[as in][]{morgan1995continuous,sainath2011making}.
+annealing~\citep[half the learning rate if no improvement on the validation set, as in][]{morgan1995continuous,sainath2011making}.
 
@@
 We would like to thank Yaroslav Ganin, David Warde-Farley for insightful discussions,
-developers of Theano~\cite{theano2016theano}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 
+developers of Theano~\citep{theano2016theano}, Blocks, and Fuel~\citep{MerrienboerBDSW15} 
 for great toolkits.

### 2017-03-21 interspeech/paper.tex

@@
     the discriminator is just a classifier network with a sigmoid output layer
-    and can be trained with gradient backpropagation. This gradient can be propagated further
+    and can be trained with gradient back-propagation. This gradient can be propagated further
     to the generator network (assuming that the output of the generator is
@@
     this objective is that the gradient of the composition~$D(G(\cdot))$ is well
-    defined. Therefore the generator~$G$ can be trained with the backpropagation
+    defined. Therefore the generator~$G$ can be trained with the back-propagation
     algorithm using the gradient signal from the discriminator~$D$.
@@
     The gradient of this term has the same sign but it has better properties.
-    We experimented with severall losses for our discriminator and chose one
+    We experimented with several losses for our discriminator and chose one
     similar to Eq.~\ref{eq:gan} but we maintain the negative sign in our notation
@@
     baseline for experiments on the Aurora-4 dataset. Recently, 
-    in a work by~\cite{yusuke2016adversarial} a multilayer sigmoidal network was trained 
+    in a work by~\cite{yusuke2016adversarial} a multi-layer sigmoidal network was trained 
     in an adversarial fashion on an in-house transcription task corrupted by noise.

### 2017-03-21 interspeech/paper.tex
initial related work citations

@@
     %TODO
+    To be robust with respect to different recording conditions and speakers,
+    one can either aim to adapt the model parameters to these new situations or
+    to learn representations which are invariant to them.
+    Unfortunately, many of the adaptation methods which have been designed for
+    GMM-HMM systems cannot be applied to DNN-based systems. For this reason, a
+    large body of recent work has investigated new adaptation methods for neural
+    networks.
 
@@
     noise reduction is discussed by~\cite{yu2013feature} and this work sets a 
-    baseline for experiments on the Aurora-4 dataset. Recently, 
-    in a work by~\cite{yusuke2016adversarial} a multi-layer sigmoidal network was trained 
+    baseline for experiments on the Aurora-4 dataset.
+
+    Some of the linear and affine transformations used for GMM adaptation can
+    still be applied to neural networks when one limits these adaptations to the
+    very last layer, as was shown in~ \cite{yao2012adaptation}. The
+    improvements seemed to rely mostly on the adaptation of the bias parameters.
+
+    Many neural network speaker/condition adaptation methods are based on
+    i-vectors. One way to do this is by providing the i-vectors as additional
+    inputs \cite{senior2014improving}.
+
+    In the specific case of noise robustness, one can use speech enhancement to
+    obtain more robust features and neural networks have been used for this for
+    decades \cite{knecht1995neural}. A further step in this direction is to train
+    systems to recognize speech and perform speech enhancement jointly \cite{narayanan2014joint}.
+
+    Another strategy is to retrain the model parameters using a very small
+    number of utterances from the new domain while making sure that the model
+    doesn't stray too far away from the parameter values that were obtained from
+    the train data \cite{yu2013kl}.
+    The adaptation to new domains can also be kept under control by limiting
+    the number of parameters which are adapted to the new data or by limiting the
+    adaptation to a rescaling of the hidden unit activations
+    \cite{swietojanski2014learning}.
+
+    Recently, in a work by~\cite{yusuke2016adversarial} a multi-layer sigmoidal network was trained 
     in an adversarial fashion on an in-house transcription task corrupted by noise.

### 2017-03-21 interspeech/paper.tex
more text and some fixes

@@
     %TODO
-    To be robust with respect to different recording conditions and speakers,
+    Neural networks display some robustness towards different recording
+    conditions and speaker by themselves and the effectiveness of representations
+    produced by a neural network for internal 
+    noise reduction is discussed by~\cite{yu2013feature}. This work sets a 
+    baseline for experiments on the Aurora-4 dataset.
+
+    To be even more robust with respect to different recording conditions and speakers,
     one can either aim to adapt the model parameters to these new situations or
     to learn representations which are invariant to them.
+    Most approaches so far, are based on adaptation.
     Unfortunately, many of the adaptation methods which have been designed for
@@
 
-    The effectiveness of representations produced by a neural network for internal 
-    noise reduction is discussed by~\cite{yu2013feature} and this work sets a 
-    baseline for experiments on the Aurora-4 dataset.
-
     Some of the linear and affine transformations used for GMM adaptation can
     still be applied to neural networks when one limits these adaptations to the
-    very last layer, as was shown in~ \cite{yao2012adaptation}. The
+    very last layer, as was shown in~\cite{yao2012adaptation}. The
     improvements seemed to rely mostly on the adaptation of the bias parameters.
+    A clear downside of this approach is that it doesn't utilize the
+    representational power provided by the non-linear multi-layer structure of deep
+    neural networks.
 
-    Many neural network speaker/condition adaptation methods are based on
-    i-vectors. One way to do this is by providing the i-vectors as additional
-    inputs \cite{senior2014improving}.
+    Many neural network speaker adaptation methods are based on
+    i-vectors. The most common way to exploit i-vectors is by providing them as additional
+    inputs \cite{senior2014improving,saon2013speaker}. This means that i-vectors also need to
+    be available during testing and this may not always be practical.
 
@@
     systems to recognize speech and perform speech enhancement jointly \cite{narayanan2014joint}. 
+    We argue that speech enhancement is a more difficult task to learn than
+    invariance to certain noise conditions and more likely to suffer from
+    overfitting when the amount of available noisy data is limited.
 
-    Another strategy is to retrain the model parameters using a very small
+    Another adaptation strategy is to retrain the model parameters using a very small
     number of utterances from the new domain while making sure that the model
@@
     \cite{swietojanski2014learning}.
+    The most important difference between these approaches and our work is that
+    we don't adapt the model parameters at all after training. In many
+    situations this may not be possible and retraining of models may require more
+    from the hardware on which the speech recognition system is implemented.
 
@@
     in an adversarial fashion on an in-house transcription task corrupted by noise.
-
-    %TODO
-    \cite{saon2017english} use a form of adversarial training with a network
-    that predicts an i-vector using the mean square error loss.
+    This work is very similar to our approach but we evaluate our methods on
+    more challenging benchmark while investigating different numbers of noise
+    conditions. Our work also differs due to the use of more modern rectifier
+    activation based classification networks.
+    Finally, in other very recent work, \cite{saon2017english} a form of adversarial training with a network
+    that predicts an i-vector using the mean square error loss was used.
 

### 2017-03-21 interspeech/paper.tex
Some cosmetic changes

@@
     Likelihood Linear Regression (MLLR),  feature-based MLLR and 
-    iVectors~\citep{saon2013speaker} have been proposed to handle speaker 
+    i-vectors~\citep{saon2013speaker} have been proposed to handle speaker 
     variability; and Noise Adaptive Training \citep[NAT;][]{kalinli2010noise}
@@
 
+    % dima: I want to add something like this:
+    %Another point of view on the GAN training is that the generator is matching
+    %the distribution of the data with the distribution of the generator using
+    %a discriminator to provide the training signal. This approach can be generalized
+    %on any distributions to be matched.
     Prior work by~\cite{ganin2014unsupervised} proposed a method for training a network 
@@
 \label{sec:relatedwork}
-    %TODO
     Neural networks display some robustness towards different recording
@@
     i-vectors. The most common way to exploit i-vectors is by providing them as additional
-    inputs \cite{senior2014improving,saon2013speaker}. This means that i-vectors also need to
+    inputs~\citep{senior2014improving,saon2013speaker}. This means that i-vectors also need to
     be available during testing and this may not always be practical.
@@
     obtain more robust features and neural networks have been used for this for
-    decades \cite{knecht1995neural}. A further step in this direction is to train
-    systems to recognize speech and perform speech enhancement jointly \cite{narayanan2014joint}. 
+    decades~\citep{knecht1995neural}. A further step in this direction is to train
+    systems to recognize speech and perform speech enhancement jointly~\citep{narayanan2014joint}. 
     We argue that speech enhancement is a more difficult task to learn than
@@
     doesn't stray too far away from the parameter values that were obtained from
-    the train data \cite{yu2013kl}.
+    the train data~\citep{yu2013kl}.
     The adaptation to new domains can also be kept under control by limiting
     the number of parameters which are adapted to the new data or by limiting the
-    adaptation to a rescaling of the hidden unit activations
-    \cite{swietojanski2014learning}.
+    adaptation to a rescaling of the hidden unit activations~\citep{swietojanski2014learning}.
     The most important difference between these approaches and our work is that
@@
     activation based classification networks.
-    Finally, in other very recent work, \cite{saon2017english} a form of adversarial training with a network
+    Finally, in other very recent work~\citep{saon2017english} a form of adversarial training with a network
     that predicts an i-vector using the mean square error loss was used.

### 2017-03-21 interspeech/paper.tex

@@
 
-% This is more of an explanation of the method rather than a literature review.
-    % Perhaps we should change the section title.
 \section{Background}
@@
 
+    %TODO
     % dima: I want to add something like this:
@@
 momentum-accelerated stochastic gradient descent for 15 epochs with new-bob 
-annealing~\citep[half the learning rate if no improvement on the validation set, as in][]{morgan1995continuous,sainath2011making}.
+annealing~\citep[the learning rate is halfed if no improvement on the validation set, as in][]{morgan1995continuous,sainath2011making}.
 
+% TODO
+% dima: we need to emphasize that the unseen noises are very important
 To evaluate the impact of our method on generalization to unseen noises,
@@
 noise types compared with multi-condition training.
+% dima: we need to add that this is good news and very reasonable
 

### 2017-03-21 interspeech/paper.tex
first attempt

@@
     %on any distributions to be matched.
+    Optimization of the GAN objective corresponds to a minimization of the
+    so-called Jenson-Shannon divergence between the data distribution and the
+    distribution represented by the generator. This divergence is zero if and
+    only if the two distributions are identical. From this perspective, we can
+    say that the discriminator provides a training signal which allows
+    one to match distributions in general. In this work we are interested in
+    matching feature distributions of different noise conditions.
+
     Prior work by~\cite{ganin2014unsupervised} proposed a method for training a network 
@@
     overfitting when the amount of available noisy data is limited.
+    Another downside of enhancement approaches is that there is no reason to
+    expect them to generalize well to noise conditions that were not available during
+    training.
 
@@
 % what kind of frequency characteristics?
-filtered to simulate the frequency characteristics. The training
+filtered to simulate the phone quality recording using P.341 \citep{parihar2002aurora}. The training
 data contains 4400 clean utterances and 446 utterances for each noise condition,
@@
 % dima: we need to emphasize that the unseen noises are very important
-To evaluate the impact of our method on generalization to unseen noises,
+To evaluate the impact of our method on generalization to \emph{unseen} noises (the
+most typical situation in practice),
 we performed 6 experiments with different sets of noises seen during training.
@@
 noise types compared with multi-condition training.
+In practice, it's only possible to train on a small fraction of all the possible
+noise conditions in the world, so this apparent ability to generalize to unseen
+conditions is a promising result.
 % dima: we need to add that this is good news and very reasonable

### 2017-03-21 interspeech/paper.tex
Add more acknowledgements

@@
     labeled with classes of interest and separate domain (image background) labels. 
-    The network has a $Y$-like structure: the image is fed to the
+    The network has a $Y$-shaped structure: the image is fed to the
     first network which produces a hidden representation $h$. Then this 
-    representation $h$ is input to two separate networks: a domain classifier network (D) and 
-    a target classifier network (R). The goal of training is to learn a hidden 
+    representation $h$ is input to two separate networks: a domain classifier network ($D$) and 
+    a target classifier network ($R$). The goal of training is to learn a hidden 
     representation that is invariant to the domain labels and still performs well on 
@@
 utterance has been 
+% TODO!!!!
 % what kind of frequency characteristics?
@@
 
-% TODO
-% dima: we need to emphasize that the unseen noises are very important
 To evaluate the impact of our method on generalization to \emph{unseen} noises (the
@@
 for great toolkits. 
+The authors would like to acknowledge the support of the following agencies for
+research funding and computing support: NSERC, Calcul Qu\'{e}bec, Compute Canada,
+the Canada Research Chairs and CIFAR. Dmitriy Serdyuk thanks Samsung for partial
+financial support of this work.
 

### 2017-03-21 interspeech/paper.tex
Rearrange figure and add equation

@@
 \label{sec:background}
-\begin{figure*}
-    \centering
-
-    \captionsetup[subfigure]{oneside,margin={0.3cm,0cm}}
-    \begin{subfigure}[b]{0.49\linewidth}
-    \centering
-        \includegraphics[width=0.4\linewidth]{model.pdf}
-        \caption{The model consists of three neural networks. The encoder $E$ produces
-        the intermediate representation $h$ which used in the recognizer $R$ and 
-        in the domain discriminator $D$. The hidden representation $h$ is trained to improve
-        the recognition and minimize the domain discriminator accuracy. The domain discriminator
-        is a classifier trained to maximize its accuracy on the noise type
-        classification task.}
-        \label{fig:model}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.5\linewidth}
-        \centering
-        \includegraphics[width=0.7\linewidth]{wer_avg.pdf}
-        \includegraphics[width=0.7\linewidth]{wer_seen_unseen.pdf}
-        \caption{Top: Average performance of the baseline multi-condition and invariance model varying with  the number of noise
-            conditions used for training. Bottom: Average performance on seen versus unseen noise conditions.
-            Testing was performed on all wv1 conditions (Sennheiser microphone).
-            }
-        \label{fig:results}
-    \end{subfigure}
-    \caption{The model and summarized results.}
-\end{figure*}
 \emph{Generative Adversarial Networks} consist of two networks: the generator and the discriminator. 
@@
     \begin{align}
+    \begin{split}
         \min_G \max_D V(D, G) = &\mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
             &\mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
+    \end{split}
     \label{eq:gan}
@@
     the \emph{gradient reverse method} makes domain specific distributions similar to each other.
+\begin{figure}[t]
+    \centering
+
+    \centering
+        \includegraphics[width=0.4\linewidth]{model.pdf}
+        \caption{The model consists of three neural networks. The encoder $E$ produces
+        the intermediate representation $h$ which used in the recognizer $R$ and 
+        in the domain discriminator $D$. The hidden representation $h$ is trained to improve
+        the recognition and minimize the domain discriminator accuracy. The domain discriminator
+        is a classifier trained to maximize its accuracy on the noise type
+        classification task.}
+        \label{fig:model}
+\end{figure}
+\begin{figure}[t]
+    \begin{subfigure}[b]{\linewidth}
+        \centering
+        \includegraphics[width=1.1\linewidth]{wer_avg.pdf}
+        \caption{Average performance of the baseline multi-condition and invariance model varying with  the number of noise
+            conditions used for training.}
+        \label{fig:results}
+    \end{subfigure}
+    \begin{subfigure}[b]{\linewidth}
+        \centering
+        \includegraphics[width=1.1\linewidth]{wer_seen_unseen.pdf}
+        \caption{Average performance on seen versus unseen noise conditions.
+            Testing was performed on all wv1 conditions (Sennheiser microphone).
+            }
+    \end{subfigure}
+    \caption{Summarized results.}
+    \label{fig:results}
+\end{figure}
 
@@
 used the noise condition as the domain label, merging all noise types into one label
-and defining `clean' as the other label. Our training loss function is Eq.~\ref{eq:grm} with 
-$L_3$ set to $d\log(1 - \hat{d}) + (1-d)\log(\hat{d})$ for stability during training. 
-$L_3$ term maximizes the probability
+and defining `clean' as the other label. Our training loss function is  
+\begin{align}
+    \begin{split}
+    L = L_1(\hat{y}, y; \theta_R, \theta_E) + 
+    \alpha L_2(\hat{d}, d; \theta_D) -\\
+    \beta [d\log(1 - \hat{d}) + (1-d)\log(\hat{d})],
+    \end{split}
+    \label{eq:our}
+\end{align}
+where $d$ is the domain label and $\hat{d}$ is the predicted domain.
+The~$L_3$ term from Eq.~\ref{eq:grm} is defined according to the standard GAN
+objective for stability during training. 
+This term maximizes the probability
 of an incorrect domain classification in contrast to the gradient reverse where the 

### 2017-03-21 interspeech/paper.tex
Resolved
More acknowledgement

@@
 utterance has been 
-% TODO!!!!
-% what kind of frequency characteristics?
 filtered to simulate the phone quality recording using P.341 \citep{parihar2002aurora}. The training
@@
 the Canada Research Chairs and CIFAR. Dmitriy Serdyuk thanks Samsung for partial
-financial support of this work.
+financial support of this work and IBM Watson where he performed this work 
+during an internship.
 

### 2017-03-21 interspeech/paper.tex
some more intro and background

--- 2017-03-21 interspeech/paper.tex
@@
     representations.
-    % Phil: The part below should include speaker invariance and mention the WSJ task
+    At the same time, the more complex relation between the parameters of deep
+    neural networks and the functions or distributions they represent, makes it
+    impossible to simply use the same types of adaptation and robustness methods
+    that have been developed for GMM-based systems.
+
     The main idea of this work is to force the acoustic model 
@@
     gradient reverse method proposed by~\cite{goodfellow2014generative} 
-    and~\cite{ganin2014unsupervised} respectively 
-    (Section~\ref{sec:background}). Then we discuss previous work in 
-    Section~\ref{sec:relatedwork}. 
-    We present results on the Aurora-4 speech 
+    and~\cite{ganin2014unsupervised} respectively. 
+    First, we discuss these two methods in Section~\ref{sec:background}.
+    Then, we discuss previous work in Section~\ref{sec:relatedwork}. 
+    Finally, we present results on the Aurora-4 speech 
     recognition task in Section~\ref{sec:experiments} and summarize 
@@
 \end{figure*}
-\emph{Generative Adversarial Networks} consist of two networks: the generator and the discriminator. 
+\emph{Generative Adversarial Networks} are data generating neural networks that
+    have been especially successful at producing small images.
+    Unlike most generative models, they are not trained by optimizing the
+    likelihood of the data, but by the optimization of a so-called
+    \emph{adversarial} objective. The training of Generative Adversarial
+    Networks, involves two neural networks: the generator and the discriminator. 
     The generator network $G$ has an
-    input of randomly-generated feature vectors and is asked to produce a
-    sample, e.g.\ an image, similar to the images in the training set. The discriminator network $D$
+    input of randomly-generated feature vectors $\bm{z}$ and is asked to produce a
+    sample $\bm{x}$, e.g.\ an image, similar to the images in the training set. The discriminator network $D$
     can either receive a generated image from the generator $G$ or an image
@@
     \begin{align}
-        \min_G \max_D V(D, G) = &\mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
+        V(D, G) = &\mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
             &\mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
@@
     labeled with classes of interest and separate domain (image background) labels. 
-    The network has a $Y$-shaped structure: the image is fed to the
+    The network has a $Y$-like structure: the image is fed to the
     first network which produces a hidden representation $h$. Then this 
-    representation $h$ is input to two separate networks: a domain classifier network ($D$) and 
-    a target classifier network ($R$). The goal of training is to learn a hidden 
+    representation $h$ is input to two separate networks: a domain classifier network (D) and 
+    a target classifier network (R). The goal of training is to learn a hidden 
     representation that is invariant to the domain labels and still performs well on 
@@
 utterance has been 
-% TODO!!!!
 % what kind of frequency characteristics?
@@
 
+% TODO
+% dima: we need to emphasize that the unseen noises are very important
 To evaluate the impact of our method on generalization to \emph{unseen} noises (the
@@
 for great toolkits.
-The authors would like to acknowledge the support of the following agencies for
-research funding and computing support: NSERC, Calcul Qu\'{e}bec, Compute Canada,
-the Canada Research Chairs and CIFAR. Dmitriy Serdyuk thanks Samsung for partial
-financial support of this work.
 

### 2017-03-21 interspeech/paper.tex

--- 2017-03-21 interspeech/paper.tex
@@
     The generator network $G$ has an
-    input of randomly-generated feature vectors and is asked to produce a
-    sample, e.g.\ an image, similar to the images in the training set. The discriminator network $D$
+    input of randomly-generated feature vectors $\bm{z}$ and is asked to produce a
+    sample $\bm{x}$, e.g.\ an image, similar to the images in the training set. The discriminator network $D$
     can either receive a generated image from the generator $G$ or an image
@@
     \begin{split}
-        \min_G \max_D V(D, G) = &\mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
+        V(D, G) = &\mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
             &\mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
@@
 
-    %TODO
-    % dima: I want to add something like this:
-    %Another point of view on the GAN training is that the generator is matching
-    %the distribution of the data with the distribution of the generator using
-    %a discriminator to provide the training signal. This approach can be generalized
-    %on any distributions to be matched.
     Optimization of the GAN objective corresponds to a minimization of the
@@
 utterance has been 
-filtered to simulate the phone quality recording using P.341 \citep{parihar2002aurora}. The training
+filtered to simulate the phone quality recording using P.341 \citep{pariharsignal}. The training
 data contains 4400 clean utterances and 446 utterances for each noise condition,

### 2017-03-21 interspeech/paper.tex
Fix typo

@@
     to do its best to recognize if there was a deception, similar to adversarial game-theoretic settings.    
-    Formally, the objective function of GAN training is
+    Formally, the objective function of GAN training is $\min_G \max_D V(D, G)$ with
     \begin{align}
@@
 momentum-accelerated stochastic gradient descent for 15 epochs with new-bob 
-annealing~\citep[the learning rate is halfed if no improvement on the validation set, as in][]{morgan1995continuous,sainath2011making}.
+annealing~\citep[the learning rate is halved if no improvement on the validation set, as in][]{morgan1995continuous,sainath2011making}.
 

### 2017-03-21 interspeech/paper.tex
Various changes

@@
   $^\dagger$CIFAR Fellow}
-\email{serdyuk@iro.umontreal.ca}
+  \email{\{serdyuk,brakelp\}@iro.umontreal.ca, \{kaudhkha,bhuvana,sthomas\}@us.ibm.com}
 
@@
     \end{equation}
-    The gradient of this term has the same sign but it has better properties.
+    The gradient of this term has the same sign but it has better properties in
+    some cases.
     We experimented with several losses for our discriminator and chose one
@@
     Optimization of the GAN objective corresponds to a minimization of the
-    so-called Jenson-Shannon divergence between the data distribution and the
+    so-called Jensen-Shannon divergence between the data distribution and the
     distribution represented by the generator. This divergence is zero if and
@@
     the \emph{gradient reverse method} makes domain specific distributions similar to each other.
-\begin{figure}[t]
-    \centering
-
-    \centering
-        \includegraphics[width=0.4\linewidth]{model.pdf}
-        \caption{The model consists of three neural networks. The encoder $E$ produces
-        the intermediate representation $h$ which used in the recognizer $R$ and 
-        in the domain discriminator $D$. The hidden representation $h$ is trained to improve
-        the recognition and minimize the domain discriminator accuracy. The domain discriminator
-        is a classifier trained to maximize its accuracy on the noise type
-        classification task.}
-        \label{fig:model}
-\end{figure}
-\begin{figure}[t]
-    \begin{subfigure}[b]{\linewidth}
-        \centering
-        \includegraphics[width=1.1\linewidth]{wer_avg.pdf}
-        \caption{Average performance of the baseline multi-condition and invariance model varying with  the number of noise
-            conditions used for training.}
-        \label{fig:results}
-    \end{subfigure}
-    \begin{subfigure}[b]{\linewidth}
-        \centering
-        \includegraphics[width=1.1\linewidth]{wer_seen_unseen.pdf}
-        \caption{Average performance on seen versus unseen noise conditions.
-            Testing was performed on all wv1 conditions (Sennheiser microphone).
-            }
-    \end{subfigure}
-    \caption{Summarized results.}
-    \label{fig:results}
-\end{figure}
 
@@
 
-\begin{table*}[t]
+\begin{figure}[t]
     \centering
-    \caption{Average word error rates (WER\%) on Aurora-4 dataset on all test conditions,
-        including seen and unseen noise and unseen microphone. The first column
-        specifies the number of noise conditions used for the training. The
-        results in the last row are from a preliminary experiment with
-        layer-wise pre-training, close to state-of-the-art
-        model and a corresponding invariance training starting with a pretrained model.}
-    \label{tab:results}
-    \begin{tabular}{r|cc||cc|cc|cc|cc}
-    %\multirow{2}{*}{Subset} & \multicolumn{2}{c}{Orig}  & \multicolumn{2}{|c}{Inv}\\ 
-    %           & wv1  & wv2  & wv1  & wv2 \\
-        Noise       &Inv&BL&  \multicolumn{2}{c|}{A} & \multicolumn{2}{c|}{B} & \multicolumn{2}{c|}{C} & \multicolumn{2}{c}{D}\\
-               & & &  Inv & BL & Inv & BL & Inv & BL & Inv & BL\\
-    \hline
-    1           &16.36        &18.14 &6.54&7.57    &12.71& 14.09   & 11.45&   13.10    & 22.47 &   24.80    \\
-    2           &15.56        &17.39 &5.90&  6.58 &   11.69   &13.28   &11.12   &13.51   &21.79   &23.96 \\
-    3           &14.24        &14.67 &5.45 & 5.08&    10.76&   12.44&   9.75&    9.84 &   19.93&   19.30\\
-    4           &13.61        &13.84 & 5.08 &5.29    &9.73    &9.97    &9.49    &9.56    &19.49   &19.90\\         
-    5           &13.41        &13.02 & 5.12 &5.34    &9.52    &9.42    &9.55    &8.67    &19.33   &18.65\\         
-    6           &12.62        &12.60 & 4.80 &4.61    &9.04    &8.86    &8.76    &8.59    &18.16   &18.21\\
-    \hline\hline
-    6* &11.85        &11.99    &4.52    &4.76    &8.76    &8.76    &7.79    &8.57    &16.84&    16.99
-    \end{tabular}
-\end{table*}
+
+    \centering
+        \includegraphics[width=0.4\linewidth]{model.pdf}
+        \caption{The model consists of three neural networks. The encoder $E$ produces
+        the intermediate representation $h$ which used in the recognizer $R$ and 
+        in the domain discriminator $D$. The hidden representation $h$ is trained to improve
+        the recognition and minimize the domain discriminator accuracy. The domain discriminator
+        is a classifier trained to maximize its accuracy on the noise type
+        classification task.}
+        \label{fig:model}
+\end{figure}
+\begin{figure}[t]
+    \begin{subfigure}[b]{\linewidth}
+        \centering
+        \includegraphics[width=1.1\linewidth]{wer_avg.pdf}
+        \caption{Average performance of the baseline multi-condition and invariance model varying with  the number of noise
+            conditions used for training.}
+        \label{fig:results}
+    \end{subfigure}
+    \begin{subfigure}[b]{\linewidth}
+        \centering
+        \includegraphics[width=1.1\linewidth]{wer_seen_unseen.pdf}
+        \caption{Average performance on seen versus unseen noise conditions.
+            Testing was performed on all wv1 conditions (Sennheiser microphone).
+            }
+    \end{subfigure}
+    \caption{Summarized results.}
+    \label{fig:results}
+\end{figure}
 
@@
 the same trend as the non-pretrained networks.
+\begin{table*}[t]
+    \centering
+    \caption{Average word error rates (WER\%) on Aurora-4 dataset on all test conditions,
+        including seen and unseen noise and unseen microphone. The first column
+        specifies the number of noise conditions used for the training. The
+        results in the last row are from a preliminary experiment with
+        layer-wise pre-training, close to state-of-the-art
+        model and a corresponding invariance training starting with a pretrained model.}
+    \label{tab:results}
+    \begin{tabular}{r|cc||cc|cc|cc|cc}
+    %\multirow{2}{*}{Subset} & \multicolumn{2}{c}{Orig}  & \multicolumn{2}{|c}{Inv}\\ 
+    %           & wv1  & wv2  & wv1  & wv2 \\
+        Noise       &Inv&BL&  \multicolumn{2}{c|}{A} & \multicolumn{2}{c|}{B} & \multicolumn{2}{c|}{C} & \multicolumn{2}{c}{D}\\
+               & & &  Inv & BL & Inv & BL & Inv & BL & Inv & BL\\
+    \hline
+    1           &16.36        &18.14 &6.54&7.57    &12.71& 14.09   & 11.45&   13.10    & 22.47 &   24.80    \\
+    2           &15.56        &17.39 &5.90&  6.58 &   11.69   &13.28   &11.12   &13.51   &21.79   &23.96 \\
+    3           &14.24        &14.67 &5.45 & 5.08&    10.76&   12.44&   9.75&    9.84 &   19.93&   19.30\\
+    4           &13.61        &13.84 & 5.08 &5.29    &9.73    &9.97    &9.49    &9.56    &19.49   &19.90\\         
+    5           &13.41        &13.02 & 5.12 &5.34    &9.52    &9.42    &9.55    &8.67    &19.33   &18.65\\         
+    6           &12.62        &12.60 & 4.80 &4.61    &9.04    &8.86    &8.76    &8.59    &18.16   &18.21\\
+    \hline\hline
+    6* &11.85        &11.99    &4.52    &4.76    &8.76    &8.76    &7.79    &8.57    &16.84&    16.99
+    \end{tabular}
+\end{table*}
 
-\section{Discussion}
+\section{Conclusion}
 \label{sec:discussion}
@@
     experiments show that in contrast to the image recognition task, in speech 
-    recognition, the domain adaptation network suffers from underfitting. Therefore, the 
-    gradient of the $L_3$ term in Eq.~\ref{eq:grm} is unreliable and noisy. Future 
+    recognition, the task of the domain classification ($D$) network is harder. Therefore, the 
+    gradient of the $L_3$ term in Eq.~\ref{eq:grm} is noisier. Future 
     research includes enhancements to the domain adaptation network while exploring 

### 2017-03-21 interspeech/paper.tex
Return Phil's changes

@@
 \label{sec:background}
-\emph{Generative Adversarial Networks} consist of two networks: the generator and the discriminator. 
+\emph{Generative Adversarial Networks} are data generating neural networks that
+    have been especially successful at producing small images.
+    Unlike most generative models, they are not trained by optimizing the
+    likelihood of the data, but by the optimization of a so-called
+    \emph{adversarial} objective. The training of Generative Adversarial
+    Networks, involves two neural networks: the generator and the discriminator.
+
     The generator network $G$ has an

### 2017-03-21 interspeech/paper.tex
small change in conclusion

@@
     the word error rate when a small number of noise types are seen during training. Our 
-    experiments show that in contrast to the image recognition task, in speech 
-    recognition, the task of the domain classification ($D$) network is harder. Therefore, the 
+    experiments show that, relative to the image recognition tasks reported
+    in \citet{Yaroslav}, the task of the domain classification ($D$) network is
+    harder for speech recognition. Therefore, the 
     gradient of the $L_3$ term in Eq.~\ref{eq:grm} is noisier. Future 

### 2017-03-21 interspeech/paper.tex
fix citation

--- 2017-03-21 interspeech/paper.tex
@@
     the word error rate when a small number of noise types are seen during training. Our 
-    experiments show that in contrast to the image recognition task, in speech 
-    recognition, the task of the domain classification ($D$) network is harder. Therefore, the 
+    experiments show that, relative to the image recognition tasks reported
+    in \citet{ganin2014unsupervised}, the task of the domain classification ($D$) network is
+    harder for speech recognition. Therefore, the 
     gradient of the $L_3$ term in Eq.~\ref{eq:grm} is noisier. Future 

### 2017-03-21 interspeech/paper.tex
asfd

--- 2017-03-21 interspeech/paper.tex
@@
     \hline
-    1           &16.36        &18.14 &6.54&7.57    &12.71& 14.09   & 11.45&   13.10    & 22.47 &   24.80    \\
-    2           &15.56        &17.39 &5.90&  6.58 &   11.69   &13.28   &11.12   &13.51   &21.79   &23.96 \\
-    3           &14.24        &14.67 &5.45 & 5.08&    10.76&   12.44&   9.75&    9.84 &   19.93&   19.30\\
-    4           &13.61        &13.84 & 5.08 &5.29    &9.73    &9.97    &9.49    &9.56    &19.49   &19.90\\         
-    5           &13.41        &13.02 & 5.12 &5.34    &9.52    &9.42    &9.55    &8.67    &19.33   &18.65\\         
-    6           &12.62        &12.60 & 4.80 &4.61    &9.04    &8.86    &8.76    &8.59    &18.16   &18.21\\
+    1 condition           &16.36        &18.14 &6.54&7.57    &12.71& 14.09   & 11.45&   13.10    & 22.47 &   24.80    \\
+    2 conditions           &15.56        &17.39 &5.90&  6.58 &   11.69   &13.28   &11.12   &13.51   &21.79   &23.96 \\
+    3 conditions           &14.24        &14.67 &5.45 & 5.08&    10.76&   12.44&   9.75&    9.84 &   19.93&   19.30\\
+    4 conditions           &13.61        &13.84 & 5.08 &5.29    &9.73    &9.97    &9.49    &9.56    &19.49   &19.90\\         
+    5 conditions           &13.41        &13.02 & 5.12 &5.34    &9.52    &9.42    &9.55    &8.67    &19.33   &18.65\\         
+    6 conditions           &12.62        &12.60 & 4.80 &4.61    &9.04    &8.86    &8.76    &8.59    &18.16   &18.21\\
     \hline\hline
-    6* &11.85        &11.99    &4.52    &4.76    &8.76    &8.76    &7.79    &8.57    &16.84&    16.99
+    6* conditions &11.85        &11.99    &4.52    &4.76    &8.76    &8.76    &7.79    &8.57    &16.84&    16.99\\
+    \hline\hline
+     Yu et al. (2013a)  &\multicolumn{2}{c|}{13.4} & \multicolumn{2}{c|}{5.6} & \multicolumn{2}{c|}{8.8} &\multicolumn{2}{c|}{8.9}&\multicolumn{2}{c}{20.0}
     \end{tabular}

### 2017-03-21 interspeech/paper.tex

@@
     the word error rate when a small number of noise types are seen during training. Our 
-    experiments show that in contrast to the image recognition task, in speech 
-    recognition, the task of the domain classification ($D$) network is harder. Therefore, the 
+    experiments show that, relative to the image recognition tasks reported
+    by \citet{ganin2014unsupervised}, the task of the domain classification ($D$) network is
+    harder for speech recognition. Therefore, the 
     gradient of the $L_3$ term in Eq.~\ref{eq:grm} is noisier. Future 

### 2017-03-22 interspeech/paper.tex
Change citation style

--- 2017-03-21 interspeech/paper.tex
@@
 
-\usepackage{natbib}
 \usepackage{INTERSPEECH_v2}
@@
     Likelihood Linear Regression (MLLR),  feature-based MLLR and 
-    i-vectors~\citep{saon2013speaker} have been proposed to handle speaker 
-    variability; and Noise Adaptive Training \citep[NAT;][]{kalinli2010noise}
-    and Vector Taylor Series \citep[VTS;][]{un1998speech} to handle 
+    i-vectors~\cite{saon2013speaker} have been proposed to handle speaker 
+    variability; and Noise Adaptive Training~(NAT;~\cite{kalinli2010noise})
+    and Vector Taylor Series~(VTS;~\cite{un1998speech}) to handle 
     environment variability.
@@
     Network~(DNN) acoustic models for ASR investigated 
-    in~\citep{hinton2012deep,seide2011conversational,sainath2011making}, 
-    and in more recent works~\citep{miao2015eesen,sainath2015learning} 
+    in~\cite{hinton2012deep}\cite{seide2011conversational}\cite{sainath2011making}, 
+    and in more recent works~\cite{miao2015eesen}\cite{sainath2015learning} 
     complicated structure of acoustic conditions is modeled within a single network.
@@
 
+    This work is a continuation of our preliminary results presented as a workshop 
+    paper~\cite{serdyuk2016invariant}.
     The main idea of this work is to force the acoustic model 
@@
     i-vectors. The most common way to exploit i-vectors is by providing them as additional
-    inputs~\citep{senior2014improving,saon2013speaker}. This means that i-vectors also need to
+    inputs~\cite{senior2014improving,saon2013speaker}. This means that i-vectors also need to
     be available during testing and this may not always be practical.
@@
     obtain more robust features and neural networks have been used for this for
-    decades~\citep{knecht1995neural}. A further step in this direction is to train
-    systems to recognize speech and perform speech enhancement jointly~\citep{narayanan2014joint}. 
+    decades~\cite{knecht1995neural}. A further step in this direction is to train
+    systems to recognize speech and perform speech enhancement jointly~\cite{narayanan2014joint}. 
     We argue that speech enhancement is a more difficult task to learn than
@@
     doesn't stray too far away from the parameter values that were obtained from
-    the train data~\citep{yu2013kl}.
+    the train data~\cite{yu2013kl}.
     The adaptation to new domains can also be kept under control by limiting
     the number of parameters which are adapted to the new data or by limiting the
-    adaptation to a rescaling of the hidden unit activations~\citep{swietojanski2014learning}.
+    adaptation to a rescaling of the hidden unit activations~\cite{swietojanski2014learning}.
     The most important difference between these approaches and our work is that
@@
     activation based classification networks.
-    Finally, in other very recent work~\citep{saon2017english} a form of adversarial training with a network
+    Finally, in other very recent work~\cite{saon2017english} a form of adversarial training with a network
     that predicts an i-vector using the mean square error loss was used.
@@
 We experimentally evaluated our approach   
-on the well-benchmarked Aurora-4 \citep{parihar2002aurora} noisy speech recognition task. Aurora-4
+on the well-benchmarked Aurora-4 \cite{parihar2002aurora} noisy speech recognition task. Aurora-4
 is based on the Wall Street Journal corpus (WSJ0). It contains noises of 
@@
 utterance has been 
-filtered to simulate the phone quality recording using P.341 \citep{pariharsignal}. The training
+filtered to simulate the phone quality recording using P.341 \cite{pariharsignal}. The training
 data contains 4400 clean utterances and 446 utterances for each noise condition,
@@
 momentum-accelerated stochastic gradient descent for 15 epochs with new-bob 
-annealing~\citep[the learning rate is halved if no improvement on the validation set, as in][]{morgan1995continuous,sainath2011making}.
+annealing~\cite[the learning rate is halved if no improvement on the validation set, as in][]{morgan1995continuous,sainath2011making}.
 
@@
     the word error rate when a small number of noise types are seen during training. Our 
-    experiments show that in contrast to the image recognition task, in speech 
-    recognition, the task of the domain classification ($D$) network is harder. Therefore, the 
+    experiments show that, relative to the image recognition tasks reported
+    by \cite{ganin2014unsupervised}, the task of the domain classification ($D$) network is
+    harder for speech recognition. Therefore, the 
     gradient of the $L_3$ term in Eq.~\ref{eq:grm} is noisier. Future 
@@
 We would like to thank Yaroslav Ganin, David Warde-Farley for insightful discussions,
-developers of Theano~\citep{theano2016theano}, Blocks, and Fuel~\citep{MerrienboerBDSW15} 
+developers of Theano~\cite{theano2016theano}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 
 for great toolkits. 
@@
 
-\bibliographystyle{IEEEtranN}
+\bibliographystyle{IEEEtran}
 
